### 2022-11-24 doc/paper.tex

\documentclass[onecolumn,10pt,cleanfoot]{asme2ej}

\usepackage{graphicx} %% for loading jpg figures
\usepackage{bm}
\usepackage{nicefrac}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{listings}
\usepackage{tablefootnote}
\usepackage{float}
\usepackage{xcolor}
\usepackage{xurl}

\title{Dimention reduction for strong vs weak learners in pneumonia X-ray image prediction}

%%% first author
\author{Jonatan H. Hanssen
    \affiliation{
	Bachelor Student, Robotics and \\
	Intelligent Systems\\ \\[-10pt]
	Department of Informatics\\ \\[-10pt]
	The faculty of Mathematics and \\
	Natural Sciences\\ \\[-10pt]
    Email: jonatahh@ifi.uio.no
    }
}

\author{Eric E. Reber
    \affiliation{
	Bachelor Student, Robotics and \\
	Intelligent Systems\\ \\[-10pt]
	Department of Informatics\\ \\[-10pt]
	The faculty of Mathematics and \\
	Natural Sciences\\ \\[-10pt]
    Email: ericer@ifi.uio.no
    }
}

\author{Gregor Kajda
    \affiliation{
	Bachelor Student, Robotics and \\
	Intelligent Systems\\ \\[-10pt]
	Department of Informatics\\ \\[-10pt]
	The faculty of Mathematics and \\
	Natural Sciences\\ \\[-10pt]
    Email: grzegork@ifi.uio.no
    }
}

\begin{document}

\maketitle

\section{Abstract}

\section{Introduction}
Images provide a challenge in machine learning as there are a large number of features present in the data which slow down our training. This is usually solved by first reducing the dimentionality of our data before feeding it into our learner of choice. In this paper, we will explore how convolution stacks up against principle component analysis as a method of dimention reduction for image data when coupled with a strong or a weak learner. We will investigate how these architectures perform for a simple image classification problem such as MNIST versus a more difficult classification problem in the form of pneumonia prediction from X-ray images.

More specifically, we will compare a traditional convolutional neural network with an architecture of a convolutional stage fed into a random forest, and against an architecture consisting of a PCA stage fed into either a neural network or a random forest with the aim of decreasing the computational cost or increasing the accuracy accross different datasets. Our investigation can be simplified into one question: what is the best way to tackle the challenges associated with image classification?

First, we will give an explanation of the theory and method used in this paper, followed up with a detailed discussion of our results. Finally, we will arrive at a conclusion which summarizes our core results and lessons.

\section{Method}

\subsection{Theory}

### 2022-11-24 doc/paper.tex
Add paper files
Added theory to paper

@@
 \usepackage{xurl}
-
-\title{Dimention reduction for strong vs weak learners in pneumonia X-ray image prediction}
 
@@
 
+% \title{Dimention reduction for strong vs weak learners in pneumonia X-ray image prediction}
+% \title{Liminations of computing compunations using tensorflow implemenation}
+\title{Hybrid models as an alternative to traditional Convolutional Neural Networks for image classification}
+
 \maketitle
@@
 
+We found a dog driving a car down Hollywood Boulevard.
+It had sunglasses on, but didn't know where to go due to demensionality reduction.
+It was computing many computinations but due to severe liminations it had problems with binaly classifination and a large emesee.
+
 \section{Introduction}
-Images provide a challenge in machine learning as there are a large number of features present in the data which slow down our training.
-This is usually solved by first reducing the dimentionality of our data before feeding it into our learner of choice.
-In this paper, we will explore how convolution stacks up against principle component analysis as a method of dimention reduction for image data when coupled with a strong or a weak learner.
-We will investigate how these architectures perform for a simple image classification problem such as MNIST versus a more difficult classification problem in the form of pneumonia prediction from X-ray images.
 
-More specifically, we will compare a traditional convolutional neural network with an architecture of a convolutional stage fed into a random forest, and against an architecture consisting of a PCA stage fed into either a neural network or a random forest with the aim of decreasing the computational cost or increasing the accuracy accross different datasets.
+The task of image recognition and classification is a great challenge in machine learning, as there are a large number of features present in the data which slow down our training.
+For a relatively small one-channel image of $200$ by $200$ pixels, the naive approach of using every pixel as a feature becomes unfeasable, as we now have $40000$ features.
+Clearly, some form of dimension reduction must be used to rectify this problem, and Convolutional Neural Networks are regarded as the gold standard for this.
+CNNs are able to reduce the dimensionality while extracting useful information in each layer, for example by finding edges in the first layer, combinations of edges that make a shape in the next layer, and combinations of shapes that resemble an object in the last layer.
+However, CNNs have many hyperparameters that need to be adjusted to achieve optimal performance, and a slow and computationally intensive training process.
+In this paper, we explore alternative models for image classification, using Principal Component Analysis and Random Forests in combination with parts of the traditional CNN model.
+We will investigate how these architectures perform for a simple image classification problem such as the MNIST dataset versus a more difficult classification problem in the form of pneumonia prediction from X-ray images.
+
+More specifically, we will compare four different architectures.
+First we will use a traditional CNN, where a series convolutional layers is followed by a series fully connected layers.
+Then we will modify our CNN to create a CNN-RF model, where the features extracted by the convolutional layers are fed into a Random Forest instead of a series of fully connected layers.
+Finally, we will make use of Principal Component Analysis instead of convolution for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest.
+In doing this, we hope to answer the following question: what is the best way to tackle the challenges associated with image classification?
+
+% More specifically, we will compare a traditional convolutional neural network with an architecture of a convolutional stage fed into a random forest, and against an architecture consisting of a PCA stage fed into either a neural network or a random forest with the aim of decreasing the computational cost or increasing the accuracy accross different datasets.
 Our investigation can be simplified into one question: what is the best way to tackle the challenges associated with image classification?
@@
 \subsection{Theory}
+
+This sections covers the theoretic background which underlies the methods used in this paper.
+
+\subsubsection{Convolutional Neural Networks}
+
+A convolutional neural network.
+
+\subsubsection{Principal Component Analysis}
+
+Principal Component Analysis is a dimensionality reduction method which projects our dataset onto a lower dimension, while capturing as much of the variance of our data as possible.
+The central idea is to change the basis we use to represent our dataset, using vectors pointing in the principal directions of our dataset, instead of the unit vectors in feature space.
+The principal directions point in the direction of the variance of our dataset.
+By simply changing our basis to these vectors, called principal components, we have not achieved any dimensionality reduction, but by discarding directions of very little variance, we can reduce dimensionality without losing much variance (and thus information).
+By performing a Singular Value Decomposition we have everything we need to achieve this.
+The SVD is as follows:
+
+\begin{equation}
+X = UDV^T
+\end{equation}
+
+Here, the coloumns $U$ is known as the left singular vectors of $X$, $D$ is a diagonal matrix containing the singular values of $X$ (the square roots of the eigenvalues of $X^TX$) and the coloumns in $V$ are known as the right singular values.
+The right singular values are actually the principal components we need, ordered by the variance in that direction.
+Thus, we can project our dataset down to a subspace spanned by a subset of these vectors (for example the first 3 vectors) and reduce the dimensionality of our dataset while losing as little information as possible.
+Furthermore, the coloumn vectors of $V$ are ortogonal, and as such we can project our dataset down to this subspace with a single matrix multiplication:
+
+\begin{equation}
+\hat{X} = XV
+\end{equation}
+
+% Principal Component Analysis is a dimensionality reduction method performs a change of basis into a lower dimensional subspace of our feature space, while capturing as much of the variance of our dataset as possible.
+To do this, we can zero-center our datThis can be done by performing a Singular Value Decomposition of the design matrix, yielding the following decomposition.
+
+\subsubsection{Decision Trees}
+
+A decision tree is an intiutive way of performing both classification and regression tasks, by asking a series of yes or no questions about an instance in the dataset.
+These question will split the original dataset, the root node, into child nodes which contain subsets of the dataset, and these nodes will be split again, creating a tree structure which terminates in leaf nodes which imply a choice of class.
+
+In a binary classification problem, the algorithm for creating a decision tree is quite simple.
+First, consider each feature of the dataset, and attempt to split all instances into two child nodes based on a simple true/false statement about this feature.
+Having done this for every feature, calculate the degree of homogeneity\footnote{This can described by calculating the Gini-factor, or by calculating the entropy of the set} of the two child nodes created, when considering the actual class of the instances.
+The feature which created the most homegenous split, is the feature which has the largest effect on the final class, and as such should be the first feature we use to split the dataset.
+After this, we have two subsets, and we perform the same algorithm on these subsets again.
+This algorithm terminates when it is not possible to split the data further, because all nodes only contain instances belonging to the same class.
+In practice, it is advised to stop the algorithm earlier by introducing limits to the depth of the tree, or limits to how few instances can be in one node.
+Once the algorithm has terminated, we have a series of leaf nodes containing subsets of the data.
+
+Once a tree has been built, prediction is done by taking the instance and propagating it down the tree by answering the true/false statements that have been decided on in the creation of the tree.
+Once we have reached a leaf node, we predict that this instance is the class to which the majority of the instances in the leaf node belong to.
+
+Decision trees are intuitive and computationally cheap to create, not requiring any training due to the deterministic nature of the algorithm.
+Furthermore, they require little tuning and have few hyperparameters.
+However, they are prone to overfitting.
+
+\subsubsection{Random Forests}
+
+A Random Forest is a type of ensemble method used to improve the performance of a single decision tree, by using bagging.
+Instead of using a single tree, a random forest consists of many similar trees, and the final class is determined majority voting by all the trees.
+Like decision trees, they are relatively simple to understand.
+
+To create a random forest, we perform a bootstrap of the original dataset, creating many similar instances with small permutations introduced by resampling with replacement.
+We now create one decision tree per bootstrap, only using a random subset of the total amount of features every time we perform a split\footnote{The number of features used is often set to $\sqrt{p}$}.
+The result of this is that we create many slightly different trees, each slightly suboptimal in their splits.
+These trees give predictions which are worse than simply creating a single decision tree using all the features and all the data, but when basing our prediction of the majority vote of all the trees, we are able to make predictions which are more generalizable and less prone to overfitting.
+
+\subsubsection{Datasets}
+
+Below is a description of each of the datasets used in this report.
+
+\subsubsection{MNIST}
+
+The MNIST dataset is well known multiclass classification problem used to test machine learning algorithms.
+It consists of $70000$ black and white images of hand drawn digits, each $28$ by $28$ pixels.
+Each image is labeled with one of ten classes, corresponding to the digit it depicting.
+The dataset is divided into $60000$ training images and $10000$ validation images.This dataset is essentially uniformly distributed, with all classes containing around $9 - 11\%$ of the total dataset.
+
+\subsubsection{Guangzhou Chest X-Ray Pneumonia Dataset}
+
+The Guangzhou Chest X-Ray Pneumonia Dataset consists of $5856$ X-Ray images of children between the ages of one to five.
+The images were taken from the Woman and Children's Medical Center in Guangzhou, China.
+The dataset is divided into a training set of $5232$ images and $624$.
+The dataset is not evenly distributed, having $74.2\%$ instances of pneumonia in the training set and $62.5\%$ instances of pneumonia in the validation set.
+
+\subsection{Implementation}
+
+Our implementation can be found at \texttt{github.com/Gregz9/CNN\_vs\_weak.git}.
+
+All neural network stages were implemented using the Tensorflow library.
+PCA was implemented as a custom layer using the Keras API, or by mapping the operation to the whole dataset using Tensorflow's \texttt{dataset.map} functionality, depending on the model.
+Random Forests were implemented using the Tensorflow Decision Forests library.
+
+\section{Results}
+
+We found so much stuff man
+
+\subsection{Traditional Convolutional Neural Network}
+
+\subsection{Convolution followed by a Random Forest}
+
+\subsection{Principal Component Analysis followed by a Neural Network}
+
+\subsection{Principal Component Analysis followed by a Random Forest}
+
+\section{Conclusion}
+
+I have realized that I am cooler than Eric.
+
+\section*{Appendix}
+
+Here are some pictures of Eric Reber, formerly known as the criminal Ed "Gurgler" Bickley of Texas State Penistentiary.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=3.25in]{/home/jona/pictures/cat.jpg}}
+\caption{Eric Reber in his most devious form}
+\label{real1msetraintest}
+\end{figure}
+
+\end{document}

### 2022-11-24 doc/paper.tex
Add more theory

@@
 
-A convolutional neural network.
+Convolutional Neural Networks are neural networks which are specialized to work on images, or other types of data where the structure of the data carries meaning.
+A normal neural network treats each input indepentently, and inputs are flattened such that any spatial information is lost.
+CNNs, on the other hand, explicitly take the spatial dimensions into account, by using convolutional kernels and pooling layers as part of their architecture.
+
+The eponymous operation of convolution is a method which is used much in image processing.
+It is performed by using a convolutional kernel, which is applied to each pixel of the image.
+The kernel is a matrix of values, and the output of a particular pixel is the sum of all the pixels in its neighbourhood multiplied by the overlapping value in the kernel.
+As such, the value in any one position is dependent on the kernel, but also on the pixels around this position.
+This allows us to extract valuable information about images, such as edges or patterns.
+A basic kernel could be the Laplacian kernel (Eq.~\ref{laplace}), which will output a high value when the pixel in the middle of the kernel has a value which differs greatly from the pixels in a 3 by 3 neighbourhood around it.
+
+\begin{equation}
+\label{laplace}
+laplace : 
+\begin{bmatrix}
+-1 & -1 & -1 \\
+-1 & 8 & -1 \\
+-1 & -1 & -1
+\end{bmatrix}
+\end{equation}
+
+The mathematical equation for the convolution of a kernel $w$ on an image $f(x,y)$ (represented as a discrete function in two variables) is as follows:
+
+\begin{equation}
+w \star f(x,y) = \sum^a_{s=-a} \sum^b_{t=-b}f(x+s, y+t)\footnote{In reality, we should have $x-s$ and $y-t$, as correlation involves rotating the kernel by 180 degrees.
+However, the term convolution is often used for both the mathematical operations of correlation and convolution\cite[160]{dip}}
+\end{equation}
+
+In a convolutional neural network, the values of each kernel is initialized randomly, and training involves tuning these values.
+When performing convolution at the edges of an image, some values will be undefined.
 
@@
 
+\bibliographystyle{apalike}
+\bibliography{bibliography}
+
 \section*{Appendix}

### 2022-11-25 src/conv_nn_pneu.py
Added conv rf for pneumonia. Not good results

@@ -28,8 +28,6 @@ test_ds = tf.keras.utils.image_dataset_from_directory(
     color_mode="grayscale",
 )
 
-normalization_layer = layers.Rescaling(1.0 / 255)
-
 AUTOTUNE = tf.data.AUTOTUNE
 
 train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)

### 2022-11-25 src/conv_rf_mnist.py

@@ -56,6 +56,7 @@ train_features = feature_extractor(x_train[:10000])
 test_features = feature_extractor(x_test)
 
 forest = RandomForestClassifier(random_state=1337)
+forest.compile(metrics=["accuracy"])
 
 forest.fit(train_features, y_train[:10000])
 

### 2022-11-25 src/conv_rf_pneu.py

@@ -1,0 +1,95 @@
+import tensorflow as tf
+import os
+import numpy as np
+from tensorflow.keras import layers
+from PIL import Image
+import time
+import matplotlib.pyplot as plt
+from utils import *
+import tensorflow_decision_forests as tfds
+
+batch_size = 128
+epochs = 3
+mnist = tf.keras.datasets.mnist
+
+filedir = os.path.dirname(__file__)
+
+TRAINDIR = filedir + "/../data/chest_xray/train"
+TESTDIR = filedir + "/../data/chest_xray/test"
+BATCHSIZE = 128
+IMG_HEIGHT = 200
+IMG_WIDTH = 200
+
+train_ds = tf.keras.utils.image_dataset_from_directory(
+    TRAINDIR,
+    labels="inferred",
+    seed=1337,
+    image_size=(IMG_HEIGHT, IMG_WIDTH),
+    batch_size=BATCHSIZE,
+    color_mode="grayscale",
+)
+
+val_ds = tf.keras.utils.image_dataset_from_directory(
+    TESTDIR,
+    labels="inferred",
+    seed=1337,
+    image_size=(IMG_HEIGHT, IMG_WIDTH),
+    batch_size=BATCHSIZE,
+    color_mode="grayscale",
+)
+
+AUTOTUNE = tf.data.AUTOTUNE
+
+train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
+val_ds = val_ds.cache().prefetch(buffer_size=AUTOTUNE)
+
+
+model = tf.keras.Sequential(
+    [
+        layers.Rescaling(1.0 / 255),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Flatten(),
+        layers.Dense(128, activation="relu"),
+        layers.Dense(1),
+    ]
+)
+
+
+model.compile(
+    optimizer="adam",
+    loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
+    metrics=["accuracy"],
+)
+
+model.fit(train_ds, validation_data=val_ds, epochs=epochs)
[…]

### 2022-11-25 src/conv_rf_mnist.py
Testing on converting numpy to tf dataset

@@ -14,8 +14,16 @@ mnist = tf.keras.datasets.mnist
 
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
 
+print(x_train.shape)
+
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
+
+print(x_train.shape)
+
+train_ds = tf.data.Dataset.from_tensor_slices((x_train, y_train))
+val_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
+print(train_ds)
 
 model = tf.keras.Sequential(
     [
@@ -39,25 +47,26 @@ model.compile(
     metrics=["accuracy"],
 )
 
-feature_extractor = tf.keras.Model(
-    inputs=model.inputs,
-    outputs=model.layers[-2].output,
-)
 
 model.fit(
-    x_train,
-    y_train,
-    validation_data=(x_test, y_test),
+    train_ds,
+    validation_data=val_ds,
     epochs=epochs,
     batch_size=batch_size,
 )
 
-train_features = feature_extractor(x_train[:10000])
-test_features = feature_extractor(x_test)
+feature_extractor = tf.keras.Model(
+    inputs=model.inputs,
+    outputs=model.layers[-3].output,
+)
+
+features_train = train_ds.map(lambda batch, label: (feature_extractor(batch), label))
+features_test = val_ds.map(lambda batch, label: (feature_extractor(batch), label))
 
 forest = RandomForestClassifier(random_state=1337)
 forest.compile(metrics=["accuracy"])
 
-forest.fit(train_features, y_train[:10000])
+forest.fit(features_train, y_train[:10000])
 
+print(forest.score(test_features, y_test))
 print(forest.score(test_features, y_test))

### 2022-11-25 src/conv_rf_pneu.py

@@ -42,6 +42,7 @@ AUTOTUNE = tf.data.AUTOTUNE
 
 train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
 val_ds = val_ds.cache().prefetch(buffer_size=AUTOTUNE)
+print(train_ds)
 
 
 model = tf.keras.Sequential(

### 2022-11-25 src/conv_rf_mnist.py
Convert to tf dataset succesful

@@ -6,7 +6,7 @@ from PIL import Image
 import time
 import matplotlib.pyplot as plt
 from utils import *
-from sklearn.ensemble import RandomForestClassifier
+import tensorflow_decision_forests as tfds
 
 batch_size = 128
 epochs = 1
@@ -19,11 +19,17 @@ print(x_train.shape)
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
 
+# y_train = tf.reshape(y_train, shape=[-1, 1])
+# y_test = tf.reshape(y_test, shape=[-1, 1])
+
 print(x_train.shape)
 
 train_ds = tf.data.Dataset.from_tensor_slices((x_train, y_train))
 val_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
 print(train_ds)
+
+train_ds = train_ds.batch(batch_size)
+val_ds = val_ds.batch(batch_size)
 
 model = tf.keras.Sequential(
     [
@@ -52,7 +58,7 @@ model.fit(
     train_ds,
     validation_data=val_ds,
     epochs=epochs,
-    batch_size=batch_size,
+    # batch_size=batch_size,
 )
 
 feature_extractor = tf.keras.Model(
@@ -63,10 +69,12 @@ feature_extractor = tf.keras.Model(
 features_train = train_ds.map(lambda batch, label: (feature_extractor(batch), label))
 features_test = val_ds.map(lambda batch, label: (feature_extractor(batch), label))
 
-forest = RandomForestClassifier(random_state=1337)
+forest = tfds.keras.RandomForestModel(
+    verbose=1, max_depth=16, random_seed=1337, check_dataset=False
+)
+forest.fit(x=features_train)
+
 forest.compile(metrics=["accuracy"])
 
-forest.fit(features_train, y_train[:10000])
-
-print(forest.score(test_features, y_test))
-print(forest.score(test_features, y_test))
+print(forest.evaluate(features_train, return_dict=True))
+print(forest.evaluate(features_test, return_dict=True))

### 2022-11-25 src/conv_rf_pneu.py

@@ -92,5 +92,5 @@ forest.fit(x=features_train)
 
 forest.compile(metrics=["accuracy"])
 
-# print(forest.evaluate(features_train, return_dict=True))
+print(forest.evaluate(features_train, return_dict=True))
 print(forest.evaluate(features_test, return_dict=True))

### 2022-11-25 src/conv_nn_mnist.py
Update datasets

@@ -16,6 +16,15 @@ mnist = tf.keras.datasets.mnist
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
 
+print(x_train.shape)
+
+train_ds = tf.data.Dataset.from_tensor_slices((x_train, y_train))
+val_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
+print(train_ds)
+
+train_ds = train_ds.batch(batch_size)
+val_ds = val_ds.batch(batch_size)
+
 model = tf.keras.Sequential(
     [
         tf.keras.layers.Rescaling(1.0 / 255),
@@ -38,11 +47,9 @@ model.compile(
 )
 
 model.fit(
-    x_train,
-    y_train,
-    validation_data=(x_test, y_test),
+    train_ds,
+    validation_data=val_ds,
     epochs=epochs,
-    batch_size=batch_size,
 )
 
 # ------- plotting pca ------------

### 2022-11-25 src/conv_nn_mnist.py
Eremove uneccesary coed

@@ -27,16 +27,16 @@ val_ds = val_ds.batch(batch_size)
 
 model = tf.keras.Sequential(
     [
-        tf.keras.layers.Rescaling(1.0 / 255),
-        tf.keras.layers.Conv2D(32, 3, activation="relu"),
-        tf.keras.layers.MaxPooling2D(),
-        tf.keras.layers.Conv2D(32, 3, activation="relu"),
-        tf.keras.layers.MaxPooling2D(),
-        tf.keras.layers.Conv2D(32, 3, activation="relu"),
-        tf.keras.layers.MaxPooling2D(),
-        tf.keras.layers.Flatten(),
-        tf.keras.layers.Dense(128, activation="relu"),
-        tf.keras.layers.Dense(10),
+        layers.Rescaling(1.0 / 255),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Conv2D(32, 3, activation="relu"),
+        layers.MaxPooling2D(),
+        layers.Flatten(),
+        layers.Dense(128, activation="relu"),
+        layers.Dense(10),
     ]
 )
 

### 2022-11-25 src/pca_nn_mnist.py
Remove old code

@@ -19,6 +19,8 @@ x_test_flat = tf.reshape(x_test, shape=[-1, 784])
 
 W = PCA_fit(x_train_flat, n_components)
 
+# train_ds.map(lambda batch, label: (tf.matmul(batch, W), label))
+
 model = tf.keras.Sequential(
     [
         PCALayer(W),
@@ -42,59 +44,3 @@ model.fit(
     epochs=6,
     batch_size=64,
 )
-
-# ------- plotting pca ------------
-# x_train_remade = pca.inverse_transform(x_train_pca)
-#
-# plt.subplot(421)
-# plt.title("Actual instance, \n 784 features")
-# plt.imshow(x_train[0])
-# plt.subplot(422)
-# plt.title(f"Constructed using only the \n {n_components} first principal components")
-# plt.imshow(tf.reshape(x_train_remade[0], (28, 28)))
-#
-# plt.subplot(423)
-# plt.imshow(x_train[1])
-# plt.subplot(424)
-# plt.imshow(tf.reshape(x_train_remade[1], (28, 28)))
-#
-# plt.subplot(425)
-# plt.imshow(x_train[2])
-# plt.subplot(426)
-# plt.imshow(tf.reshape(x_train_remade[2], (28, 28)))
-#
-# plt.subplot(427)
-# plt.imshow(x_train[3])
-# plt.subplot(428)
-# plt.imshow(tf.reshape(x_train_remade[3], (28, 28)))
-#
-# plt.show()
-
-# class PCALayer(layers.Layer):
-#     def __init__(self, num_outputs):
-#         super(PCALayer, self).__init__()
-#         self.num_outputs = num_outputs
-#         self.trainable = False
-#
-#     def call(self, inputs):
-#         batch_size = inputs.shape[0] or 1
-#         features = inputs.shape[1] * inputs.shape[2]
-#
-#         x_list = []
-#         for i in range(batch_size):
-#             x_list.append(tf.reshape(inputs[i], (features,)))
-#
-#         X = tf.stack(x_list)
-#
-#         means = tf.reduce_mean(X)
-#         print(means)
-#
-#         S, U, V = tf.linalg.svd(X)
-#
-#         slice_index = min(self.num_outputs, batch_size)
-#         S = S[:slice_index]
-#         S = tf.linalg.diag(S)
-#         U = U[:, :slice_index]
-#         output = tf.linalg.matmul(U, S)
-#
-#         return output

### 2022-11-25 doc/paper.tex
Start on stochastic pca, more text

@@
 Convolutional Neural Networks are neural networks which are specialized to work on images, or other types of data where the structure of the data carries meaning.
-A normal neural network treats each input indepentently, and inputs are flattened such that any spatial information is lost.
+A normal neural network treats each input independently, and inputs are flattened such that any spatial information is lost.
 CNNs, on the other hand, explicitly take the spatial dimensions into account, by using convolutional kernels and pooling layers as part of their architecture.
@@
 A decision tree is an intiutive way of performing both classification and regression tasks, by asking a series of yes or no questions about an instance in the dataset.
-These question will split the original dataset, the root node, into child nodes which contain subsets of the dataset, and these nodes will be split again, creating a tree structure which terminates in leaf nodes which imply a choice of class.
+These questions will split the original dataset, the root node, into child nodes which contain subsets of the dataset, and these nodes will be split again, creating a tree structure which terminates in leaf nodes which imply a choice of class.
 
-In a binary classification problem, the algorithm for creating a decision tree is quite simple.
+For classification problems, the algorithm for creating a decision tree is quite simple.
 First, consider each feature of the dataset, and attempt to split all instances into two child nodes based on a simple true/false statement about this feature.
@@
 We now create one decision tree per bootstrap, only using a random subset of the total amount of features every time we perform a split\footnote{The number of features used is often set to $\sqrt{p}$}.
-The result of this is that we create many slightly different trees, each slightly suboptimal in their splits.
-These trees give predictions which are worse than simply creating a single decision tree using all the features and all the data, but when basing our prediction of the majority vote of all the trees, we are able to make predictions which are more generalizable and less prone to overfitting.
+The result of this is that we create many slightly different trees.
+These trees give predictions which are worse than simply creating a single decision tree using all the features and all the data, but when basing our prediction on the majority vote of all the trees, we are able to make predictions which are more generalizable and less prone to overfitting.
 
@@
 
-The MNIST dataset is well known multiclass classification problem used to test machine learning algorithms.
+The MNIST dataset is a well known multiclass classification problem used to test machine learning algorithms.
 It consists of $70000$ black and white images of hand drawn digits, each $28$ by $28$ pixels.
-Each image is labeled with one of ten classes, corresponding to the digit it depicting.
-The dataset is divided into $60000$ training images and $10000$ validation images.This dataset is essentially uniformly distributed, with all classes containing around $9 - 11\%$ of the total dataset.
+Each image is labeled with one of ten classes, corresponding to the digit it depicts.
+The dataset is divided into $60000$ training images and $10000$ validation images.
+This dataset is essentially uniformly distributed, with all classes containing around $9 - 11\%$ of the total dataset.
 
@@
 The images were taken from the Woman and Children's Medical Center in Guangzhou, China.
-The dataset is divided into a training set of $5232$ images and $624$.
+The dataset is divided into a training set of $5232$ images and $624$ validation images.
 The dataset is not evenly distributed, having $74.2\%$ instances of pneumonia in the training set and $62.5\%$ instances of pneumonia in the validation set.

### 2022-11-25 src/utils.py

@@ -1,5 +1,37 @@
 import tensorflow as tf
 from tensorflow.keras import layers
+import numpy as np
+
+
+def PCA_stoch(X, n_components):
+    print("eric")
+
+
+def PCA_stoch(X, n_components, iterations=10, eta=1e-5, convergence=1e-5, epochs=10):
+    n, d = X.shape
+    W_t = np.random.rand(d, n_components) - 0.5
+    print(W_t.shape)
+    W_t = W_t / np.linalg.norm(W_t, axis=0)
+    X = X - np.mean(X, axis=0)
+
+    for e in range(epochs):
+
+        W = W_t
+
+        for t in range(iterations):
+            i = np.random.randint(n)
+            print(f"{W.shape=}")
+            print(f"{X[i].shape=}")
+            print(f"{X[i].T.dot(W).shape=}")
+            W_mark = W + eta * (X[i] * (X[i].T.dot(W) - X[i].T.dot(W_t)))
+            W_mark = W_mark / np.linalg.norm(W_mark, axis=0)
+            W = _W
+
+        d = np.linalg.norm(W_t - W)
+        W_t = W
+
+        if d < rate:
+            return
 
 
 def PCA_fit(X, n_components):

### 2022-11-28 doc/paper.tex
Add more theory on convolution

@@
 In a convolutional neural network, the values of each kernel is initialized randomly, and training involves tuning these values.
-When performing convolution at the edges of an image, some values will be undefined.
+When performing convolution at the edges of an image, some values in the image will be undefined (i.e, some values of the kernel will not overlap with the image).
+To solve this problem, we can either ignore any positions where the kernel is not fully contained within the image, or introduce padding, where we pad the image with values (usually just zeroes) so that more of the image can be convolved.
+By choosing the amount of padding, we can decide how close the dimensions of the result of our convolution are to the original dimensions.
+Another parameter which has an effect on the output dimensions is the {\it stride} parameter, which decides the size of the step between each convolution operation.
+By having a stride higher than one, we are no longer evaluating at every position.
+This reduces the computational complexity at the cost of not extracting our features as finely \cite[343]{gbc}.
+This will also downsample the result of our convolution to a smaller dimension than the input.
+
+The convolution stage is often followed by an activation function, like the RELU function.
+After this, it is common to use a {\it pooling layer}.
+This layer replaces the output at a certain location with a summary statistic of the nearby outputs \cite[335]{gbc}.
+We could for example replace the value at a given position with the maximum value in a neighbourhood around this position, which is known as {\it max pooling}.
+By using pooling, we are making our output invariant to small translations of the input.
+This is useful if we care more about whether a feature is present rather than exactly where it is \cite[336]{gbc}.
+Furthermore, by using a stride parameter here as well, we are able to downsample our output.
+
+After several layers of convolution, the input has been usually been downsampled, and important features about the input have hopefully been extracted by use of convolution and pooling.
+The output is then flattened, and the features are sent to one or more fully connected layers.
+The outputs of these layers are then used to make a prediction.
 

### 2022-11-28 src/conv_nn_mnist.py
Remove old code

@@ -51,59 +51,3 @@ model.fit(
     validation_data=val_ds,
     epochs=epochs,
 )
-
-# ------- plotting pca ------------
-# x_train_remade = pca.inverse_transform(x_train_pca)
-#
-# plt.subplot(421)
-# plt.title("Actual instance, \n 784 features")
-# plt.imshow(x_train[0])
-# plt.subplot(422)
-# plt.title(f"Constructed using only the \n {n_components} first principal components")
-# plt.imshow(tf.reshape(x_train_remade[0], (28, 28)))
-#
-# plt.subplot(423)
-# plt.imshow(x_train[1])
-# plt.subplot(424)
-# plt.imshow(tf.reshape(x_train_remade[1], (28, 28)))
-#
-# plt.subplot(425)
-# plt.imshow(x_train[2])
-# plt.subplot(426)
-# plt.imshow(tf.reshape(x_train_remade[2], (28, 28)))
-#
-# plt.subplot(427)
-# plt.imshow(x_train[3])
-# plt.subplot(428)
-# plt.imshow(tf.reshape(x_train_remade[3], (28, 28)))
-#
-# plt.show()
-
-# class PCALayer(layers.Layer):
-#     def __init__(self, num_outputs):
-#         super(PCALayer, self).__init__()
-#         self.num_outputs = num_outputs
-#         self.trainable = False
-#
-#     def call(self, inputs):
-#         batch_size = inputs.shape[0] or 1
-#         features = inputs.shape[1] * inputs.shape[2]
-#
-#         x_list = []
-#         for i in range(batch_size):
-#             x_list.append(tf.reshape(inputs[i], (features,)))
-#
-#         X = tf.stack(x_list)
-#
-#         means = tf.reduce_mean(X)
-#         print(means)
-#
-#         S, U, V = tf.linalg.svd(X)
-#
-#         slice_index = min(self.num_outputs, batch_size)
-#         S = S[:slice_index]
-#         S = tf.linalg.diag(S)
-#         U = U[:, :slice_index]
-#         output = tf.linalg.matmul(U, S)
-#
-#         return output

### 2022-11-28 src/utils.py
Some stuff with VR-PCA that wont be used

@@ -8,30 +8,31 @@ def PCA_stoch(X, n_components):
 
 
 def PCA_stoch(X, n_components, iterations=10, eta=1e-5, convergence=1e-5, epochs=10):
-    n, d = X.shape
-    W_t = np.random.rand(d, n_components) - 0.5
-    print(W_t.shape)
-    W_t = W_t / np.linalg.norm(W_t, axis=0)
-    X = X - np.mean(X, axis=0)
+    print(f"{X.shape=}")
+    m, n = X.shape
+    W_prev_epoch = np.eye(m, n_components)
 
     for e in range(epochs):
+        print(e)
 
-        W = W_t
+        W_prev_iter = W_prev_epoch
+        u_mark = np.zeros((m, n_components))
+        for i in range(n):
+            u_mark += np.outer(X[:, i], (X[:, i].T @ W_prev_epoch)) / n
 
         for t in range(iterations):
+            print(t)
             i = np.random.randint(n)
-            print(f"{W.shape=}")
-            print(f"{X[i].shape=}")
-            print(f"{X[i].T.dot(W).shape=}")
-            W_mark = W + eta * (X[i] * (X[i].T.dot(W) - X[i].T.dot(W_t)))
-            W_mark = W_mark / np.linalg.norm(W_mark, axis=0)
-            W = _W
 
-        d = np.linalg.norm(W_t - W)
-        W_t = W
+            X_i = X[:, i].reshape(X[:, i].shape[0], 1)
+            W_mark = W_prev_iter + eta * (
+                X_i * (X_i.T @ W_prev_iter - X_i.T @ W_prev_epoch) + u_mark
+            )
+            W_prev_iter, _ = np.linalg.qr(W_mark)
 
-        if d < rate:
-            return
+        W_prev_epoch = W_prev_iter
+
+    return W_prev_epoch
 
 
 def PCA_fit(X, n_components):

### 2022-11-28 src/pca_nn_mnist.py
Add stochastic pca

@@ -6,6 +6,7 @@ from PIL import Image
 import time
 import matplotlib.pyplot as plt
 from utils import *
+from sklearn.decomposition import PCA
 
 mnist = tf.keras.datasets.mnist
 
@@ -17,13 +18,18 @@ n_components = 100
 x_train_flat = tf.reshape(x_train, shape=[-1, 784])
 x_test_flat = tf.reshape(x_test, shape=[-1, 784])
 
-W = PCA_fit(x_train_flat, n_components)
+pca = PCA(n_components=n_components, svd_solver="randomized")
 
-# train_ds.map(lambda batch, label: (tf.matmul(batch, W), label))
+print("Fitting PCA")
+start = time.time()
+pca.fit(x_train_flat)
+
+pca.transform(x_train_flat)
+pca.transform(x_test_flat)
+
 
 model = tf.keras.Sequential(
     [
-        PCALayer(W),
         layers.Dense(n_components, activation="relu"),
         layers.Dense(n_components, activation="relu"),
         layers.Dense(n_components, activation="relu"),

### 2022-11-28 src/pca_nn_mnist.py
Add hyperparameter tuning

@@ -7,6 +7,7 @@ import time
 import matplotlib.pyplot as plt
 from utils import *
 from sklearn.decomposition import PCA
+import keras_tuner as kt
 
 mnist = tf.keras.datasets.mnist
 
@@ -14,6 +15,37 @@ mnist = tf.keras.datasets.mnist
 x_train, x_test = x_train / 255.0, x_test / 255.0
 
 n_components = 100
+
+
+def model_builder(hp):
+    hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
+    model = tf.keras.Sequential(
+        [
+            layers.Dense(units=hp_units, activation="relu"),
+            layers.Dense(units=hp_units, activation="relu"),
+            layers.Dense(units=hp_units, activation="relu"),
+            layers.Dense(10),
+        ]
+    )
+
+    hp_learning_rate = hp.Choice("learning_rate", values=[1e-2, 1e-3, 1e-4])
+
+    model.compile(
+        optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
+        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+        metrics=["accuracy"],
+    )
+
+    return model
+
+
+tuner = kt.Hyperband(
+    model_builder,
+    objective="val_accuracy",
+    max_epochs=10,
+    factor=3,
+)
+
 
 x_train_flat = tf.reshape(x_train, shape=[-1, 784])
 x_test_flat = tf.reshape(x_test, shape=[-1, 784])
@@ -27,21 +59,11 @@ pca.fit(x_train_flat)
 pca.transform(x_train_flat)
 pca.transform(x_test_flat)
 
+tuner.search(x_train_flat, y_train, epochs=10, validation_split=0.2)
+best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
 
-model = tf.keras.Sequential(
-    [
-        layers.Dense(n_components, activation="relu"),
-        layers.Dense(n_components, activation="relu"),
-        layers.Dense(n_components, activation="relu"),
-        layers.Dense(10),
-    ]
-)
-
-model.compile(
-    optimizer="adam",
-    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
-    metrics=["accuracy"],
-)
+print(best_hps.get("units"))
+print(best_hps.get("learning_rate"))
 
 model.fit(
     x_train_flat,

### 2022-11-28 src/conv_nn_pneu.py
change PCA

@@ -1,8 +1,11 @@
 import tensorflow as tf
 import os
 from tensorflow.keras import layers
+from tensorflow.keras import regularizers
 
 filedir = os.path.dirname(__file__)
+
+tf.keras.utils.set_random_seed(1336)
 
 TRAINDIR = filedir + "/../data/chest_xray/train"
 TESTDIR = filedir + "/../data/chest_xray/test"
@@ -33,18 +36,20 @@ AUTOTUNE = tf.data.AUTOTUNE
 train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
 test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
 
+kernel = tf.keras.regularizers.L2(l2=0.001)
+bias = tf.keras.regularizers.L2(l2=0.001)
 
 model = tf.keras.Sequential(
     [
         tf.keras.layers.Rescaling(1.0 / 255),
-        tf.keras.layers.Conv2D(32, 3, activation="relu"),
+        tf.keras.layers.Conv2D(32, 7, activation="relu", kernel_regularizer=kernel, bias_regularizer=bias),
         tf.keras.layers.MaxPooling2D(),
-        tf.keras.layers.Conv2D(32, 3, activation="relu"),
+        tf.keras.layers.Conv2D(32, 7, activation="relu", kernel_regularizer=kernel, bias_regularizer=bias),
         tf.keras.layers.MaxPooling2D(),
-        tf.keras.layers.Conv2D(32, 3, activation="relu"),
+        tf.keras.layers.Conv2D(32, 7, activation="relu", kernel_regularizer=kernel, bias_regularizer=bias),
         tf.keras.layers.MaxPooling2D(),
         tf.keras.layers.Flatten(),
-        tf.keras.layers.Dense(128, activation="relu"),
+        tf.keras.layers.Dense(128, activation="relu", kernel_regularizer=kernel, bias_regularizer=bias),
         tf.keras.layers.Dense(1),
     ]
 )

### 2022-11-28 src/pca_nn_mnist.py

@@ -8,17 +8,19 @@ import matplotlib.pyplot as plt
 from utils import *
 from sklearn.decomposition import PCA
 
+tf.keras.utils.set_random_seed(1336)
 mnist = tf.keras.datasets.mnist
 
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
 x_train, x_test = x_train / 255.0, x_test / 255.0
 
-n_components = 100
+n_components = 20
 
 x_train_flat = tf.reshape(x_train, shape=[-1, 784])
 x_test_flat = tf.reshape(x_test, shape=[-1, 784])
 
-pca = PCA(n_components=n_components, svd_solver="randomized")
+pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
+# pca = PCA(n_components=n_components, svd_solver="full", random_state=1336)
 
 print("Fitting PCA")
 start = time.time()
@@ -50,3 +52,4 @@ model.fit(
     epochs=6,
     batch_size=64,
 )
+print(f"Time taken: {time.time() - start}")

### 2022-11-28 src/pca_nn_pneu.py

@@ -2,12 +2,16 @@ import tensorflow as tf
 import os
 import numpy as np
 from tensorflow.keras import layers
+from tensorflow.keras import regularizers
 from PIL import Image
 import time
 import matplotlib.pyplot as plt
 from utils import *
+from sklearn.decomposition import PCA
 
 filedir = os.path.dirname(__file__)
+
+tf.keras.utils.set_random_seed(1336)
 
 TRAINDIR = filedir + "/../data/chest_xray/train"
 TESTDIR = filedir + "/../data/chest_xray/test"
@@ -37,29 +41,52 @@ val_ds = tf.keras.utils.image_dataset_from_directory(
 # grab a subset for PCA calculation
 x_list = []
 i = 0
-for batch, _ in train_ds:
-    x_list.append(tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0)
-    i += 1
-    if i > 20:
-        break
+X_train = None
+X_test = None
+y_train = None
+y_test = None
 
-X_subset = tf.concat(x_list, axis=0)
-n_components = 1000
-
-W = PCA_fit(X_subset, n_components)
-
-
+n_components = 5000
+kernel = tf.keras.regularizers.L2(l2=0.001)
+bias = tf.keras.regularizers.L2(l2=0.001)
 model = tf.keras.Sequential(
     [
-        layers.Rescaling(1.0 / 255),
-        layers.Flatten(),
-        PCALayer(W),
-        layers.Dense(n_components, activation="relu"),
-        layers.Dense(n_components, activation="relu"),
-        layers.Dense(n_components, activation="relu"),
+        layers.Dense(n_components, kernel_regularizer=kernel, bias_regularizer=bias, activation="relu"),
+        layers.Dense(n_components, kernel_regularizer=kernel, bias_regularizer=bias, activation="relu"),
+        layers.Dense(n_components, kernel_regularizer=kernel, bias_regularizer=bias, activation="relu"),
         layers.Dense(1),
     ]
 )
+
+for batch, labels in train_ds:
+    if X_train is None:
+        X_train = tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0
+        y_train = labels
+
+    else:
+        X_train = tf.concat([X_train, (tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0)], axis=0)
+        y_train = tf.concat([y_train, labels], axis=0)
+
[…]

### 2022-11-28 src/conv_nn_mnist.py
Add hyperparameter tuning and crossvalidation

@@ -6,6 +6,8 @@ from PIL import Image
 import time
 import matplotlib.pyplot as plt
 from utils import *
+from sklearn.model_selection import KFold
+import keras_tuner as kt
 
 batch_size = 128
 epochs = 6
@@ -16,38 +18,105 @@ mnist = tf.keras.datasets.mnist
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
 
-print(x_train.shape)
+# print(x_train.shape)
+#
+# train_ds = tf.data.Dataset.from_tensor_slices((x_train, y_train))
+# val_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
+# print(train_ds)
+#
+# train_ds = train_ds.batch(batch_size)
+# val_ds = val_ds.batch(batch_size)
 
-train_ds = tf.data.Dataset.from_tensor_slices((x_train, y_train))
-val_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
-print(train_ds)
 
-train_ds = train_ds.batch(batch_size)
-val_ds = val_ds.batch(batch_size)
+def model_builder(hp):
+    hp_size = hp.Int("units", min_value=3, max_value=7, step=2)
 
-model = tf.keras.Sequential(
-    [
-        layers.Rescaling(1.0 / 255),
-        layers.Conv2D(32, 3, activation="relu"),
-        layers.MaxPooling2D(),
-        layers.Conv2D(32, 3, activation="relu"),
-        layers.MaxPooling2D(),
-        layers.Conv2D(32, 3, activation="relu"),
-        layers.MaxPooling2D(),
-        layers.Flatten(),
-        layers.Dense(128, activation="relu"),
-        layers.Dense(10),
-    ]
+    model = tf.keras.Sequential(
+        [
+            layers.Rescaling(1.0 / 255),
+            layers.Conv2D(32, hp_size, activation="relu", padding="same"),
+            layers.MaxPooling2D(),
+            layers.Conv2D(32, hp_size, activation="relu", padding="same"),
+            layers.MaxPooling2D(),
+            layers.Conv2D(32, hp_size, activation="relu", padding="same"),
+            layers.MaxPooling2D(),
+            layers.Flatten(),
+            layers.Dense(128, activation="relu"),
+            layers.Dense(10),
+        ]
+    )
+
[…]

### 2022-11-28 src/pca_nn_mnist.py

@@ -17,37 +17,6 @@ x_train, x_test = x_train / 255.0, x_test / 255.0
 
 n_components = 20
 
-
-def model_builder(hp):
-    hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
-    model = tf.keras.Sequential(
-        [
-            layers.Dense(units=hp_units, activation="relu"),
-            layers.Dense(units=hp_units, activation="relu"),
-            layers.Dense(units=hp_units, activation="relu"),
-            layers.Dense(10),
-        ]
-    )
-
-    hp_learning_rate = hp.Choice("learning_rate", values=[1e-2, 1e-3, 1e-4])
-
-    model.compile(
-        optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
-        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
-        metrics=["accuracy"],
-    )
-
-    return model
-
-
-tuner = kt.Hyperband(
-    model_builder,
-    objective="val_accuracy",
-    max_epochs=10,
-    factor=3,
-)
-
-
 x_train_flat = tf.reshape(x_train, shape=[-1, 784])
 x_test_flat = tf.reshape(x_test, shape=[-1, 784])
 
@@ -61,11 +30,54 @@ pca.fit(x_train_flat)
 pca.transform(x_train_flat)
 pca.transform(x_test_flat)
 
-tuner.search(x_train_flat, y_train, epochs=10, validation_split=0.2)
-best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
+# def model_builder(hp):
+#     hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
+#     model = tf.keras.Sequential(
+#         [
+#             layers.Dense(units=hp_units, activation="relu"),
+#             layers.Dense(units=hp_units, activation="relu"),
+#             layers.Dense(units=hp_units, activation="relu"),
+#             layers.Dense(10),
+#         ]
+#     )
+#
+#     hp_learning_rate = hp.Choice("learning_rate", values=[1e-2, 1e-3, 1e-4])
+#
+#     model.compile(
+#         optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
+#         loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+#         metrics=["accuracy"],
+#     )
+#
+#     return model
+#
+#
[…]

### 2022-11-29 doc/paper.tex
More paper
The most important commit yet

@@
 
-\author{Eric E. Reber
+% \author{Eric E. Reber
+%     \affiliation{
+% 	Bachelor Student, Robotics and \\
+% 	Intelligent Systems\\ \\[-10pt]
+% 	Department of Informatics\\ \\[-10pt]
+% 	The faculty of Mathematics and \\
+% 	Natural Sciences\\ \\[-10pt]
+%     Email: ericer@ifi.uio.no
+%     }
+% }
+
+\author{Ed "Gurgler" Bickley
     \affiliation{
-	Bachelor Student, Robotics and \\
-	Intelligent Systems\\ \\[-10pt]
-	Department of Informatics\\ \\[-10pt]
-	The faculty of Mathematics and \\
-	Natural Sciences\\ \\[-10pt]
-    Email: ericer@ifi.uio.no
+	Inmate, Texas State \\
+	Penistentiary\\ \\[-10pt]
+	Huntsville, Texas\\ \\[-10pt]
+	United States of America\\
+    Email: thegurgler@hotmail.com
     }
@@
 
+For very large design matrices, the computational complexity of calculating the singular value decomposition can be significant.
+However, there exists stochastic methods to achieve good approximations, like Randomized PCA \cite[227]{halko}.
+
 % Principal Component Analysis is a dimensionality reduction method performs a change of basis into a lower dimensional subspace of our feature space, while capturing as much of the variance of our dataset as possible.
@@
 
-\subsection{Traditional Convolutional Neural Network}
+\subsection{MNIST}
 
-\subsection{Convolution followed by a Random Forest}
+\subsubsection{Traditional Convolutional Neural Network}
 
-\subsection{Principal Component Analysis followed by a Neural Network}
+\subsubsection{Convolution followed by a Random Forest}
 
-\subsection{Principal Component Analysis followed by a Random Forest}
+\subsubsection{Principal Component Analysis followed by a Neural Network}
+
+\subsubsection{Principal Component Analysis followed by a Random Forest}
+
+\subsection{Guangzhou Pneumonia Data}
+
+Hopefully this is real
+
+\subsubsection{Traditional Convolutional Neural Network}
+
+\subsubsection{Convolution followed by a Random Forest}
+
+\subsubsection{Principal Component Analysis followed by a Neural Network}
+
+\subsubsection{Principal Component Analysis followed by a Random Forest}
 

### 2022-11-29 src/conv_nn_mnist.py
Add hyperparameter tuning

@@ -29,16 +29,16 @@ x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
 
 
 def model_builder(hp):
-    hp_size = hp.Int("units", min_value=3, max_value=7, step=2)
+    hp_width = hp.Choice("width", values=[3, 5, 7])
 
     model = tf.keras.Sequential(
         [
             layers.Rescaling(1.0 / 255),
-            layers.Conv2D(32, hp_size, activation="relu", padding="same"),
+            layers.Conv2D(32, hp_width, activation="relu", padding="same"),
             layers.MaxPooling2D(),
-            layers.Conv2D(32, hp_size, activation="relu", padding="same"),
+            layers.Conv2D(32, hp_width, activation="relu", padding="same"),
             layers.MaxPooling2D(),
-            layers.Conv2D(32, hp_size, activation="relu", padding="same"),
+            layers.Conv2D(32, hp_width, activation="relu", padding="same"),
             layers.MaxPooling2D(),
             layers.Flatten(),
             layers.Dense(128, activation="relu"),
@@ -65,12 +65,10 @@ tuner = kt.Hyperband(
     directory="conv_nn_mnist-params",
 )
 
-print("searching")
 tuner.search(x_train, y_train, epochs=10, validation_split=0.2)
-print("done")
 best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
-print(best_hps.get("units"))
-print(best_hps.get("learning_rate"))
+width = best_hps.get("width")
+learning_rate = best_hps.get("learning_rate")
 
 kf = KFold(n_splits=5)  # random_state=1336)
 
@@ -86,11 +84,11 @@ for train_index, test_index in kf.split(x_train, y_train):
     model = tf.keras.Sequential(
         [
             layers.Rescaling(1.0 / 255),
-            layers.Conv2D(32, 3, activation="relu"),
+            layers.Conv2D(32, width, activation="relu"),
             layers.MaxPooling2D(),
-            layers.Conv2D(32, 3, activation="relu"),
+            layers.Conv2D(32, width, activation="relu"),
             layers.MaxPooling2D(),
-            layers.Conv2D(32, 3, activation="relu"),
+            layers.Conv2D(32, width, activation="relu"),
             layers.MaxPooling2D(),
             layers.Flatten(),
             layers.Dense(128, activation="relu"),
[…]

### 2022-11-29 src/pca_nn_mnist.py
Add crossval to pca_nn_mnist

@@ -27,63 +27,99 @@ print("Fitting PCA")
 start = time.time()
 pca.fit(x_train_flat)
 
-pca.transform(x_train_flat)
-pca.transform(x_test_flat)
+x_train_pca = pca.transform(x_train_flat)
+x_train_pca = pca.transform(x_test_flat)
 
-# def model_builder(hp):
-#     hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
-#     model = tf.keras.Sequential(
-#         [
-#             layers.Dense(units=hp_units, activation="relu"),
-#             layers.Dense(units=hp_units, activation="relu"),
-#             layers.Dense(units=hp_units, activation="relu"),
-#             layers.Dense(10),
-#         ]
-#     )
-#
-#     hp_learning_rate = hp.Choice("learning_rate", values=[1e-2, 1e-3, 1e-4])
-#
-#     model.compile(
-#         optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
-#         loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
-#         metrics=["accuracy"],
-#     )
-#
-#     return model
-#
-#
-# tuner = kt.Hyperband(
-#     model_builder,
-#     objective="val_accuracy",
-#     max_epochs=10,
-#     factor=3,
-# )
-#
-# tuner.search(x_train_flat, y_train, epochs=10, validation_split=0.2)
-# best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
-# print(best_hps.get("units"))
-# print(best_hps.get("learning_rate"))
 
-model = tf.keras.Sequential(
-    [
-        layers.Dense(224, activation="relu"),
-        layers.Dense(224, activation="relu"),
-        layers.Dense(224, activation="relu"),
-        layers.Dense(10),
-    ]
+def model_builder(hp):
+    hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
+    model = tf.keras.Sequential(
+        [
+            layers.Dense(units=hp_units, activation="relu"),
+            layers.Dense(units=hp_units, activation="relu"),
+            layers.Dense(units=hp_units, activation="relu"),
+            layers.Dense(10),
+        ]
+    )
+
+    hp_learning_rate = hp.Choice("learning_rate", values=[1e-2, 1e-3, 1e-4])
+
+    model.compile(
[…]

### 2022-11-29 src/test_pca.py
Add test pca script

@@ -1,0 +1,69 @@
+import tensorflow as tf
+import os
+import numpy as np
+from tensorflow.keras import layers
+from PIL import Image
+import time
+import matplotlib.pyplot as plt
+from utils import *
+from sklearn.decomposition import PCA
+
+from tensorflow.python.ops.numpy_ops import np_config
+
+np_config.enable_numpy_behavior()
+
+mnist = tf.keras.datasets.mnist
+
+(x_train, y_train), (x_test, y_test) = mnist.load_data()
+x_train, x_test = x_train / 255.0, x_test / 255.0
+
+x_train_flat = x_train.reshape((x_train.shape[0], 784))
+x_test_flat = x_test.reshape((x_test.shape[0], 784))
+x_train_flat = x_train_flat
+
+
+pca35 = PCA(n_components=35, svd_solver="randomized")
+pca50 = PCA(n_components=50, svd_solver="randomized")
+pca100 = PCA(n_components=100, svd_solver="randomized")
+
+pca35.fit(x_train_flat)
+temp = pca35.transform(x_train_flat)
+x35 = pca35.inverse_transform(temp)
+
+pca50.fit(x_train_flat)
+temp = pca50.transform(x_train_flat)
+x50 = pca50.inverse_transform(temp)
+
+pca100.fit(x_train_flat)
+temp = pca100.transform(x_train_flat)
+x100 = pca100.inverse_transform(temp)
+
+
+# ------- plotting pca ------------
+plt.subplot(431)
+plt.title("Actual instance, \n 784 features")
+plt.imshow(tf.reshape(x_train[0], (28, 28)))
+plt.subplot(432)
+plt.title(f"Constructed using only the \n {n_components} first principal components")
+plt.imshow(tf.reshape(x35[0], (28, 28)))
+plt.subplot(433)
+plt.title(
+    f"Constructed using only the \n {n_components} first principal components but its stochastic"
+)
+plt.imshow(tf.reshape(x50[0], (28, 28)))
+
+plt.subplot(434)
+plt.imshow(tf.reshape(x_train[1], (28, 28)))
+plt.subplot(435)
+plt.imshow(tf.reshape(x35[1], (28, 28)))
+plt.subplot(436)
+plt.imshow(tf.reshape(x50[1], (28, 28)))
+
+plt.subplot(437)
+plt.imshow(tf.reshape(x_train[2], (28, 28)))
[…]

### 2022-11-29 src/test_pca.py
Add more plotting

@@ -40,30 +40,35 @@ x100 = pca100.inverse_transform(temp)
 
 
 # ------- plotting pca ------------
-plt.subplot(431)
+plt.subplot(341)
 plt.title("Actual instance, \n 784 features")
 plt.imshow(tf.reshape(x_train[0], (28, 28)))
-plt.subplot(432)
-plt.title(f"Constructed using only the \n {n_components} first principal components")
-plt.imshow(tf.reshape(x35[0], (28, 28)))
-plt.subplot(433)
-plt.title(
-    f"Constructed using only the \n {n_components} first principal components but its stochastic"
-)
+plt.subplot(342)
+plt.title("PCA with 100 components")
+plt.imshow(tf.reshape(x100[0], (28, 28)))
+plt.subplot(343)
+plt.title("PCA with 50 components")
 plt.imshow(tf.reshape(x50[0], (28, 28)))
+plt.subplot(344)
+plt.title("PCA with 35 components")
+plt.imshow(tf.reshape(x35[0], (28, 28)))
 
-plt.subplot(434)
+plt.subplot(345)
 plt.imshow(tf.reshape(x_train[1], (28, 28)))
-plt.subplot(435)
-plt.imshow(tf.reshape(x35[1], (28, 28)))
-plt.subplot(436)
+plt.subplot(346)
+plt.imshow(tf.reshape(x100[1], (28, 28)))
+plt.subplot(347)
 plt.imshow(tf.reshape(x50[1], (28, 28)))
+plt.subplot(348)
+plt.imshow(tf.reshape(x35[1], (28, 28)))
 
-plt.subplot(437)
+plt.subplot(3, 4, 9)
 plt.imshow(tf.reshape(x_train[2], (28, 28)))
-plt.subplot(438)
+plt.subplot(3, 4, 10)
 plt.imshow(tf.reshape(x35[2], (28, 28)))
-plt.subplot(439)
+plt.subplot(3, 4, 11)
 plt.imshow(tf.reshape(x50[2], (28, 28)))
+plt.subplot(3, 4, 12)
+plt.imshow(tf.reshape(x100[2], (28, 28)))
 
 plt.show()

### 2022-11-29 src/test_pca.py
Fix plots

@@ -12,6 +12,7 @@ from tensorflow.python.ops.numpy_ops import np_config
 
 np_config.enable_numpy_behavior()
 
+
 mnist = tf.keras.datasets.mnist
 
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
@@ -22,13 +23,13 @@ x_test_flat = x_test.reshape((x_test.shape[0], 784))
 x_train_flat = x_train_flat
 
 
-pca35 = PCA(n_components=35, svd_solver="randomized")
+pca10 = PCA(n_components=10, svd_solver="randomized")
 pca50 = PCA(n_components=50, svd_solver="randomized")
 pca100 = PCA(n_components=100, svd_solver="randomized")
 
-pca35.fit(x_train_flat)
-temp = pca35.transform(x_train_flat)
-x35 = pca35.inverse_transform(temp)
+pca10.fit(x_train_flat)
+temp = pca10.transform(x_train_flat)
+x10 = pca10.inverse_transform(temp)
 
 pca50.fit(x_train_flat)
 temp = pca50.transform(x_train_flat)
@@ -40,18 +41,19 @@ x100 = pca100.inverse_transform(temp)
 
 
 # ------- plotting pca ------------
+plt.tick_params(colors="1")
 plt.subplot(341)
-plt.title("Actual instance, \n 784 features")
+plt.title("Actual instance, \n 784 features", size=22)
 plt.imshow(tf.reshape(x_train[0], (28, 28)))
 plt.subplot(342)
-plt.title("PCA with 100 components")
+plt.title("PCA with \n 100 components", size=22)
 plt.imshow(tf.reshape(x100[0], (28, 28)))
 plt.subplot(343)
-plt.title("PCA with 50 components")
+plt.title("PCA with \n 50 components", size=22)
 plt.imshow(tf.reshape(x50[0], (28, 28)))
 plt.subplot(344)
-plt.title("PCA with 35 components")
-plt.imshow(tf.reshape(x35[0], (28, 28)))
+plt.title("PCA with \n 10 components", size=22)
+plt.imshow(tf.reshape(x10[0], (28, 28)))
 
 plt.subplot(345)
 plt.imshow(tf.reshape(x_train[1], (28, 28)))
[…]

### 2022-11-29 src/utils.py
Add confusion matrxi

@@ -6,6 +6,25 @@ import numpy as np
 def PCA_stoch(X, n_components):
     print("eric")
 
+def plot_confusion(confusion_matrix: np.ndarray, title=None):
+    fontsize = 40
+
+    sns.set(font_scale=3)
+    sns.heatmap(
+        confusion_matrix,
+        annot=True,
+        fmt=".2%",
+        cmap="Blues",
+    )
+    if title:
+        plt.title(title)
+    else:
+        plt.title("Confusion matrix")
+
+    plt.xlabel("Predicted class")
+    plt.ylabel("True class")
+    plt.show()
+
 
 def PCA_stoch(X, n_components, iterations=10, eta=1e-5, convergence=1e-5, epochs=10):
     print(f"{X.shape=}")

### 2022-11-29 src/utils.py
Adding funcs to utils.py

@@ -81,6 +81,17 @@ class PCALayer(layers.Layer):
         return tf.matmul(inputs, self.W)
 
 
+def conf_mat(preds, labl, numn_cls): 
+    return tf.math.confusion_matrix(labels=labl, predictions=preds, num_classes=num_cls)
+
+def perc(matrix): 
+    conf = np.zeros(matrix.shape) 
+
+    for i in range(matrix.shape[0]): 
+        for j in range(matrix.shape[1]): 
+            conf[i][j] = matrix[i][j] / tf.reduce_sum(matrix[i], 0).numpy()
+    return conf
+
 # PCA implemented using tensors, to be able to run on gpu
 
 # class PCA:

### 2022-11-29 src/conv_nn_mnist.py
Adding figures

@@ -8,7 +8,7 @@ import matplotlib.pyplot as plt
 from utils import *
 from sklearn.model_selection import KFold
 import keras_tuner as kt
-
+import seaborn as sns
 tf.keras.utils.set_random_seed(1336)
 batch_size = 128
 epochs = 6
@@ -35,10 +35,10 @@ def model_builder(hp):
     model = tf.keras.Sequential(
         [
             layers.Rescaling(1.0 / 255),
-            layers.Conv2D(32, hp_width, activation="relu", padding="same"),
-            layers.MaxPooling2D(),
-            layers.Conv2D(32, hp_width, activation="relu", padding="same"),
-            layers.MaxPooling2D(),
+            # layers.Conv2D(32, hp_width, activation="relu", padding="same"),
+            # layers.MaxPooling2D(),
+            # layers.Conv2D(32, hp_width, activation="relu", padding="same"),
+            # layers.MaxPooling2D(),
             layers.Conv2D(32, hp_width, activation="relu", padding="same"),
             layers.MaxPooling2D(),
             layers.Flatten(),
@@ -73,52 +73,46 @@ learning_rate = best_hps.get("learning_rate")
 print(width)
 print(learning_rate)
 
-kf = KFold(n_splits=5)  # random_state=1336)
 
+model = tf.keras.Sequential(
+    [
+        layers.Rescaling(1.0 / 255),
+        # layers.Conv2D(32, width, activation="relu", padding='same'),
+        # layers.MaxPooling2D(),
+        # layers.Conv2D(32, width, activation="relu", padding='same'),
+        # layers.MaxPooling2D(),
+        layers.Conv2D(32, width, activation="relu", padding='same'),
+        layers.MaxPooling2D(),
+        layers.Flatten(),
+        layers.Dense(128, activation="relu"),
+        layers.Dense(10),
+    ]
+)
 
-avg_accuracy = 0
-for train_index, test_index in kf.split(x_train, y_train):
-    train_start = train_index[0]
-    train_stop = train_index[-1]
+model.compile(
+    optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
+    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+    metrics=["accuracy"],
+)
 
-    test_start = test_index[0]
-    test_stop = test_index[-1]
+model.fit(
+    x_train,
+    y_train,
+    validation_data=(x_test, y_test),
+    batch_size=batch_size,
[…]

### 2022-11-29 src/pca_nn_mnist.py

@@ -70,56 +70,50 @@ print(units)
 print(learning_rate)
 folds = 5
 
-kf = KFold(n_splits=folds)  # random_state=1336)
 
-avg_accuracy = 0
-for train_index, test_index in kf.split(x_train_flat, y_train):
-    train_start = train_index[0]
-    train_stop = train_index[-1]
+pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
+pca.fit(x_train_flat)
 
-    test_start = test_index[0]
-    test_stop = test_index[-1]
+x_train_pca = pca.transform(x_train_flat)
+x_test_pca = pca.transform(x_test_flat)
 
-    pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
+y_train_cv = y_train
 
-    pca.fit(x_train_flat[train_start:train_stop])
 
-    x_train_pca = pca.transform(x_train_flat[train_start:train_stop])
-    x_holdout_pca = pca.transform(x_train_flat[test_start:test_stop])
+model = tf.keras.Sequential(
+    [
+        layers.Dense(units, activation="relu"),
+        layers.Dense(units, activation="relu"),
+        layers.Dense(units, activation="relu"),
+        layers.Dense(10),
+    ]
+)
 
-    y_train_cv = y_train[train_start:train_stop]
-    y_holdout = y_train[test_start:test_stop]
+model.compile(
+    optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
+    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+    metrics=["accuracy"],
+)
 
-    model = tf.keras.Sequential(
-        [
-            layers.Dense(units, activation="relu"),
-            layers.Dense(units, activation="relu"),
-            layers.Dense(units, activation="relu"),
-            layers.Dense(10),
-        ]
-    )
+model.fit(
+    x_train_pca,
+    y_train,
+    validation_data=(x_test_pca, y_test),
+    epochs=6,
+    batch_size=64,
+)
 
-    model.compile(
-        optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
-        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
-        metrics=["accuracy"],
-    )
+results = model.evaluate(
+    x_test_pca,
+    y_test,
+    return_dict=True,
+)
 
-    model.fit(
[…]

### 2022-11-29 src/utils.py

@@ -1,7 +1,9 @@
 import tensorflow as tf
 from tensorflow.keras import layers
 import numpy as np
-
+import seaborn as sns
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm 
 
 def PCA_stoch(X, n_components):
     print("eric")
@@ -9,12 +11,14 @@ def PCA_stoch(X, n_components):
 def plot_confusion(confusion_matrix: np.ndarray, title=None):
     fontsize = 40
 
-    sns.set(font_scale=3)
+    sns.set(font_scale=1.5)
     sns.heatmap(
         confusion_matrix,
         annot=True,
         fmt=".2%",
         cmap="Blues",
+        norm=LogNorm(),
+
     )
     if title:
         plt.title(title)
@@ -81,7 +85,7 @@ class PCALayer(layers.Layer):
         return tf.matmul(inputs, self.W)
 
 
-def conf_mat(preds, labl, numn_cls): 
+def conf_mat(preds, labl, num_cls): 
     return tf.math.confusion_matrix(labels=labl, predictions=preds, num_classes=num_cls)
 
 def perc(matrix): 

### 2022-11-29 src/pca_rf_pneu.py
Small changes

@@ -41,7 +41,7 @@ i = 0
 for batch, _ in train_ds:
     x_list.append(tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0)
     i += 1
-    if i > 40:
+    if i > 20:
         break
 
 X_subset = tf.concat(x_list, axis=0)

### 2022-11-30 src/pca_rf_mnist.py
add docs

@@ -10,6 +10,9 @@ import tensorflow_decision_forests as tfdf
 from sklearn.decomposition import PCA
 
 tf.keras.utils.set_random_seed(1336)
+"""
+PCA random forest used for MNIST dataset. Builds and fits data.
+"""
 
 batch_size = 128
 epochs = 6

### 2022-11-30 src/conv_rf_mnist.py
more docs

@@ -9,6 +9,9 @@ from utils import *
 import tensorflow_decision_forests as tfdf
 
 tf.keras.utils.set_random_seed(1336)
+"""
+Convolutional random forest used for MNIST dataset. Builds and fits data.
+"""
 
 batch_size = 128
 epochs = 1
@@ -66,7 +69,10 @@ features_test = val_ds.map(lambda batch, label: (feature_extractor(batch), label
 # tuner.choice("max_depth", [4, 6, 8, 10, 12, 14, 16, 18, 20])
 
 forest = tfdf.keras.RandomForestModel(
-    verbose=1, max_depth=25, random_seed=1337, num_trees=300#, tuner=tuner#, check_dataset=False
+    verbose=1,
+    max_depth=25,
+    random_seed=1337,
+    num_trees=300,  # , tuner=tuner#, check_dataset=False
 )
 forest.fit(x=features_train)
 

### 2022-11-30 src/conv_rf_pneu.py

@@ -8,6 +8,11 @@ import matplotlib.pyplot as plt
 from utils import *
 import tensorflow_decision_forests as tfds
 
+tf.keras.utils.set_random_seed(1336)
+"""
+Convolutional random forest used for pneumonia dataset. Builds and fits data.
+"""
+
 batch_size = 128
 epochs = 3
 mnist = tf.keras.datasets.mnist

### 2022-11-30 src/conv_nn_better.py
Convolution models

@@ -3,11 +3,13 @@ import os
 from tensorflow.keras import layers
 from tensorflow.keras import regularizers
 import tensorflow_datasets as tfds
-
+import keras_tuner as kt
 filedir = os.path.dirname(__file__)
 from functools import partial
 
-tf.keras.utils.set_random_seed(1336)
+seed = 1337
+tf.keras.utils.set_random_seed(seed)
+tf.config.experimental.enable_op_determinism()
 
 TRAINDIR = filedir + "/../data/chest_xray/train"
 TESTDIR = filedir + "/../data/chest_xray/test"
@@ -17,8 +19,20 @@ IMG_WIDTH = 227
 
 train_ds = tf.keras.utils.image_dataset_from_directory(
     TRAINDIR,
+    validation_split=0.1, 
+    subset="training",
     labels="inferred",
-    seed=1337,
+    seed=seed,
+    image_size=(IMG_HEIGHT, IMG_WIDTH),
+    batch_size=BATCHSIZE,
+    color_mode="grayscale",
+)
+val_ds = tf.keras.utils.image_dataset_from_directory(
+    TRAINDIR,
+    validation_split=0.1, 
+    subset="validation",
+    labels="inferred",
+    seed=seed,
     image_size=(IMG_HEIGHT, IMG_WIDTH),
     batch_size=BATCHSIZE,
     color_mode="grayscale",
@@ -27,7 +41,7 @@ train_ds = tf.keras.utils.image_dataset_from_directory(
 test_ds = tf.keras.utils.image_dataset_from_directory(
     TESTDIR,
     labels="inferred",
-    seed=1337,
+    seed=seed,
     image_size=(IMG_HEIGHT, IMG_WIDTH),
     batch_size=BATCHSIZE,
     color_mode="grayscale",
@@ -48,7 +62,7 @@ test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
 
 
 def model_builder(hp):
-    hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
+    hp_lambda = hp.Choice("lambda", values=[0.0, 1e-5, 1e-4, 1e-3])
     hp_learning_rate = hp.Choice("learning_rate", values=[1e-3, 1e-2, 1e-1])
 
     kernel = tf.keras.regularizers.L2(l2=hp_lambda)
@@ -63,8 +77,6 @@ def model_builder(hp):
                 strides=(4, 4),
                 padding="valid",
                 activation="relu",
[…]

### 2022-11-30 src/conv_nn_pneu.py

@@ -4,8 +4,11 @@ from tensorflow.keras import layers
 from tensorflow.keras import regularizers
 import tensorflow_datasets as tfds
 filedir = os.path.dirname(__file__)
+import numpy as np
 
-tf.keras.utils.set_random_seed(1336)
+seed = 1336
+tf.keras.utils.set_random_seed(seed)
+tf.config.experimental.enable_op_determinism()
 
 TRAINDIR = filedir + "/../data/chest_xray/train"
 TESTDIR = filedir + "/../data/chest_xray/test"

### 2022-11-30 src/conv_nn_better.py
Fixed conv_nn_better.py to be a deterministic code process

@@ -7,7 +7,7 @@ import keras_tuner as kt
 filedir = os.path.dirname(__file__)
 from functools import partial
 
-seed = 1337
+seed = 1336
 tf.keras.utils.set_random_seed(seed)
 tf.config.experimental.enable_op_determinism()
 
@@ -22,7 +22,7 @@ train_ds = tf.keras.utils.image_dataset_from_directory(
     validation_split=0.1, 
     subset="training",
     labels="inferred",
-    seed=seed,
+    seed=1337,
     image_size=(IMG_HEIGHT, IMG_WIDTH),
     batch_size=BATCHSIZE,
     color_mode="grayscale",
@@ -32,7 +32,7 @@ val_ds = tf.keras.utils.image_dataset_from_directory(
     validation_split=0.1, 
     subset="validation",
     labels="inferred",
-    seed=seed,
+    seed=1337,
     image_size=(IMG_HEIGHT, IMG_WIDTH),
     batch_size=BATCHSIZE,
     color_mode="grayscale",
@@ -41,7 +41,7 @@ val_ds = tf.keras.utils.image_dataset_from_directory(
 test_ds = tf.keras.utils.image_dataset_from_directory(
     TESTDIR,
     labels="inferred",
-    seed=seed,
+    seed=1337,
     image_size=(IMG_HEIGHT, IMG_WIDTH),
     batch_size=BATCHSIZE,
     color_mode="grayscale",
@@ -147,7 +147,7 @@ tuner = kt.Hyperband(
     directory="conv_nn_pneu-params",
 )
 
-tuner.search(train_ds, epochs=10, validation_data=val_ds)
+tuner.search(train_ds, epochs=10, validation_data=val_ds, class_weight=class_weight)
 best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
 lam = best_hps.get("lambda")
 learning_rate = best_hps.get("learning_rate")

### 2022-11-30 src/pca_nn_pneu.py
add yolo based algorithm

@@ -3,6 +3,7 @@ import os
 import numpy as np
 from tensorflow.keras import layers
 from tensorflow.keras import regularizers
+import keras_tuner as kt
 from PIL import Image
 import time
 import matplotlib.pyplot as plt
@@ -16,114 +17,187 @@ tf.keras.utils.set_random_seed(1336)
 PCA neural network used for pneumonia dataset. Builds and fits data, takes time.
 """
 
-TRAINDIR = filedir + "/../data/chest_xray/train"
-TESTDIR = filedir + "/../data/chest_xray/test"
-BATCHSIZE = 128
-IMG_HEIGHT = 200
-IMG_WIDTH = 200
+def model_builder(hp):
+    hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
+    hp_learning_rate = hp.Choice("learning_rate", values=[1e-3, 1e-2, 1e-1])
 
-train_ds = tf.keras.utils.image_dataset_from_directory(
-    TRAINDIR,
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=128,
-    color_mode="grayscale",
-)
+    kernel = tf.keras.regularizers.L2(l2=hp_lambda)
+    bias = tf.keras.regularizers.L2(l2=hp_lambda)
 
-val_ds = tf.keras.utils.image_dataset_from_directory(
-    TESTDIR,
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=128,
-    color_mode="grayscale",
-)
+    model = tf.keras.Sequential(
+        [
+            layers.Dense(
+                n_components,
+                kernel_regularizer=kernel,
+                bias_regularizer=bias,
+                activation="relu",
+            ),
+            layers.Dense(
+                n_components,
+                kernel_regularizer=kernel,
+                bias_regularizer=bias,
+                activation="relu",
+            ),
+            layers.Dense(
+                n_components,
+                kernel_regularizer=kernel,
+                bias_regularizer=bias,
+                activation="relu",
+            ),
+            layers.Dense(1),
+        ]
+    )
+
+    model.compile(
+        optimizer=tf.keras.optimizers.Adamax(learning_rate=hp_learning_rate),
+        loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
+        metrics=["accuracy"],
[…]

### 2022-11-30 src/pca_nn_pneu.py
This is jonatan on the pc rn

@@ -144,7 +144,7 @@ with tf.device('/gpu:0'):
         objective="val_accuracy",
         max_epochs=10,
         factor=3,
-        directory="conv_nn_pneu-params",
+        directory="pca_nn_pneu-params",
     )
 
     tuner.search(

### 2022-11-30 src/conv_exp.py
File using ResNet architecture with params set to achieve 89% accuracy

@@ -1,0 +1,83 @@
+import tensorflow as tf
+import os
+from tensorflow.keras import layers
+from tensorflow.keras import regularizers
+import tensorflow_datasets as tfds
+filedir = os.path.dirname(__file__)
+import numpy as np
+from CNN_models import DefaultConv2D, ResidualUnit
+"""
+Convolutional neural network used for pneumonia dataset. Builds and fits data.
+"""
+seed = 1337
+tf.keras.utils.set_random_seed(seed)
+tf.config.experimental.enable_op_determinism()
+
+TRAINDIR = filedir + "/../data/chest_xray/train"
+TESTDIR = filedir + "/../data/chest_xray/test"
+BATCHSIZE = 4
+IMG_HEIGHT = 227
+IMG_WIDTH = 227
+
+train_ds = tf.keras.utils.image_dataset_from_directory(
+    TRAINDIR,
+    labels="inferred",
+    seed=seed,
+    image_size=(IMG_HEIGHT, IMG_WIDTH),
+    batch_size=BATCHSIZE,
+    color_mode="grayscale",
+)
+
+test_ds = tf.keras.utils.image_dataset_from_directory(
+    TESTDIR,
+    labels="inferred",
+    seed=seed,
+    image_size=(IMG_HEIGHT, IMG_WIDTH),
+    batch_size=BATCHSIZE,
+    color_mode="grayscale",
+)
+
+COUNT_NORMAL = 1071
+COUNT_PNEUMONIA = 3114
+
+weight_for_0 = (1 / COUNT_NORMAL) * (COUNT_NORMAL + COUNT_PNEUMONIA) / 2.0
+weight_for_1 = (1 / COUNT_PNEUMONIA) * (COUNT_NORMAL + COUNT_PNEUMONIA) / 2.0
+
+class_weight = {0: weight_for_0, 1: weight_for_1}
+
+AUTOTUNE = tf.data.AUTOTUNE
+
+train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
+test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
+
+model = tf.keras.Sequential([
+    #tf.keras.layers.Rescaling(1.0/255),
+    DefaultConv2D(64, kernel_size=7, strides=2),
+    tf.keras.layers.BatchNormalization(), 
+    tf.keras.layers.Activation('relu'),
+    tf.keras.layers.MaxPool2D(pool_size=3, strides=2, padding='same'),
+    ])
[…]

### 2022-12-01 doc/paper.tex
More results

@@
 \section{Introduction}
+
+Some might say that Minsky himself used PCA back in the stone age, but we have made numerous improvements to the PCA algorithm:
+
+\begin{itemize}
+  \item - XD random
+  \item - \cite{halko}
+\end{itemize}
 
@@
 Each image is labeled with one of ten classes, corresponding to the digit it depicts.
-The dataset is divided into $60000$ training images and $10000$ validation images.
+The dataset is divided into $60000$ training images and $10000$ test images.
 This dataset is essentially uniformly distributed, with all classes containing around $9 - 11\%$ of the total dataset.
@@
 The images were taken from the Woman and Children's Medical Center in Guangzhou, China.
-The dataset is divided into a training set of $5232$ images and $624$ validation images.
-The dataset is not evenly distributed, having $74.2\%$ instances of pneumonia in the training set and $62.5\%$ instances of pneumonia in the validation set.
+The dataset is divided into a training set of $5232$ images and $624$ test images.
+The dataset is not evenly distributed, having $74.2\%$ instances of pneumonia in the training set and $62.5\%$ instances of pneumonia in the testing set.
+The data was collected and labeled as a part of the article {\it Identifying Medical Diagnoses and Treatable Diseases by Image-Based Deep Learning}, published in {\it Cell} \cite{xray}.
 
@@
 
+For hyperparameter tuning, we used the Keras Tuner library \cite{kerastuner}.
+
 \section{Results}
@@
 
+To test our hybrid models against traditional methods, we applied them all to the MNIST dataset.
+
 \subsubsection{Traditional Convolutional Neural Network}
+
+As is well known, convolutional neural networks perform excellent on the MNIST dataset, achieving test accuracies rivaling humans.
+As an initial guess for the architecture, we used only a single convolutional layer\footnote{We follow the terminology used in Goodfellow et Al.
+\cite[336]{gbc}, where what is denoted as a single convolutional layer includes the convolutional, detector and pooling stage} followed by a dense layer.
+The convolutional layer had $32$ filters.
+By gridsearch, we found the optimal filter width to be $7$ and the optimal learning rate to be $0.001$.
+With these hyperparameters we achieved a test accuracy of $98.8\%$, a very good result.
+Based on this result, we conclude that no more hidden layers or regularization parameters need to be included.
+Looking at the heatmap (Fig.~\ref{mnistheatmap_cnn}), we can see that our predictions are excellent in all classes, never misclassifying any single class more than $2.4\%$ of the time.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=8in]{figure/conf_cnn_MNIST.png}}
+\caption{Confusion matrix for MNIST dataset classified using a CNN.
+From this figure, we can see that most misclassifications occur for digits that look similar: eights are sometimes misidentified as zeroes, and fours and nines are sometimes misidentified as eachother.}
+\label{mnistheatmap_cnn}
+\end{figure}
 
@@
 
+Given that we have achieved excellent results by using a CNN, we may don't expect to achieve much better results by sending the extracted features to a random forest.
+Regardless, our testing found that by using the same CNN model we found previously,
+
 \subsubsection{Principal Component Analysis followed by a Neural Network}
+
+Replacing the convolutional stage with a simpler form of dimensionality reduction, namely PCA, we can see that we are able to reduce the dimensionality of our dataset while retaining a large amount of the information.
+Looking at figure~\ref{pcacomp}, we can see that even when using only the $50$ first principal components, the numbers are still readable.
+From this, we can assume that we can reduce the amount of features sent into our neural net from $784$ to less than $100$ features, and still expect reasonable results.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=5in]{figure/pcacomp.png}}
+\caption{Comparison showing the amount of information lost by reducing the amount of principal components}
+\label{pcacomp}
+\end{figure}
+
+Using different amount of principal components, we sent the reduced dataset into a neural network with three hidden layers.
+After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $97.8\%$.
+This is comparable to the accuracy obtained by a traditional convolutional neural network.
+When considering the computational effort, we need to consider the performance benefit that modern GPU's bring, being specifically designed to run neural networks efficiently.
+This is very apparent in our testing: When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds.
+This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time.
+However, when using a GPU with CUDA cores, the story is quite different: Here, the CNN used only $6$ seconds, while the PCA-NN model took $12$ seconds\footnote{The fact that this model took longer than the CNN is most likely because the PCA function we use only runs on the CPU, meaning that only the neural net part is actually running on the GPU}.
+Looking at the confusion matrix (Fig.~\ref{mnistheatmap_pcann}), we can see that we are again achieving very few misclassifications using the PCA-NN model.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=8in]{figure/conf_pca_nn_MNIST.png}}
+\caption{Confusion matrix for MNIST dataset classified using PCA followed by a neural network.
+Here we can see that misclassifications are slightly more common than when using a traditional CNN, but the results are still good.}
+\label{mnistheatmap_pcann}
+\end{figure}
 
@@
 
+Satisfied with the results achieved by replacing the convolutional stage with Principal Component Analysis, we attempt to sidestep neural networks all together, and use PCA followed by a random forest.
+Here, we achieved a reasonable result of $94.7\%$ accuracy on the test set.
+Although this is not as impressive as the previous models, there
+
 \subsection{Guangzhou Pneumonia Data}
 
-Hopefully this is real
+As previously stated, the Guangzhou Pneumonia Dataset was collected and labeled as part of the article {\it Identifying Medical Diagnoses and Treatable Diseases by Image-Based Deep Learning}.
+In this article, the authors used transfer learning using a network previously trained on ImageNet to classify the x-ray images.
+With this method, they achieved a test accuracy of $92.8\%$ \cite[1127]{xray}.
+We now turn our attention to this dataset, and see if we can achieve similar results using our methods.
 
 \subsubsection{Traditional Convolutional Neural Network}
+
+Determining whether an x-ray shows a case of pneumonia or not is clearly a much more difficult classification problem than our previous dataset.
+Therefore, we initially had problems designing an architecture which achieved much higher than $80\%$ test accuracy.
+Clearly, we are not able to use a simple one convolutional layer model like we were able to on the MNIST dataset, and designing a general convolutional neural network architecture is a very complex task.
+Thus, we turn to the literature, and explore models which have been proven to be effective at image recognition.
+Out of the architectures presented in BOOK \cite{geron}, we decided to try AlexNet, which won the ILSVRC 2012 challenge and is a relatively shallow network having $11$ hidden layers.
+Using this model, we achieved a testing accuracy of $87.8\%$.
+Comparing our result to those obtained by Kermany et Al., we see that our model is slightly more ass.
 
@@
 
+% \begin{table}[h]
+% \caption{Accuracy by }
+% \begin{center}
+% \label{datasettable}
+% \begin{tabular}{| c | l | l | l |}
+% \hline
+% Dataset & Number of instances & Features & Type \\
+% \hline
+% Franke Function (raw data) & 400 & 2 & regression \\
+% Franke Function (polynomial features) & 400 & 56 & regression \\
+% Wisconsin Breast Cancer & 569 & 30 & binary classification \\
+% MNIST 8x8 & 1797 & 64 & multiclass classification (10 classes) \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
+
 I have realized that I am cooler than Eric.
+
+The children of Guangzhou province were saved to a certain extent. They all lived happily ever after.
 

### 2022-12-08 src/conv_rf_pneu.py
Try new method

@@ -237,7 +237,7 @@ model.evaluate(test_ds, batch_size=BATCHSIZE)
 
 feature_extractor = tf.keras.Model(
     inputs=model.inputs,
-    outputs=model.layers[-4].output,
+    outputs=model.layers[-1].output,
 )
 
 features_train = train_ds.map(lambda batch, label: (feature_extractor(batch), label))

### 2022-12-08 src/pca_nn_pneu.py
Remove bad code

@@ -17,6 +17,7 @@ tf.keras.utils.set_random_seed(1336)
 PCA neural network used for pneumonia dataset. Builds and fits data, takes time.
 """
 
+
 def model_builder(hp):
     hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
     hp_learning_rate = hp.Choice("learning_rate", values=[1e-3, 1e-2, 1e-1])
@@ -56,6 +57,7 @@ def model_builder(hp):
 
     return model
 
+
 with tf.device("/cpu:0"):
     TRAINDIR = filedir + "/../data/chest_xray/train"
     TESTDIR = filedir + "/../data/chest_xray/test"
@@ -94,9 +96,6 @@ with tf.device("/cpu:0"):
     train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
     test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
 
-
-
-# grab a subset for PCA calculation
     x_list = []
     i = 0
     X_train = None
@@ -128,7 +127,6 @@ with tf.device("/cpu:0"):
             )
             y_test = tf.concat([y_test, labels], axis=0)
 
-
     pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
 
     print("Fitting PCA")
@@ -138,7 +136,7 @@ with tf.device("/cpu:0"):
     pca.transform(X_train)
     pca.transform(X_test)
 
-with tf.device('/gpu:0'):
+with tf.device("/gpu:0"):
     tuner = kt.Hyperband(
         model_builder,
         objective="val_accuracy",
@@ -186,7 +184,7 @@ with tf.device('/gpu:0'):
     model.compile(
         optimizer=tf.keras.optimizers.Adamax(learning_rate=learning_rate),
         loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
-        metrics=["accuracy", "precision", "recall"],
+        metrics=["accuracy"],
     )
 
     model.fit(

### 2022-12-08 src/pca_rf_pneu.py
Add pca rf pneu new version

@@ -2,75 +2,146 @@ import tensorflow as tf
 import os
 import numpy as np
 from tensorflow.keras import layers
+from tensorflow.keras import regularizers
+import keras_tuner as kt
 from PIL import Image
 import time
 import matplotlib.pyplot as plt
 from utils import *
-import tensorflow_decision_forests as tfds
+from sklearn.decomposition import PCA
 
 filedir = os.path.dirname(__file__)
 
 tf.keras.utils.set_random_seed(1336)
 """
-PCA random forest used for pneumonia dataset. Builds and fits data.
+PCA neural network used for pneumonia dataset. Builds and fits data, takes time.
 """
 
-TRAINDIR = filedir + "/../data/chest_xray/train"
-TESTDIR = filedir + "/../data/chest_xray/test"
-BATCHSIZE = 128
-IMG_HEIGHT = 200
-IMG_WIDTH = 200
 
-train_ds = tf.keras.utils.image_dataset_from_directory(
-    TRAINDIR,
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=128,
-    color_mode="grayscale",
-)
+def model_builder(hp):
+    hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
+    hp_learning_rate = hp.Choice("learning_rate", values=[1e-3, 1e-2, 1e-1])
 
-val_ds = tf.keras.utils.image_dataset_from_directory(
-    TESTDIR,
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=128,
-    color_mode="grayscale",
-)
+    kernel = tf.keras.regularizers.L2(l2=hp_lambda)
+    bias = tf.keras.regularizers.L2(l2=hp_lambda)
 
+    model = tf.keras.Sequential(
+        [
+            layers.Dense(
+                n_components,
+                kernel_regularizer=kernel,
+                bias_regularizer=bias,
+                activation="relu",
+            ),
+            layers.Dense(
+                n_components,
+                kernel_regularizer=kernel,
+                bias_regularizer=bias,
+                activation="relu",
+            ),
+            layers.Dense(
+                n_components,
+                kernel_regularizer=kernel,
[…]

### 2022-12-08 src/pca_nn_pneu.py
Add checkpoint

@@ -187,6 +187,15 @@ with tf.device("/gpu:0"):
         metrics=["accuracy"],
     )
 
+    checkpoint_filepath = "/tmp/checkpoint"
+    model_checkpoint_callback = tf.keras.callbacks.ModelCheckpoint(
+        filepath=checkpoint_filepath,
+        save_weights_only=True,
+        monitor="val_accuracy",
+        mode="max",
+        save_best_only=True,
+    )
+
     model.fit(
         X_train,
         y_train,
@@ -194,32 +203,10 @@ with tf.device("/gpu:0"):
         epochs=6,
         batch_size=64,
         class_weight=class_weight,
+        callbacks=[model_checkpoint_callback],
     )
+    model.load_weights(checkpoint_filepath)
+
+    model.evaluate(test_ds, batch_size=BATCHSIZE)
+
     print(f"Time taken: {time.time() - start}")
-
-# ------- plotting pca ------------
-# x_train_remade = pca.inverse_transform(x_train_pca)
-
-# plt.subplot(421)
-# plt.title("Actual instance, \n 784 features")
-# plt.imshow(x_train[0])
-# plt.subplot(422)
-# plt.title(f"Constructed using only the \n {n_components} first principal components")
-# plt.imshow(tf.reshape(x_train_remade[0], (28, 28)))
-#
-# plt.subplot(423)
-# plt.imshow(x_train[1])
-# plt.subplot(424)
-# plt.imshow(tf.reshape(x_train_remade[1], (28, 28)))
-#
-# plt.subplot(425)
-# plt.imshow(x_train[2])
-# plt.subplot(426)
-# plt.imshow(tf.reshape(x_train_remade[2], (28, 28)))
-#
-# plt.subplot(427)
-# plt.imshow(x_train[3])
-# plt.subplot(428)
-# plt.imshow(tf.reshape(x_train_remade[3], (28, 28)))
-#
-# plt.show()

### 2022-12-08 src/pca_nn_pneu.py
Update to work

@@ -133,8 +133,8 @@ with tf.device("/cpu:0"):
     start = time.time()
     pca.fit(X_train)
 
-    pca.transform(X_train)
-    pca.transform(X_test)
+    X_train = pca.transform(X_train)
+    X_test = pca.transform(X_test)
 
 with tf.device("/gpu:0"):
     tuner = kt.Hyperband(

### 2022-12-08 src/pca_rf_pneu.py

@@ -9,6 +9,7 @@ import time
 import matplotlib.pyplot as plt
 from utils import *
 from sklearn.decomposition import PCA
+import tensorboard_decision_forests as tfdf
 
 filedir = os.path.dirname(__file__)
 
@@ -133,8 +134,8 @@ with tf.device("/cpu:0"):
     start = time.time()
     pca.fit(X_train)
 
-    pca.transform(X_train)
-    pca.transform(X_test)
+    X_train = pca.transform(X_train)
+    X_test = pca.transform(X_test)
 
     forest = tfdf.keras.RandomForestModel(
         verbose=1, max_depth=16, random_seed=1337, check_dataset=False

### 2022-12-08 src/pca_nn_pneu.py
GregSearch MVP god tier algorithm

@@ -21,6 +21,7 @@ PCA neural network used for pneumonia dataset. Builds and fits data, takes time.
 def model_builder(hp):
     hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
     hp_learning_rate = hp.Choice("learning_rate", values=[1e-3, 1e-2, 1e-1])
+    hp_units = hp.Choice("units", values=[9, 15, 17, 24, 30])
 
     kernel = tf.keras.regularizers.L2(l2=hp_lambda)
     bias = tf.keras.regularizers.L2(l2=hp_lambda)
@@ -28,19 +29,19 @@ def model_builder(hp):
     model = tf.keras.Sequential(
         [
             layers.Dense(
-                n_components,
+                hp_units,
                 kernel_regularizer=kernel,
                 bias_regularizer=bias,
                 activation="relu",
             ),
             layers.Dense(
-                n_components,
+                hp_units,
                 kernel_regularizer=kernel,
                 bias_regularizer=bias,
                 activation="relu",
             ),
             layers.Dense(
-                n_components,
+                hp_units,
                 kernel_regularizer=kernel,
                 bias_regularizer=bias,
                 activation="relu",
@@ -103,7 +104,7 @@ with tf.device("/cpu:0"):
     y_train = None
     y_test = None
 
-    n_components = 1000
+    n_components = 9
 
     for batch, labels in train_ds:
         if X_train is None:
@@ -151,28 +152,30 @@ with tf.device("/gpu:0"):
     best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
     lam = best_hps.get("lambda")
     learning_rate = best_hps.get("learning_rate")
+    units = best_hps.get("units")
 
     print(lam)
     print(learning_rate)
+    print(units)
 
     kernel = tf.keras.regularizers.L2(l2=lam)
     bias = tf.keras.regularizers.L2(l2=lam)
     model = tf.keras.Sequential(
         [
             layers.Dense(
-                n_components,
+                units,
                 kernel_regularizer=kernel,
                 bias_regularizer=bias,
                 activation="relu",
             ),
             layers.Dense(
-                n_components,
+                units,
                 kernel_regularizer=kernel,
                 bias_regularizer=bias,
                 activation="relu",
             ),
             layers.Dense(
-                n_components,
+                units,
                 kernel_regularizer=kernel,
[…]

### 2022-12-08 src/pca_nn_pneu.py
Add stuff

@@ -133,8 +133,8 @@ with tf.device("/cpu:0"):
     start = time.time()
     pca.fit(X_train)
 
-    X_train = pca.transform(X_train)
-    X_test = pca.transform(X_test)
+    X_train_pca = pca.transform(X_train)
+    X_test_pca = pca.transform(X_test)
 
 with tf.device("/gpu:0"):
     tuner = kt.Hyperband(
@@ -146,7 +146,7 @@ with tf.device("/gpu:0"):
     )
 
     tuner.search(
-        X_train, y_train, epochs=10, validation_split=0.2, class_weight=class_weight
+        X_train_pca, y_train, epochs=10, validation_split=0.2, class_weight=class_weight
     )
     best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
     lam = best_hps.get("lambda")
@@ -197,9 +197,9 @@ with tf.device("/gpu:0"):
     )
 
     model.fit(
-        X_train,
+        X_train_pca,
         y_train,
-        validation_data=(X_test, y_test),
+        validation_data=(X_test_pca, y_test),
         epochs=6,
         batch_size=64,
         class_weight=class_weight,
@@ -207,6 +207,16 @@ with tf.device("/gpu:0"):
     )
     model.load_weights(checkpoint_filepath)
 
-    model.evaluate(test_ds, batch_size=BATCHSIZE)
+    model.evaluate(X_test_pca, y_test, batch_size=BATCHSIZE)
+
+    X_train_pca_restored = pca.inverse_transform(X_test_pca)
+
+    plt.subplot(121)
+    plt.title("Actual instance 51k features", size=22)
+    plt.imshow(tf.reshape(X_train[0], (227, 227)))
+
+    plt.subplot(122)
+    plt.title("PCA with 9 components", size=22)
+    plt.imshow(tf.reshape(X_train_pca_restored[0], (227, 227)))
 
     print(f"Time taken: {time.time() - start}")

### 2022-12-08 src/pca_nn_pneu.py
Fix dimensions

@@ -63,8 +63,8 @@ with tf.device("/cpu:0"):
     TRAINDIR = filedir + "/../data/chest_xray/train"
     TESTDIR = filedir + "/../data/chest_xray/test"
     BATCHSIZE = 128
-    IMG_HEIGHT = 200
-    IMG_WIDTH = 200
+    IMG_HEIGHT = 227
+    IMG_WIDTH = 227
 
     train_ds = tf.keras.utils.image_dataset_from_directory(
         TRAINDIR,
@@ -203,7 +203,7 @@ with tf.device("/gpu:0"):
         X_train_pca,
         y_train,
         validation_data=(X_test_pca, y_test),
-        epochs=6,
+        epochs=15,
         batch_size=64,
         class_weight=class_weight,
         callbacks=[model_checkpoint_callback],

### 2022-12-08 src/pca_nn_pneu.py
More chad moves abound

@@ -108,23 +108,23 @@ with tf.device("/cpu:0"):
 
     for batch, labels in train_ds:
         if X_train is None:
-            X_train = tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0
+            X_train = tf.reshape(batch, shape=[-1, IMG_WIDTH * IMG_HEIGHT]) / 255.0
             y_train = labels
 
         else:
             X_train = tf.concat(
-                [X_train, (tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0)], axis=0
+                [X_train, (tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0)], axis=0
             )
             y_train = tf.concat([y_train, labels], axis=0)
 
     for batch, labels in test_ds:
         if X_test is None:
-            X_test = tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0
+            X_test = tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0
             y_test = labels
 
         else:
             X_test = tf.concat(
-                [X_test, (tf.reshape(batch, shape=[-1, 200 * 200]) / 255.0)], axis=0
+                [X_test, (tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0)], axis=0
             )
             y_test = tf.concat([y_test, labels], axis=0)
 
@@ -221,5 +221,6 @@ with tf.device("/gpu:0"):
     plt.subplot(122)
     plt.title("PCA with 9 components", size=22)
     plt.imshow(tf.reshape(X_train_pca_restored[0], (227, 227)))
+    plt.show()
 
     print(f"Time taken: {time.time() - start}")

### 2022-12-08 src/pca_nn_pneu.py
Add more plots

@@ -113,7 +113,11 @@ with tf.device("/cpu:0"):
 
         else:
             X_train = tf.concat(
-                [X_train, (tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0)], axis=0
+                [
+                    X_train,
+                    (tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0),
+                ],
+                axis=0,
             )
             y_train = tf.concat([y_train, labels], axis=0)
 
@@ -124,7 +128,11 @@ with tf.device("/cpu:0"):
 
         else:
             X_test = tf.concat(
-                [X_test, (tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0)], axis=0
+                [
+                    X_test,
+                    (tf.reshape(batch, shape=[-1, IMG_HEIGHT * IMG_WIDTH]) / 255.0),
+                ],
+                axis=0,
             )
             y_test = tf.concat([y_test, labels], axis=0)
 
@@ -211,16 +219,29 @@ with tf.device("/gpu:0"):
     model.load_weights(checkpoint_filepath)
 
     model.evaluate(X_test_pca, y_test, batch_size=BATCHSIZE)
+    print(f"Time taken: {time.time() - start}")
 
     X_train_pca_restored = pca.inverse_transform(X_test_pca)
 
-    plt.subplot(121)
-    plt.title("Actual instance 51k features", size=22)
+    plt.subplot(421)
+    plt.title(f"Actual instance 51k features. Class {Y_train[0]}", size=18)
     plt.imshow(tf.reshape(X_train[0], (227, 227)))
-
-    plt.subplot(122)
-    plt.title("PCA with 9 components", size=22)
+    plt.subplot(422)
+    plt.title("PCA with 9 components", size=18)
     plt.imshow(tf.reshape(X_train_pca_restored[0], (227, 227)))
-    plt.show()
 
-    print(f"Time taken: {time.time() - start}")
+    plt.subplot(421)
+    plt.title(f"Actual instance 51k features. Class {Y_train[1]}", size=18)
+    plt.imshow(tf.reshape(X_train[1], (227, 227)))
+    plt.subplot(422)
+    plt.title("PCA with 9 components", size=18)
+    plt.imshow(tf.reshape(X_train_pca_restored[1], (227, 227)))
+
[…]

### 2022-12-08 src/pca_nn_pneu.py
Just some params

@@ -149,13 +149,13 @@ with tf.device("/gpu:0"):
     tuner = kt.Hyperband(
         model_builder,
         objective="val_accuracy",
-        max_epochs=10,
+        max_epochs=20,
         factor=3,
         directory="pca_nn_pneu-params",
     )
 
     tuner.search(
-        X_train_pca, y_train, epochs=10, validation_split=0.2, class_weight=class_weight
+        X_train_pca, y_train, epochs=20, validation_split=0.2, class_weight=class_weight
     )
     best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
     lam = best_hps.get("lambda")
@@ -212,7 +212,7 @@ with tf.device("/gpu:0"):
         y_train,
         validation_data=(X_test_pca, y_test),
         epochs=15,
-        batch_size=64,
+        batch_size=BATCHSIZE, #64,
         class_weight=class_weight,
         callbacks=[model_checkpoint_callback],
     )
@@ -223,25 +223,25 @@ with tf.device("/gpu:0"):
 
     X_train_pca_restored = pca.inverse_transform(X_test_pca)
 
-    plt.subplot(421)
-    plt.title(f"Actual instance 51k features. Class {Y_train[0]}", size=18)
+    plt.subplot(321)
+    plt.title(f"Actual instance 51k features. Class {y_train[0]}", size=18)
     plt.imshow(tf.reshape(X_train[0], (227, 227)))
-    plt.subplot(422)
+    plt.subplot(322)
     plt.title("PCA with 9 components", size=18)
     plt.imshow(tf.reshape(X_train_pca_restored[0], (227, 227)))
 
-    plt.subplot(421)
-    plt.title(f"Actual instance 51k features. Class {Y_train[1]}", size=18)
-    plt.imshow(tf.reshape(X_train[1], (227, 227)))
-    plt.subplot(422)
+    plt.subplot(323)
+    plt.title(f"Actual instance 51k features. Class {y_train[5]}", size=18)
+    plt.imshow(tf.reshape(X_train[5], (227, 227)))
+    plt.subplot(324)
     plt.title("PCA with 9 components", size=18)
-    plt.imshow(tf.reshape(X_train_pca_restored[1], (227, 227)))
+    plt.imshow(tf.reshape(X_train_pca_restored[5], (227, 227)))
 
-    plt.subplot(421)
[…]

### 2022-12-08 doc/paper.tex
Update paper

@@
 
-We found a dog driving a car down Hollywood Boulevard.
+% We found a dog driving a car down Hollywood Boulevard.
 It had sunglasses on, but didn't know where to go due to demensionality reduction.
 It was computing many computinations but due to severe liminations it had problems with binaly classifination and a large emesee.
+Kurwa.
 
@@
 
-The eponymous operation of convolution is a method which is used much in image processing.
+The eponymous operation of convolution is a method which is often used in image processing.
 It is performed by using a convolutional kernel, which is applied to each pixel of the image.
@@
 A basic kernel could be the Laplacian kernel (Eq.~\ref{laplace}), which will output a high value when the pixel in the middle of the kernel has a value which differs greatly from the pixels in a 3 by 3 neighbourhood around it.
+As such, this kernel can be used to detect edges, as edges imply large local changes in pixel intensity.
+The effect of this kernel can be seen in figure ~\ref{millie}.
 
@@
 
+\begin{figure}[H]
+\centerline{\includegraphics[width=8in]{figure/millie.png}}
+\caption{The result of applying the kernel in Eq.~\ref{laplace} to an image.
+Here we can see that areas of relatively constant intensity disappear (the floor, the back part of the box), while sharp edges elicit a response (the transition between carpet and floor, the holes in the box).
+However, a lot of noise is also present.
+This can be solved with more sophisticated methods of edge detection, such as Canny's algorithm.}
+\label{millie}
+\end{figure}
+
 The mathematical equation for the convolution of a kernel $w$ on an image $f(x,y)$ (represented as a discrete function in two variables) is as follows:
@@
 w \star f(x,y) = \sum^a_{s=-a} \sum^b_{t=-b}f(x+s, y+t)\footnote{In reality, we should have $x-s$ and $y-t$, as correlation involves rotating the kernel by 180 degrees.
-However, the term convolution is often used for both the mathematical operations of correlation and convolution\cite[160]{dip}}
+However, the term convolution is often used for both the mathematical operations of correlation and convolution\cite[160]{dip}.
+Furthermore, the kernels used in CNNs have randomly initialized weights, making rotation inconsequential.}
 \end{equation}
@@
 Given that we have achieved excellent results by using a CNN, we may don't expect to achieve much better results by sending the extracted features to a random forest.
-Regardless, our testing found that by using the same CNN model we found previously,
+Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of $97.8\%$.
+This is lower than what we have achieved by simply using the features in a final dense layer, and is a somewhat disappointing result.
 
@@
 This is very apparent in our testing: When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds.
-This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time.
+This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time\footnote{Of course, you would ideally train your network on better hardware, even if the hardware making the prediction does not have a GPU.
+However, the same ratio holds when making a single prediction as opposed to training the entire model.
+That is, the PCA-NN model is much faster at making a single prediction than the CNN model}.
 However, when using a GPU with CUDA cores, the story is quite different: Here, the CNN used only $6$ seconds, while the PCA-NN model took $12$ seconds\footnote{The fact that this model took longer than the CNN is most likely because the PCA function we use only runs on the CPU, meaning that only the neural net part is actually running on the GPU}.
@@
 Here, we achieved a reasonable result of $94.7\%$ accuracy on the test set.
-Although this is not as impressive as the previous models, there
+Although this is not as impressive as the previous models, it shows that decent results can be achieved without using neural networks at all.
+However, we clearly see that traditional CNNs have performances which are hard to beat using other forms of dimensionality reduction.
 
@@
 Out of the architectures presented in BOOK \cite{geron}, we decided to try AlexNet, which won the ILSVRC 2012 challenge and is a relatively shallow network having $11$ hidden layers.
-Using this model, we achieved a testing accuracy of $87.8\%$.
-Comparing our result to those obtained by Kermany et Al., we see that our model is slightly more ass.
+Using this model, and using the GregSearch algorithm\footnote{This algorithm involves Greg trying out different values until he finds a very good one} we achieved a testing accuracy of $88.9\%$.
+Comparing our result to those obtained by Kermany et Al.
+($92.8\%$), we see that our model is a few percentage points worse, however when considering that we are not using transfer learning and are completing training on a single laptop, we consider our results to be reasonably good.
 
@@
 
+As with the MNIST dataset, we also found disappointing results when attempting to use a CNN-RF model on this dataset.
+Here, we achieved a test accuracy of $87.5\%$, which is $1.4\%$ lower than simply using the CNN as is to make the prediction.
+
 \subsubsection{Principal Component Analysis followed by a Neural Network}
+
+% Given the complexity of the network needed to achieve good results when using a CNN, we do not expect much when combining PCA, which is a much less sophisticated form of feature extraction than convolutional layers in a CNN, with a traditional neural network.
+Regardless, using a PCA-NN model, we extracted a varying amount of principal components and used these as features in a neural network composed of three dense layers.
+Using hyperparameter tuning, we found that a number of nodes in each layer equal to $1000$, a lambda of $0.001$ and a learning rate of $0.001$ gave the best result.
+However, even when using the best hyperparameters, we only achieved a test accuracy of $80.0\%$, which is substantially lower than Kermany et Al's $92.8\%$ and our own CNN model's $88.9\%$.
+This result shows why convolutional layers have become the de facto standard for image recognition, being able to extract features with much higher precision than simply calculating the principal components of a dataset.
+By the use of gradient descent, CNNs are able to tune the convolutional kernels to have the values which best fit the dataset, and as such they are able to be trained to solve many different problems.
+PCA, on the other hand, cannot be tuned or trained to better extract the relevant features, the only parameter than be changed is the amount of principal components to keep.
+With these poor results, no justification based on computational savings can be made, especially when CNNs can be trained extremely quickly on modern GPUs.
+
+Given the complexity of the network needed to achieve good results when using a CNN, one might not expect to find good results using PCA, which is a much more simplistic form of feature extraction.
+However, we have found that we can actually achieve very comparable results to AlexNet using a much simpler model.
+In our PCA-NN model, we tried different number of principal components to keep, and sent these components into a neural network with three hidden layers.
+Tuning the model, we found that optimal hyperparameters were as described by table~.\ref{pcanntable}.
+Using these parameters, we were able to achieve a test accuracy of $87.3\%$.
+This is an impressive accuracy considering the simplicity of our model, and shows that it is possible to achieve reasonable results without turning to complex convolutional neural networks.
+Looking at figure~.\ref{pneupcacomp}, we can see how much information is left when using only $9$ principal components.
+From this we can see that while much detail has been lost, something something
+
+\begin{table}[H]
+\caption{Optimal hyperparameters for the PCA-NN model}
+\begin{center}
+\label{pcanntable}
+\begin{tabular}{| c | l |}
+\hline
+Principal Components & Nodes per hidden layer & Learning Rate & Lambda \\
+\hline
+9 & 17 & 0.001 & 0.001 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
+Yet again impressed with the performance of using a simpler model, we attempt to perform the same classification without using neural networks at all.
+Using a PCA-RF model, we achieved a test accuracy of $83.7\%$.
+NEED TIME STATS FOR THE REST
+
+$87.3\%$ for PCA NN med 17
+$83.7\%$ for PCA RF med 17
+
 \section{Conclusion}
 
-% \begin{table}[h]
-% \caption{Accuracy by }
-% \begin{center}
-% \label{datasettable}
-% \begin{tabular}{| c | l | l | l |}
-% \hline
-% Dataset & Number of instances & Features & Type \\
-% \hline
-% Franke Function (raw data) & 400 & 2 & regression \\
-% Franke Function (polynomial features) & 400 & 56 & regression \\
-% Wisconsin Breast Cancer & 569 & 30 & binary classification \\
-% MNIST 8x8 & 1797 & 64 & multiclass classification (10 classes) \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
-
-I have realized that I am cooler than Eric.
-
-The children of Guangzhou province were saved to a certain extent. They all lived happily ever after.
+No child was saved. More children actually died due to our actions.
 

### 2022-12-08 src/conv_nn_catdog.py
Remove catdog

@@ -1,73 +1,0 @@
-import tensorflow as tf
-import os
-from tensorflow.keras import layers
-from tensorflow.keras import regularizers
-import tensorflow_datasets as tfds
-filedir = os.path.dirname(__file__)
-
-tf.keras.utils.set_random_seed(1336)
-
-TRAINDIR = filedir + "/../data/chest_xray/train"
-TESTDIR = filedir + "/../data/chest_xray/test"
-BATCHSIZE = 256
-IMG_HEIGHT = 227
-IMG_WIDTH = 227
-
-
-def preprocess(img, label):
-    return tf.image.resize(img, [IMG_HEIGHT, IMG_WIDTH]) / 255, label
-
-HEIGHT = 200
-WIDTH = 200
-split = ['train[:70%]', 'train[70%:]']
-
-trainDataset, testDataset = tfds.load(name='cats_vs_dogs', split=split, as_supervised=True)
-
-trainDataset = trainDataset.map(preprocess).batch(BATCHSIZE)
-testDataset = testDataset.map(preprocess).batch(BATCHSIZE)
-
-model = tf.keras.Sequential(
-    [
-        #tf.keras.layers.Rescaling(1.0 / 255),
-        tf.keras.layers.Conv2D(
-            96, 11, strides=(4,4), padding='valid', activation="relu"), #kernel_regularizer=kernel, bias_regularizer=bias),
-
-        tf.keras.layers.MaxPooling2D(pool_size=(3,3), strides=2, padding='valid'),
-        tf.keras.layers.Conv2D(
-            256, 5, strides=(1,1), padding='same', activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias), 
-          
-        tf.keras.layers.MaxPooling2D(pool_size=(3,3), strides=2, padding='valid'), # S4
-        tf.keras.layers.Conv2D(
-            384, 3, strides=(1,1), padding='same', activation="relu"), #kernel_regularizer=kernel, bias_regularizer=bias), #C5
-        tf.keras.layers.Conv2D(
-            384, 3, strides=(1,1), padding='same', activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias), #C6
-        tf.keras.layers.Conv2D(
-            256, 3, strides=(1,1), padding='same', activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias), 
-        
[…]

### 2022-12-08 src/convolutionexample.py
Add convolutionexample

@@ -1,0 +1,19 @@
+# imports
+import numpy as np
+from imageio import imread, imsave
+from scipy import signal
+import matplotlib.pyplot as plt
+
+img = imread("../data/millie2.jpg", as_gray=True)
+kernel = np.array([[-1, -1, -1], [-1, 8, -1], [-1, -1, -1]])
+img_sharp = signal.convolve2d(img, kernel, mode="same")
+
+plt.subplot(121)
+plt.title("Original Image", size=25)
+plt.imshow(img, cmap="gray")
+
+plt.subplot(122)
+plt.title("Image after convolution with a laplacian kernel", size=25)
+plt.imshow(img_sharp, cmap="gray", vmin=10, vmax=14)
+
+plt.show()

### 2022-12-08 src/conv_nn_mnist.py
Update this file to be about finished

@@ -12,11 +12,13 @@ import seaborn as sns
 
 tf.keras.utils.set_random_seed(1336)
 """
-Convolutional neural network used for MNIST dataset. Builds and fits data, takes time.
+Convolutional neural network used for MNIST dataset. Performs hyperparameter tuning,
+trains the model, prints average time for making a prediction on the entire test set
+creates a confusion matrix
 """
 
 batch_size = 128
-epochs = 6
+epochs = 1
 mnist = tf.keras.datasets.mnist
 
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
@@ -24,26 +26,14 @@ mnist = tf.keras.datasets.mnist
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
 
-# print(x_train.shape)
-#
-# train_ds = tf.data.Dataset.from_tensor_slices((x_train, y_train))
-# val_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
-# print(train_ds)
-#
-# train_ds = train_ds.batch(batch_size)
-# val_ds = val_ds.batch(batch_size)
 
-
+# Hyperparameter tuning model
 def model_builder(hp):
     hp_width = hp.Choice("width", values=[3, 5, 7])
 
     model = tf.keras.Sequential(
         [
             layers.Rescaling(1.0 / 255),
-            # layers.Conv2D(32, hp_width, activation="relu", padding="same"),
-            # layers.MaxPooling2D(),
-            # layers.Conv2D(32, hp_width, activation="relu", padding="same"),
-            # layers.MaxPooling2D(),
             layers.Conv2D(32, hp_width, activation="relu", padding="same"),
             layers.MaxPooling2D(),
             layers.Flatten(),
@@ -71,22 +61,18 @@ tuner = kt.Hyperband(
     directory="conv_nn_mnist-params",
 )
 
+# finding the best hyperparameters
 tuner.search(x_train, y_train, epochs=10, validation_split=0.2)
 best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
 width = best_hps.get("width")
 learning_rate = best_hps.get("learning_rate")
-print(width)
-print(learning_rate)
+print(f"Best kernel width found by tuner: {width}")
+print(f"Best learning rate found by tuner: {learning_rate}")
 
-
-start = time.time()
+# using the best hyperparameters
[…]

### 2022-12-08 src/conv_nn_mnist.py
Update timeit

@@ -109,8 +109,10 @@ model.fit(
 # loading the best weights from the model
 model.load_weights(checkpoint_filepath)
 
+model.evaluate(x_test, y_test)
+
 print("Timing model:")
-timeit(model.evaluate, x_test, y_test)
+timeit(model.predict, (x_test, y_test), batch_size=batch_size)
 
 predictions = tf.math.argmax(model.predict(x_test), axis=1)
 conf = conf_mat(predictions, y_test, num_cls=10)

### 2022-12-08 src/utils.py

@@ -3,10 +3,10 @@ from tensorflow.keras import layers
 import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
-from matplotlib.colors import LogNorm 
+from matplotlib.colors import LogNorm
+from typing import Callable
+import time
 
-def PCA_stoch(X, n_components):
-    print("eric")
 
 def plot_confusion(confusion_matrix: np.ndarray, title=None):
     fontsize = 40
@@ -18,7 +18,6 @@ def plot_confusion(confusion_matrix: np.ndarray, title=None):
         fmt=".2%",
         cmap="Blues",
         norm=LogNorm(),
-
     )
     if title:
         plt.title(title)
@@ -30,98 +29,36 @@ def plot_confusion(confusion_matrix: np.ndarray, title=None):
     plt.show()
 
 
-def PCA_stoch(X, n_components, iterations=10, eta=1e-5, convergence=1e-5, epochs=10):
-    print(f"{X.shape=}")
-    m, n = X.shape
-    W_prev_epoch = np.eye(m, n_components)
+def timeit(func: Callable, *args, **kwargs):
+    """Function to time another function
 
-    for e in range(epochs):
-        print(e)
+    Parameters:
+        func: function to be timed
+        args: positional arguments for function
+        kwargs: keyword arguments for function
 
-        W_prev_iter = W_prev_epoch
-        u_mark = np.zeros((m, n_components))
-        for i in range(n):
-            u_mark += np.outer(X[:, i], (X[:, i].T @ W_prev_epoch)) / n
+    Returns:
+        average number of seconds taken for the function to run
+    """
+    n = 5
+    avg = 0
+    for i in range(n):
+        start = time.time()
+        func(*args)
+        avg += (time.time() - start) / n
 
-        for t in range(iterations):
-            print(t)
-            i = np.random.randint(n)
-
-            X_i = X[:, i].reshape(X[:, i].shape[0], 1)
-            W_mark = W_prev_iter + eta * (
-                X_i * (X_i.T @ W_prev_iter - X_i.T @ W_prev_epoch) + u_mark
-            )
-            W_prev_iter, _ = np.linalg.qr(W_mark)
-
-        W_prev_epoch = W_prev_iter
-
-    return W_prev_epoch
+    print(f"Average time taken for {n} calls: {avg}")
+    return avg
 
 
-def PCA_fit(X, n_components):
[…]

### 2022-12-08 src/conv_rf_mnist.py
Add timing

@@ -30,6 +30,7 @@ test_ds = tf.data.Dataset.from_tensor_slices((x_test, y_test))
 train_ds = train_ds.batch(batch_size)
 test_ds = test_ds.batch(batch_size)
 
+# these were found from conv_nn_mnist.py
 width = 5
 learning_rate = 0.001
 
@@ -74,7 +75,7 @@ model.evaluate(test_ds, batch_size=BATCHSIZE)
 
 feature_extractor = tf.keras.Model(
     inputs=model.inputs,
-    outputs=model.layers[-3].output,
+    outputs=model.layers[-2].output,
 )
 
 features_train = train_ds.map(lambda batch, label: (feature_extractor(batch), label))
@@ -92,3 +93,14 @@ forest.compile(metrics=["accuracy"])
 
 print(forest.evaluate(features_train, return_dict=True))
 print(forest.evaluate(features_test, return_dict=True))
+
+# timing:
+
+
+def predict():
+    features_test = test_ds.map(lambda batch, label: (feature_extractor(batch), label))
+    forest.predict(features_test)
+
+
+print("Timing prediction")
+timeit(predict)

### 2022-12-08 src/pca_nn_mnist.py
Add timeit

@@ -116,3 +116,12 @@ plot_confusion(conf, title="Confusion matrix - PCA_NN - MNIST")
 print(results)
 
 print(results["accuracy"])
+
+
+def predit():
+    x_test_pca = pca.transform(x_test_flat)
+    model.predict(x_test_pca, y_test)
+
+
+print("Timing prediction")
+timeit(predict)

### 2022-12-08 src/pca_rf_mnist.py
Add timeit

@@ -26,8 +26,6 @@ n_components = 50
 x_train_flat = tf.reshape(x_train, shape=[-1, 784])
 x_test_flat = tf.reshape(x_test, shape=[-1, 784])
 
-start = time.time()
-
 pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
 pca.fit(x_train_flat)
 
@@ -35,10 +33,10 @@ x_train_pca = pca.transform(x_train_flat)
 x_test_pca = pca.transform(x_test_flat)
 
 train_ds = tf.data.Dataset.from_tensor_slices((x_train_pca, y_train))
-val_ds = tf.data.Dataset.from_tensor_slices((x_test_pca, y_test))
+test_ds = tf.data.Dataset.from_tensor_slices((x_test_pca, y_test))
 
 train_ds = train_ds.batch(batch_size)
-val_ds = val_ds.batch(batch_size)
+test_ds = test_ds.batch(batch_size)
 
 
 forest = tfdf.keras.RandomForestModel(
@@ -52,4 +50,13 @@ forest.fit(x=train_ds)
 forest.compile(metrics=["accuracy"])
 
 print(forest.evaluate(train_ds, return_dict=True))
-print(forest.evaluate(val_ds, return_dict=True))
+print(forest.evaluate(test_ds, return_dict=True))
+
+
+def predict():
+    x_test_pca = pca.transform(x_test_flat)
+    forest.predict((x_test_pca, y_test))
+
+
+print("Timing prediction")
+timeit(predict)

### 2022-12-08 src/conv_nn_better.py
Made it so that conv_nn_better is the model we use

@@ -1,235 +1,0 @@
-import tensorflow as tf
-import os
-from tensorflow.keras import layers
-from tensorflow.keras import regularizers
-import tensorflow_datasets as tfds
-import keras_tuner as kt
-
-filedir = os.path.dirname(__file__)
-from functools import partial
-
-seed = 1336
-tf.keras.utils.set_random_seed(seed)
-tf.config.experimental.enable_op_determinism()
-
-TRAINDIR = filedir + "/../data/chest_xray/train"
-TESTDIR = filedir + "/../data/chest_xray/test"
-BATCHSIZE = 256
-IMG_HEIGHT = 227
-IMG_WIDTH = 227
-
-train_ds = tf.keras.utils.image_dataset_from_directory(
-    TRAINDIR,
-    validation_split=0.1,
-    subset="training",
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=BATCHSIZE,
-    color_mode="grayscale",
-)
-val_ds = tf.keras.utils.image_dataset_from_directory(
-    TRAINDIR,
-    validation_split=0.1,
-    subset="validation",
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=BATCHSIZE,
-    color_mode="grayscale",
-)
-
-test_ds = tf.keras.utils.image_dataset_from_directory(
-    TESTDIR,
-    labels="inferred",
-    seed=1337,
-    image_size=(IMG_HEIGHT, IMG_WIDTH),
-    batch_size=BATCHSIZE,
-    color_mode="grayscale",
-)
-
-COUNT_NORMAL = 1071
-COUNT_PNEUMONIA = 3114
-
-weight_for_0 = (1 / COUNT_NORMAL) * (COUNT_NORMAL + COUNT_PNEUMONIA) / 2.0
-weight_for_1 = (1 / COUNT_PNEUMONIA) * (COUNT_NORMAL + COUNT_PNEUMONIA) / 2.0
-
-class_weight = {0: weight_for_0, 1: weight_for_1}
-
-AUTOTUNE = tf.data.AUTOTUNE
-
-train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
-test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
-
-
-def model_builder(hp):
-    hp_learning_rate = hp.Choice(
[…]

### 2022-12-08 src/conv_nn_pneu.py

@@ -3,12 +3,11 @@ import os
 from tensorflow.keras import layers
 from tensorflow.keras import regularizers
 import tensorflow_datasets as tfds
-filedir = os.path.dirname(__file__)
-import numpy as np
+import keras_tuner as kt
 
-"""
-Convolutional neural network used for pneumonia dataset. Builds and fits data.
-"""
+filedir = os.path.dirname(__file__)
+from functools import partial
+
 seed = 1336
 tf.keras.utils.set_random_seed(seed)
 tf.config.experimental.enable_op_determinism()
@@ -21,6 +20,18 @@ IMG_WIDTH = 227
 
 train_ds = tf.keras.utils.image_dataset_from_directory(
     TRAINDIR,
+    validation_split=0.1,
+    subset="training",
+    labels="inferred",
+    seed=1337,
+    image_size=(IMG_HEIGHT, IMG_WIDTH),
+    batch_size=BATCHSIZE,
+    color_mode="grayscale",
+)
+val_ds = tf.keras.utils.image_dataset_from_directory(
+    TRAINDIR,
+    validation_split=0.1,
+    subset="validation",
     labels="inferred",
     seed=1337,
     image_size=(IMG_HEIGHT, IMG_WIDTH),
@@ -50,51 +61,175 @@ AUTOTUNE = tf.data.AUTOTUNE
 train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
 test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
 
-kernel = tf.keras.regularizers.L2(l2=0.001)
-bias = tf.keras.regularizers.L2(l2=0.001)
+
+def model_builder(hp):
+    hp_learning_rate = hp.Choice(
+        "learning_rate", values=[2 * 1e-5, 2 * 1e-4, 2 * 1e-3, 2 * 1e-2, 2 * 1e-1]
+    )
+
+    model = tf.keras.Sequential(
+        [
+            tf.keras.layers.Rescaling(1.0 / 255),
+            tf.keras.layers.Conv2D(
+                96,
+                11,
+                strides=(4, 4),
+                padding="valid",
+                activation="relu",
+            ),
+            tf.keras.layers.MaxPooling2D(pool_size=(3, 3), strides=2, padding="valid"),
+            tf.keras.layers.Conv2D(
+                256,
+                5,
+                strides=(1, 1),
+                padding="same",
+                activation="relu",
+            ),
[…]

### 2022-12-08 src/conv_nn_pneu.py
Add timeit

@@ -2,8 +2,8 @@ import tensorflow as tf
 import os
 from tensorflow.keras import layers
 from tensorflow.keras import regularizers
-import tensorflow_datasets as tfds
 import keras_tuner as kt
+from utils import *
 
 filedir = os.path.dirname(__file__)
 from functools import partial
@@ -142,9 +142,9 @@ tuner = kt.Hyperband(
     directory="conv_nn_pneu-params",
 )
 
-tuner.search(train_ds, epochs=10, validation_data=val_ds, class_weight=class_weight)
-best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
-learning_rate = best_hps.get("learning_rate")
+# tuner.search(train_ds, epochs=10, validation_data=val_ds, class_weight=class_weight)
+# best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
+learning_rate = 0.001  # best_hps.get("learning_rate")
 
 print(learning_rate)
 
@@ -225,7 +225,7 @@ model.fit(
     train_ds,
     batch_size=BATCHSIZE,
     validation_data=test_ds,
-    epochs=15,
+    epochs=1,
     class_weight=class_weight,
     callbacks=[model_checkpoint_callback],
 )
@@ -233,3 +233,6 @@ model.fit(
 model.load_weights(checkpoint_filepath)
 
 model.evaluate(test_ds, batch_size=BATCHSIZE)
+
+print("Timing model:")
+timeit(model.predict, test_ds, batch_size=BATCHSIZE)

### 2022-12-08 src/conv_nn_pneu.py
Add confusion matrix

@@ -225,7 +225,7 @@ model.fit(
     train_ds,
     batch_size=BATCHSIZE,
     validation_data=test_ds,
-    epochs=1,
+    epochs=2,
     class_weight=class_weight,
     callbacks=[model_checkpoint_callback],
 )
@@ -236,3 +236,20 @@ model.evaluate(test_ds, batch_size=BATCHSIZE)
 
 print("Timing model:")
 timeit(model.predict, test_ds, batch_size=BATCHSIZE)
+
+# making confusion matrix
+y_test = None
+for _, labels in test_ds:
+    if y_test is None:
+        y_test = labels
+
+    else:
+        y_test = tf.concat([y_test, labels], axis=0)
+
+pred = model.predict(test_ds)
+pred = tf.where(pred > 0.5, 1, 0)
+pred = tf.squeeze(pred)
+conf = conf_mat(pred, y_test, 2)
+conf = perc(conf)
+
+plot_confusion(conf, title="Confusion matrix - CNN - MNIST")

### 2022-12-08 src/conv_nn_pneu.py
Fix epochs

@@ -225,7 +225,7 @@ model.fit(
     train_ds,
     batch_size=BATCHSIZE,
     validation_data=test_ds,
-    epochs=2,
+    epochs=15,
     class_weight=class_weight,
     callbacks=[model_checkpoint_callback],
 )

### 2022-12-08 src/conv_nn_pneu.py
Fix tf.where

@@ -247,7 +247,7 @@ for _, labels in test_ds:
         y_test = tf.concat([y_test, labels], axis=0)
 
 pred = model.predict(test_ds)
-pred = tf.where(pred > 0.5, 1, 0)
+pred = tf.where(pred > 0, 1, 0)
 pred = tf.squeeze(pred)
 conf = conf_mat(pred, y_test, 2)
 conf = perc(conf)

### 2022-12-08 src/conv_rf_mnist.py
Fix batchsize issue

@@ -71,7 +71,7 @@ model.fit(
 
 model.load_weights(checkpoint_filepath)
 
-model.evaluate(test_ds, batch_size=BATCHSIZE)
+model.evaluate(test_ds, batch_size=batch_size)
 
 feature_extractor = tf.keras.Model(
     inputs=model.inputs,

### 2022-12-08 src/conv_rf_pneu.py
Add timeit

@@ -255,3 +255,12 @@ forest.compile(metrics=["accuracy"])
 
 print(forest.evaluate(features_train, return_dict=True))
 print(forest.evaluate(features_test, return_dict=True))
+
+
+def predict():
+    features_test = test_ds.map(lambda batch, label: (feature_extractor(batch), label))
+    forest.predict(features_test)
+
+
+print("Timing prediction")
+timeit(predict)

### 2022-12-08 src/pca_nn_mnist.py
More timeit stuff

@@ -62,14 +62,16 @@ tuner = kt.Hyperband(
     directory="pca_nn_mnist-parameters",
 )
 
-tuner.search(x_train_pca, y_train, epochs=10, validation_split=0.2)
-best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
-units = best_hps.get("units")
-learning_rate = best_hps.get("learning_rate")
+# tuner.search(x_train_pca, y_train, epochs=10, validation_split=0.2)
+# best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
+# units = best_hps.get("units")
+# learning_rate = best_hps.get("learning_rate")
 
-print(units)
-print(learning_rate)
+# print(units)
+# print(learning_rate)
 
+units = 100
+learning_rate = 0.001
 
 start = time.time()
 pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
@@ -118,9 +120,9 @@ print(results)
 print(results["accuracy"])
 
 
-def predit():
+def predict():
     x_test_pca = pca.transform(x_test_flat)
-    model.predict(x_test_pca, y_test)
+    model.predict((x_test_pca, y_test))
 
 
 print("Timing prediction")

### 2022-12-08 src/pca_nn_pneu.py

@@ -212,36 +212,40 @@ with tf.device("/gpu:0"):
         y_train,
         validation_data=(X_test_pca, y_test),
         epochs=15,
-        batch_size=BATCHSIZE, #64,
+        batch_size=BATCHSIZE,  # 64,
         class_weight=class_weight,
         callbacks=[model_checkpoint_callback],
     )
     model.load_weights(checkpoint_filepath)
 
     model.evaluate(X_test_pca, y_test, batch_size=BATCHSIZE)
-    print(f"Time taken: {time.time() - start}")
+
+    def predict():
+        x_test_pca = pca.transform(X_test)
+        model.predict((x_test_pca, y_test))
+
+    print("Timing prediction")
+    timeit(predict)
 
     X_train_pca_restored = pca.inverse_transform(X_test_pca)
 
     plt.subplot(321)
-    plt.title(f"Actual instance 51k features. Class {y_train[0]}", size=18)
+    plt.title(f"Actual instance, 51k features. Class {y_train[0]}", size=18)
     plt.imshow(tf.reshape(X_train[0], (227, 227)))
     plt.subplot(322)
     plt.title("PCA with 9 components", size=18)
     plt.imshow(tf.reshape(X_train_pca_restored[0], (227, 227)))
 
     plt.subplot(323)
-    plt.title(f"Actual instance 51k features. Class {y_train[5]}", size=18)
+    plt.title(f"Class {y_train[5]}", size=18)
     plt.imshow(tf.reshape(X_train[5], (227, 227)))
     plt.subplot(324)
-    plt.title("PCA with 9 components", size=18)
     plt.imshow(tf.reshape(X_train_pca_restored[5], (227, 227)))
 
     plt.subplot(325)
-    plt.title(f"Actual instance 51k features. Class {y_train[8]}", size=18)
+    plt.title(f"Class {y_train[8]}", size=18)
     plt.imshow(tf.reshape(X_train[8], (227, 227)))
     plt.subplot(326)
-    plt.title("PCA with 9 components", size=18)
     plt.imshow(tf.reshape(X_train_pca_restored[8], (227, 227)))
 
     plt.show()

### 2022-12-08 src/pca_rf_mnist.py
More timeit

@@ -55,7 +55,9 @@ print(forest.evaluate(test_ds, return_dict=True))
 
 def predict():
     x_test_pca = pca.transform(x_test_flat)
-    forest.predict((x_test_pca, y_test))
+    test_ds = tf.data.Dataset.from_tensor_slices((x_test_pca, y_test))
+    test_ds = test_ds.batch(batch_size)
+    forest.predict(test_ds)
 
 
 print("Timing prediction")

### 2022-12-08 src/pca_rf_pneu.py

@@ -62,7 +62,7 @@ def model_builder(hp):
 with tf.device("/cpu:0"):
     TRAINDIR = filedir + "/../data/chest_xray/train"
     TESTDIR = filedir + "/../data/chest_xray/test"
-    BATCHSIZE = 128
+    batch_size = 128
     IMG_HEIGHT = 200
     IMG_WIDTH = 200
 
@@ -71,7 +71,7 @@ with tf.device("/cpu:0"):
         labels="inferred",
         seed=1337,
         image_size=(IMG_HEIGHT, IMG_WIDTH),
-        batch_size=BATCHSIZE,
+        batch_size=batch_size,
         color_mode="grayscale",
     )
 
@@ -80,7 +80,7 @@ with tf.device("/cpu:0"):
         labels="inferred",
         seed=1337,
         image_size=(IMG_HEIGHT, IMG_WIDTH),
-        batch_size=BATCHSIZE,
+        batch_size=batch_size,
         color_mode="grayscale",
     )
 
@@ -134,15 +134,24 @@ with tf.device("/cpu:0"):
     start = time.time()
     pca.fit(X_train)
 
-    X_train = pca.transform(X_train)
-    X_test = pca.transform(X_test)
+    X_train_pca = pca.transform(X_train)
+    X_test_pca = pca.transform(X_test)
 
     forest = tfdf.keras.RandomForestModel(
         verbose=1, max_depth=40, random_seed=1337, check_dataset=False
     )
-    forest.fit(X_train, y_train, class_weight=class_weight)
+    forest.fit(X_train_pca, y_train, class_weight=class_weight)
 
     forest.compile(metrics=["accuracy"])
 
-    print(forest.evaluate(X_train, y_train, return_dict=True))
-    print(forest.evaluate(X_test, y_test, return_dict=True))
+    print(forest.evaluate(X_train_pca, y_train, return_dict=True))
+    print(forest.evaluate(X_test_pca, y_test, return_dict=True))
+
+    def predict():
+        X_test_pca = pca.transform(X_test)
+        test_ds = tf.data.Dataset.from_tensor_slices((X_test_pca, y_test))
+        test_ds = test_ds.batch(batch_size)
+        forest.predict(test_ds)
+
+    print("Timing prediction")
+    timeit(predict)

### 2022-12-08 doc/paper.tex
Update paper
Write abstract and add tables

@@
 It was computing many computinations but due to severe liminations it had problems with binaly classifination and a large emesee.
-Kurwa.
+
+We compared convolutional neural networks, the de facto standard for image classification problems, with hybrid models created by combining neural networks with Principal Component Analysis (PCA) and Random Forests (RF).
+By doing this, we explored the tradeoff between prediction accuracy and computational cost for four different ML models: A pure CNN model, a CNN-RF model, a PCA-NN model and a PCA-RF model.
+We used these models on the well known MNIST dataset and the Guangzhou Pneumonia Dataset consisting of x-ray images of potential cases of pneumonia.
+From this, we found that pure CNNs performed best in terms of accuracy, achieving test accuracies of $98.8\%$ and $88.8\%$ on the MNIST and pneumonia dataset, respectively.
+However, we found that the PCA-NN model performed similarly, at a much lower computational cost MAYBE.
 
@@
 First we will use a traditional CNN, where a series convolutional layers is followed by a series fully connected layers.
-Then we will modify our CNN to create a CNN-RF model, where the features extracted by the convolutional layers are fed into a Random Forest instead of a series of fully connected layers.
-Finally, we will make use of Principal Component Analysis instead of convolution for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest.
+Then we will modify our CNN to create a hybrid CNN-RF model, where the features extracted by the CNN are fed into a Random Forest which makes the final prediction.
+Finally, we will make use of Principal Component Analysis instead of convolution and pooling for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest.
 In doing this, we hope to answer the following question: what is the best way to tackle the challenges associated with image classification?
@@
 Given that we have achieved excellent results by using a CNN, we may don't expect to achieve much better results by sending the extracted features to a random forest.
-Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of $97.8\%$.
-This is lower than what we have achieved by simply using the features in a final dense layer, and is a somewhat disappointing result.
+Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of
+
+% $97.8\%$. This is lower than what we have achieved by simply using the features in a final dense layer, and is a somewhat disappointing result.
 
@@
 When considering the computational effort, we need to consider the performance benefit that modern GPU's bring, being specifically designed to run neural networks efficiently.
-This is very apparent in our testing: When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds.
+This is very apparent in our testing:
+
+% When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds.
 This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time\footnote{Of course, you would ideally train your network on better hardware, even if the hardware making the prediction does not have a GPU.
@@
 However, we clearly see that traditional CNNs have performances which are hard to beat using other forms of dimensionality reduction.
+
+\subsubsection{MNIST recap}
+
+Below is table showing the accuracy and average prediction time for each model on the MNIST dataset.
+
+\begin{table}[H]
+\caption{Table showing the accuracy and average time used to make a prediction on the entire test dataset, for each model}
+\begin{center}
+\label{pcanntable}
+\begin{tabular}{| c | l | l |}
+\hline
+Model & Test accuracy & Average prediction time \\
+\hline
+CNN & 98.8\% & 123 \\
+CNN-RF & 98.8\% & 123 \\
+PCA-NN & 97.8\% & 123 \\
+PCA-RF & 94.7\% & 123 \\
+CNN (CPU only) & 98.8\% & 1000 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+From this we can see that bla bla bla
 
@@
 
-% Given the complexity of the network needed to achieve good results when using a CNN, we do not expect much when combining PCA, which is a much less sophisticated form of feature extraction than convolutional layers in a CNN, with a traditional neural network.
-Regardless, using a PCA-NN model, we extracted a varying amount of principal components and used these as features in a neural network composed of three dense layers.
-Using hyperparameter tuning, we found that a number of nodes in each layer equal to $1000$, a lambda of $0.001$ and a learning rate of $0.001$ gave the best result.
-However, even when using the best hyperparameters, we only achieved a test accuracy of $80.0\%$, which is substantially lower than Kermany et Al's $92.8\%$ and our own CNN model's $88.9\%$.
-This result shows why convolutional layers have become the de facto standard for image recognition, being able to extract features with much higher precision than simply calculating the principal components of a dataset.
-By the use of gradient descent, CNNs are able to tune the convolutional kernels to have the values which best fit the dataset, and as such they are able to be trained to solve many different problems.
-PCA, on the other hand, cannot be tuned or trained to better extract the relevant features, the only parameter than be changed is the amount of principal components to keep.
-With these poor results, no justification based on computational savings can be made, especially when CNNs can be trained extremely quickly on modern GPUs.
-
 Given the complexity of the network needed to achieve good results when using a CNN, one might not expect to find good results using PCA, which is a much more simplistic form of feature extraction.
@@
 \label{pcanntable}
-\begin{tabular}{| c | l |}
+\begin{tabular}{| l | l | l | l |}
 \hline
@@
 \end{table}
+
+% \begin{figure}[H]
+% \centerline{\includegraphics[width=8in]{figure/pneupcacomp.png}}
+% \caption{Figure showing an image from the Guangzhou dataset, compared to the PCA reconstruction of the same image using only 9 principal components}
+% \label{pneupcacomp}
+% \end{figure}
 
@@
 
+\subsubsection{Pneumonia recap}
+
+Below is table showing the accuracy and average prediction time for each model on the Guangzhou Pneumonia Dataset.
+
+\begin{table}[H]
+\caption{Table showing the accuracy and average time used to make a prediction on the entire test dataset, for each model}
+\begin{center}
+\label{pcanntable}
+\begin{tabular}{| c | l | l |}
+\hline
+Model & Test accuracy & Average prediction time \\
+\hline
+CNN & 98.8\% & 123 \\
+CNN-RF & 98.8\% & 123 \\
+PCA-NN & 97.8\% & 123 \\
+PCA-RF & 94.7\% & 123 \\
+CNN (CPU only) & 98.8\% & 1000 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \section{Conclusion}
 
-No child was saved. More children actually died due to our actions.
+No child was saved. More children actually died due to our actions. Nevermind childern actually were saved.
 

### 2022-12-09 src/conv_rf_pneu.py
Add tuner maybe

@@ -245,10 +245,12 @@ features_test = val_ds.map(lambda batch, label: (feature_extractor(batch), label
 print(features_test)
 print(features_train)
 
+tuner = tfdf.tuner.RandomSearch(num_trials=20)
+tuner.choice("max_depth", [5, 10, 15, 20, 25, 30])
+tuner.choice("min_examples", [5, 7, 9, 11, 13])
 
-forest = tfdf.keras.RandomForestModel(
-    verbose=1, max_depth=16, random_seed=1337, check_dataset=False
-)
+
+forest = tfdf.keras.RandomForestModel(tuner=tuner, check_dataset=False)
 forest.fit(x=features_train)
 
 forest.compile(metrics=["accuracy"])

### 2022-12-09 src/conv_rf_pneu.py
Fix stuff

@@ -5,6 +5,7 @@ from tensorflow.keras import regularizers
 import tensorflow_datasets as tfds
 import keras_tuner as kt
 import tensorflow_decision_forests as tfdf
+from utils import *
 
 filedir = os.path.dirname(__file__)
 from functools import partial
@@ -241,7 +242,7 @@ feature_extractor = tf.keras.Model(
 )
 
 features_train = train_ds.map(lambda batch, label: (feature_extractor(batch), label))
-features_test = val_ds.map(lambda batch, label: (feature_extractor(batch), label))
+features_test = test_ds.map(lambda batch, label: (feature_extractor(batch), label))
 print(features_test)
 print(features_train)
 

### 2022-12-09 doc/paper.tex
Update paper

@@
 
-% \author{Eric E. Reber
+\author{Eric E. Reber
+    \affiliation{
+	Bachelor Student, Robotics and \\
+	Intelligent Systems\\ \\[-10pt]
+	Department of Informatics\\ \\[-10pt]
+	The faculty of Mathematics and \\
+	Natural Sciences\\ \\[-10pt]
+    Email: ericer@ifi.uio.no
+    }
+}
+
+% \author{Ed "Gurgler" Bickley
 %     \affiliation{
-% 	Bachelor Student, Robotics and \\
-% 	Intelligent Systems\\ \\[-10pt]
-% 	Department of Informatics\\ \\[-10pt]
-% 	The faculty of Mathematics and \\
-% 	Natural Sciences\\ \\[-10pt]
-%     Email: ericer@ifi.uio.no
+% 	Inmate, Texas State \\
+% 	Penistentiary\\ \\[-10pt]
+% 	Huntsville, Texas\\ \\[-10pt]
+% 	United States of America\\
+%     Email: thegurgler@hotmail.com
 %     }
 % }
-
-\author{Ed "Gurgler" Bickley
-    \affiliation{
-	Inmate, Texas State \\
-	Penistentiary\\ \\[-10pt]
-	Huntsville, Texas\\ \\[-10pt]
-	United States of America\\
-    Email: thegurgler@hotmail.com
-    }
-}
 
@@
 
-% \title{Dimention reduction for strong vs weak learners in pneumonia X-ray image prediction}
 % \title{Liminations of computing compunations using tensorflow implemenation}
@@
 \section{Abstract}
-
-% We found a dog driving a car down Hollywood Boulevard.
-It had sunglasses on, but didn't know where to go due to demensionality reduction.
-It was computing many computinations but due to severe liminations it had problems with binaly classifination and a large emesee.
 
@@
 
-Some might say that Minsky himself used PCA back in the stone age, but we have made numerous improvements to the PCA algorithm:
-
-\begin{itemize}
-  \item - XD random
-  \item - \cite{halko}
-\end{itemize}
-
 The task of image recognition and classification is a great challenge in machine learning, as there are a large number of features present in the data which slow down our training.
 For a relatively small one-channel image of $200$ by $200$ pixels, the naive approach of using every pixel as a feature becomes unfeasable, as we now have $40000$ features.
-Clearly, some form of dimension reduction must be used to rectify this problem, and Convolutional Neural Networks are regarded as the gold standard for this.
+Clearly, some form of dimension reduction must be used to rectify this problem, and Convolutional Neural Networks (CNNs) are regarded as the gold standard for this.
 CNNs are able to reduce the dimensionality while extracting useful information in each layer, for example by finding edges in the first layer, combinations of edges that make a shape in the next layer, and combinations of shapes that resemble an object in the last layer.
 However, CNNs have many hyperparameters that need to be adjusted to achieve optimal performance, and a slow and computationally intensive training process.
+Furthermore, the prediction process itself can be slow, depending on the size of the network and the hardware availiable.
+In small embedded systems without GPUs, or in systems with strict real-time constraints, a CNN may not be the best option.
 In this paper, we explore alternative models for image classification, using Principal Component Analysis and Random Forests in combination with parts of the traditional CNN model.
-We will investigate how these architectures perform for a simple image classification problem such as the MNIST dataset versus a more difficult classification problem in the form of pneumonia prediction from X-ray images.
+We will investigate how these architectures perform for a simple image classification problem such as the MNIST dataset, as well as a more difficult classification problem in the form of pneumonia prediction from X-ray images.
 
@@
 Finally, we will make use of Principal Component Analysis instead of convolution and pooling for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest.
-In doing this, we hope to answer the following question: what is the best way to tackle the challenges associated with image classification?
-
-% More specifically, we will compare a traditional convolutional neural network with an architecture of a convolutional stage fed into a random forest, and against an architecture consisting of a PCA stage fed into either a neural network or a random forest with the aim of decreasing the computational cost or increasing the accuracy accross different datasets.
-Our investigation can be simplified into one question: what is the best way to tackle the challenges associated with image classification?
+In doing this, we hope to answer the following question: could hybrid models be a viable alternative to CNNs, either by directly outperforming them in accuracy or by offering an appealing tradeoff of performance and computational effort?
 
@@
 However, there exists stochastic methods to achieve good approximations, like Randomized PCA \cite[227]{halko}.
-
-% Principal Component Analysis is a dimensionality reduction method performs a change of basis into a lower dimensional subspace of our feature space, while capturing as much of the variance of our dataset as possible.
-To do this, we can zero-center our datThis can be done by performing a Singular Value Decomposition of the design matrix, yielding the following decomposition.
+Furthermore, when a decomposition is found, transforming any arbitrary input is trivial.
 
@@
 
-All neural network stages were implemented using the Tensorflow library.
-PCA was implemented as a custom layer using the Keras API, or by mapping the operation to the whole dataset using Tensorflow's \texttt{dataset.map} functionality, depending on the model.
+All neural network stages were implemented through the Keras API, using the Tensorflow library.
+Randomized PCA was implemented using Scikit-learn's PCA class, which follows the algorithm of Halko et Al..
 Random Forests were implemented using the Tensorflow Decision Forests library.
-
 For hyperparameter tuning, we used the Keras Tuner library \cite{kerastuner}.
@@
 
-We found so much stuff man
+With the theory and implementation explained, we can now discuss the results we achieved by applying these methods.
 
@@
 By gridsearch, we found the optimal filter width to be $7$ and the optimal learning rate to be $0.001$.
-With these hyperparameters we achieved a test accuracy of $98.8\%$, a very good result.
+With these hyperparameters we achieved a test accuracy of $99.0\%$, a very good result.
 Based on this result, we conclude that no more hidden layers or regularization parameters need to be included.
@@
 
-Given that we have achieved excellent results by using a CNN, we may don't expect to achieve much better results by sending the extracted features to a random forest.
-Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of
-
-% $97.8\%$. This is lower than what we have achieved by simply using the features in a final dense layer, and is a somewhat disappointing result.
+Given that we have achieved excellent results by using a CNN, we may not expect to achieve better results by sending the extracted features to a random forest.
+Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of $98.1\%$, which is slightly worse than a pure CNN model.
+Furthermore, this method introduced a significant computational cost to the prediction, as can be seen from table~\ref{modcompmnist}.
+From this we conclude that a CNN-RF hybrid model is inferior to a pure CNN model in this case, achieving both lower accuracies and being more computationally expensive.
 
@@
 
-Replacing the convolutional stage with a simpler form of dimensionality reduction, namely PCA, we can see that we are able to reduce the dimensionality of our dataset while retaining a large amount of the information.
+With somewhat disappointing results from introducing a different model when making the prediction, we instead attempt to perform the feature extraction and dimensionality reduction stage with a different method.
+We therefore attempt to use Principal Component Analysis, and can see that we are able to reduce the dimensionality of our dataset while retaining a large amount of the information.
 Looking at figure~\ref{pcacomp}, we can see that even when using only the $50$ first principal components, the numbers are still readable.
-From this, we can assume that we can reduce the amount of features sent into our neural net from $784$ to less than $100$ features, and still expect reasonable results.
+This implies that we are able to reduce the amount of features we send to the dense layers and still expect reasonable results.
 
@@
 This is comparable to the accuracy obtained by a traditional convolutional neural network.
-When considering the computational effort, we need to consider the performance benefit that modern GPU's bring, being specifically designed to run neural networks efficiently.
-This is very apparent in our testing:
+Furthermore, when looking at the speed at which we are able to make predictions, we see that this model makes predictions around $34\%$ faster than a pure CNN model, while sacrificing less than a single percentage point in accuracy.
+However, when using a GPU, there is very little difference in prediction time between the two models.
+Regardless, this decrease in computational effort could offer an appealing tradeoff for embedded systems which do not have access to dedicated GPUs.
 
@@
 \begin{center}
-\label{pcanntable}
-\begin{tabular}{| c | l | l |}
+\label{modcompmnist}
+\begin{tabular}{| c | l | l | l |}
 \hline
-Model & Test accuracy & Average prediction time \\
+Model & Test accuracy & Average prediction time & Avg. time when GPU accerelated \\
 \hline
-CNN & 98.8\% & 123 \\
-CNN-RF & 98.8\% & 123 \\
-PCA-NN & 97.8\% & 123 \\
-PCA-RF & 94.7\% & 123 \\
-CNN (CPU only) & 98.8\% & 1000 \\
+CNN & 99.0\% & 1.877s & 0.2895s \\
+CNN-RF & 98.9\% & 4.511s & 4.934s \\
+PCA-NN & 98.1\% & 1.244 & 0.3073s \\
+PCA-RF & 94.8\% & 1.182s & - \\
 \hline
@@
 Thus, we turn to the literature, and explore models which have been proven to be effective at image recognition.
-Out of the architectures presented in BOOK \cite{geron}, we decided to try AlexNet, which won the ILSVRC 2012 challenge and is a relatively shallow network having $11$ hidden layers.
-Using this model, and using the GregSearch algorithm\footnote{This algorithm involves Greg trying out different values until he finds a very good one} we achieved a testing accuracy of $88.9\%$.
+Out of the architectures presented in {\it Hands-On Machine Learning with Scikit-Learn and Tensorflow} \cite{geron}, we decided to try AlexNet, which won the ILSVRC 2012 challenge and is a relatively shallow network having $11$ hidden layers.
+Using this model, and using the Adam optimizer, we achieved a testing accuracy of $88.9\%$.
 Comparing our result to those obtained by Kermany et Al.
@@
 As with the MNIST dataset, we also found disappointing results when attempting to use a CNN-RF model on this dataset.
-Here, we achieved a test accuracy of $87.5\%$, which is $1.4\%$ lower than simply using the CNN as is to make the prediction.
+Here, we achieved a test accuracy of $88.1\%$, which is $0.7\%$ lower than simply using the CNN as is to make the prediction.
 
@@
 In our PCA-NN model, we tried different number of principal components to keep, and sent these components into a neural network with three hidden layers.
-Tuning the model, we found that optimal hyperparameters were as described by table~.\ref{pcanntable}.
-Using these parameters, we were able to achieve a test accuracy of $87.3\%$.
+Tuning the model, we found that optimal hyperparameters were as described by table~\ref{pcanntable}.
+Using these parameters, we were able to achieve a test accuracy of $87.8\%$.
 This is an impressive accuracy considering the simplicity of our model, and shows that it is possible to achieve reasonable results without turning to complex convolutional neural networks.
@@
 \hline
-9 & 17 & 0.001 & 0.001 \\
+9 & 9 & 0.1 & 0.0001 \\
 \hline
@@
 
-$87.3\%$ for PCA NN med 17
-$83.7\%$ for PCA RF med 17
-
 \subsubsection{Pneumonia recap}
@@
 \label{pcanntable}
-\begin{tabular}{| c | l | l |}
+\begin{tabular}{| c | l | l | l |}
 \hline
-Model & Test accuracy & Average prediction time \\
+Model & Test accuracy & Average prediction time & Avg. time when GPU accerelated \\
 \hline
-CNN & 98.8\% & 123 \\
-CNN-RF & 98.8\% & 123 \\
-PCA-NN & 97.8\% & 123 \\
-PCA-RF & 94.7\% & 123 \\
-CNN (CPU only) & 98.8\% & 1000 \\
+CNN & 88.9\% & 2.864s & 0.1179s \\
+CNN-RF & 88.1\% & 3.437s & 3.356s \\
+PCA-NN & 87.8\% & 0.3387s & 0.1823s \\
+PCA-RF & 84.9\% & 0.1667s & - \\
 \hline

### 2022-12-09 doc/paper.tex
More on paper, conclusion done

@@
 
-% \author{Ed "Gurgler" Bickley
-%     \affiliation{
-% 	Inmate, Texas State \\
-% 	Penistentiary\\ \\[-10pt]
-% 	Huntsville, Texas\\ \\[-10pt]
-% 	United States of America\\
-%     Email: thegurgler@hotmail.com
-%     }
-% }
-
 \author{Gregor Kajda
@@
 
-% \title{Liminations of computing compunations using tensorflow implemenation}
 \title{Hybrid models as an alternative to traditional Convolutional Neural Networks for image classification}
@@
 We used these models on the well known MNIST dataset and the Guangzhou Pneumonia Dataset consisting of x-ray images of potential cases of pneumonia.
-From this, we found that pure CNNs performed best in terms of accuracy, achieving test accuracies of $98.8\%$ and $88.8\%$ on the MNIST and pneumonia dataset, respectively.
-However, we found that the PCA-NN model performed similarly, at a much lower computational cost MAYBE.
+From this, we found that pure CNNs performed best in terms of accuracy, achieving test accuracies of $99.0\%$ and $88.8\%$ on the MNIST and pneumonia dataset, respectively.
+However, we found that the PCA-NN model performed similarly, at a much lower computational cost, achieving an impressive $88\%$ increase in prediction speed at a cost of only $1.1$ percentage points in accuracy.
 
@@
 The task of image recognition and classification is a great challenge in machine learning, as there are a large number of features present in the data which slow down our training.
-For a relatively small one-channel image of $200$ by $200$ pixels, the naive approach of using every pixel as a feature becomes unfeasable, as we now have $40000$ features.
+For a relatively small one-channel image of $227$ by $227$ pixels, the naive approach of using every pixel as a feature becomes unfeasable, as we now have $51$ thousand features.
 Clearly, some form of dimension reduction must be used to rectify this problem, and Convolutional Neural Networks (CNNs) are regarded as the gold standard for this.
 CNNs are able to reduce the dimensionality while extracting useful information in each layer, for example by finding edges in the first layer, combinations of edges that make a shape in the next layer, and combinations of shapes that resemble an object in the last layer.
-However, CNNs have many hyperparameters that need to be adjusted to achieve optimal performance, and a slow and computationally intensive training process.
-Furthermore, the prediction process itself can be slow, depending on the size of the network and the hardware availiable.
-In small embedded systems without GPUs, or in systems with strict real-time constraints, a CNN may not be the best option.
+However, all these layers of convolution introduce significant computational costs, and CNNs are often reliant on Graphical Processing Units (GPUs) for efficient training and prediction.
+In many embedded systems or mobile devices, the computational cost introduced by using deep CNNs may not be ideal, as the absence of dedicated hardware may lead to long prediction times.
+Furthermore, these systems may have real-time constraints, needing to make many predictions per second.
 In this paper, we explore alternative models for image classification, using Principal Component Analysis and Random Forests in combination with parts of the traditional CNN model.
@@
 First we will use a traditional CNN, where a series convolutional layers is followed by a series fully connected layers.
-Then we will modify our CNN to create a hybrid CNN-RF model, where the features extracted by the CNN are fed into a Random Forest which makes the final prediction.
+We will then modify our CNN to create a hybrid CNN-RF model, where the features extracted by the CNN are fed into a Random Forest, which makes the final prediction.
 Finally, we will make use of Principal Component Analysis instead of convolution and pooling for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest.
-In doing this, we hope to answer the following question: could hybrid models be a viable alternative to CNNs, either by directly outperforming them in accuracy or by offering an appealing tradeoff of performance and computational effort?
+In doing this, we hope to answer the following question: could hybrid models be a viable alternative to CNNs, either by directly outperforming them in accuracy, or by offering an appealing tradeoff between performance and computational effort?
 
@@
 Using different amount of principal components, we sent the reduced dataset into a neural network with three hidden layers.
-After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $97.8\%$.
+After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $98.1\%$.
 This is comparable to the accuracy obtained by a traditional convolutional neural network.
+Looking at figure~\ref{mnistheatmap_pcann}, we again see that we are making accurate predictions across all classes.
 Furthermore, when looking at the speed at which we are able to make predictions, we see that this model makes predictions around $34\%$ faster than a pure CNN model, while sacrificing less than a single percentage point in accuracy.
@@
 Regardless, this decrease in computational effort could offer an appealing tradeoff for embedded systems which do not have access to dedicated GPUs.
-
-% When using only a CPU, the PCA-NN model was much less intensive, requiring only $17$ seconds to complete training, while the convolutional neural network used $164$ seconds.
-This tells us that when computational resources are limited, simpler architectures can give reasonable results in a fraction of the time\footnote{Of course, you would ideally train your network on better hardware, even if the hardware making the prediction does not have a GPU.
-However, the same ratio holds when making a single prediction as opposed to training the entire model.
-That is, the PCA-NN model is much faster at making a single prediction than the CNN model}.
-However, when using a GPU with CUDA cores, the story is quite different: Here, the CNN used only $6$ seconds, while the PCA-NN model took $12$ seconds\footnote{The fact that this model took longer than the CNN is most likely because the PCA function we use only runs on the CPU, meaning that only the neural net part is actually running on the GPU}.
-Looking at the confusion matrix (Fig.~\ref{mnistheatmap_pcann}), we can see that we are again achieving very few misclassifications using the PCA-NN model.
 
@@
 Satisfied with the results achieved by replacing the convolutional stage with Principal Component Analysis, we attempt to sidestep neural networks all together, and use PCA followed by a random forest.
-Here, we achieved a reasonable result of $94.7\%$ accuracy on the test set.
+Here, we achieved a reasonable result of $94.8\%$ accuracy on the test set.
 Although this is not as impressive as the previous models, it shows that decent results can be achieved without using neural networks at all.
-However, we clearly see that traditional CNNs have performances which are hard to beat using other forms of dimensionality reduction.
+Furthermore, we again see a slight performance gain in our prediction, with this model being $38\%$ faster than the pure CNN model at making a prediction for the entire test dataset.
+However, the decrease in accuracy makes this tradeoff less appealing, and because the libraries we use do not offer GPU accelaration when using PCA or random forests, there is no reason to use this model when a GPU is available.
 
-\subsubsection{MNIST recap}
+\subsubsection{MNIST model comparison}
 
-Below is table showing the accuracy and average prediction time for each model on the MNIST dataset.
+With each model optimized and tested, we can compare them to eachother in terms of accuracy and prediction speed.
+Table~\ref{modcompmnist} shows the relevant values for each model.
+From this we can gain several insights.
+Firstly, we can see that we have not been able to outperform a pure CNN model with any of our hybrid models.
+This is not so surprising, as CNNs are regarded as the gold standard for image classification, for a multitude of reasons.
+However, we can also see that we are able to perform decently with all of our models, achieving test accuracies above $98\%$ for all models except the PCA-RF model, which achieved a test accuracy of $94.8\%$, a result that is still reasonable.
+Furthermore, both the PCA models were faster than the pure CNN model (without GPU acceleration), with the PCA-NN model presenting particularly appealing performance/cost-tradeoff.
 
 \begin{table}[H]
-\caption{Table showing the accuracy and average time used to make a prediction on the entire test dataset, for each model}
+\caption{Table showing the accuracy and average time used to make a prediction on the entire MNIST test dataset, for each model.
+All models have been tested on the same system, and the average speed is calculated over 5 predictions.
+Neither PCA nor RF utilize the GPU, and thus it is not possible to accelerate the PCA-RF model.}
 \begin{center}
@@
 CNN & 99.0\% & 1.877s & 0.2895s \\
-CNN-RF & 98.9\% & 4.511s & 4.934s \\
+CNN-RF & 98.9\% & 4.910s & 4.898s \\
 PCA-NN & 98.1\% & 1.244 & 0.3073s \\
@@
 \end{table}
-
-From this we can see that bla bla bla
 
@@
 Thus, we turn to the literature, and explore models which have been proven to be effective at image recognition.
-Out of the architectures presented in {\it Hands-On Machine Learning with Scikit-Learn and Tensorflow} \cite{geron}, we decided to try AlexNet, which won the ILSVRC 2012 challenge and is a relatively shallow network having $11$ hidden layers.
-Using this model, and using the Adam optimizer, we achieved a testing accuracy of $88.9\%$.
+Out of the architectures presented in {\it Hands-On Machine Learning with Scikit-Learn and Tensorflow} \cite{geron}, we decided to try AlexNet, which won the ILSVRC 2012 challenge.
+This choice of model was motivated by a few different reasons.
+Firstly, AlexNet is a relatively shallow network having only $11$ hidden layers.
+Furthermore, although more effective models have been created after 2012, many of these use more complicated architectures such as skip layers or layers which feed their outputs to all subsequent layers.
+AlexNet is both a relatively simple model to train using our available hardware, and a simple model to understand.
+Using AlexNet with the Adam optimizer, we achieved a testing accuracy of $88.9\%$.
 Comparing our result to those obtained by Kermany et Al.
@@
 Here, we achieved a test accuracy of $88.1\%$, which is $0.7\%$ lower than simply using the CNN as is to make the prediction.
+Furthermore, using a random forests increases the computational complexity considerably as well, as can be seen from Table~\ref{modcomppneu}.
 
@@
 
-Given the complexity of the network needed to achieve good results when using a CNN, one might not expect to find good results using PCA, which is a much more simplistic form of feature extraction.
+Again we then attempt to replace the feature extraction stage of a CNN with a simple application of Principal Component Analysis.
+Given the complexity of the network needed to achieve good results when using a CNN, one might not expect to find good results using PCA.
 However, we have found that we can actually achieve very comparable results to AlexNet using a much simpler model.
 In our PCA-NN model, we tried different number of principal components to keep, and sent these components into a neural network with three hidden layers.
-Tuning the model, we found that optimal hyperparameters were as described by table~\ref{pcanntable}.
+Tuning the model, we found that optimal hyperparameters were as described by Table~\ref{pcanntable}.
 Using these parameters, we were able to achieve a test accuracy of $87.8\%$.
 This is an impressive accuracy considering the simplicity of our model, and shows that it is possible to achieve reasonable results without turning to complex convolutional neural networks.
-Looking at figure~.\ref{pneupcacomp}, we can see how much information is left when using only $9$ principal components.
-From this we can see that while much detail has been lost, something something
+Furthermore, this small decrease in accuracy came with a considerable gain in prediction speed when predicting using only the CPU, with the PCA-NN model being $88\%$ faster than the CNN on average (Table.~\ref{modcomppneu}).
+Of course, prediction of pneumonia is not a problem where one would trade accuracy for prediction speed, but there are many real-time use cases for image recognition where this is a tradeoff that is preferable.
 
@@
 
-% \begin{figure}[H]
-% \centerline{\includegraphics[width=8in]{figure/pneupcacomp.png}}
-% \caption{Figure showing an image from the Guangzhou dataset, compared to the PCA reconstruction of the same image using only 9 principal components}
-% \label{pneupcacomp}
-% \end{figure}
+The fact that only $9$ components where needed is quite surprising, but could be explained by the homogeneity of the dataset.
+Unlike the MNIST dataset, where each class is very different, both the positive and negative classes of pneumonia have images which are very similar, as can be seen in figure~\ref{pneupcacomp}.
+It is then likely that the first $9$ principal components capture the variance which in large part decides whether an image displays of a case of pneumonia or not.
+In figure~\ref{pneupcacomp} we can also see the recreated images from the $9$ components.
+Although it is difficult for the human eye to differentiate the negative from the positive in these images, there is clearly enough information left, given that the model achieves the test accuracy that it does.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=6in]{figure/pneupcacomp.png}}
+\caption{Instances from the Guangzhou dataset and their PCA reconstruction using only 9 principal components}
+\label{pneupcacomp}
+\end{figure}
 
@@
 Yet again impressed with the performance of using a simpler model, we attempt to perform the same classification without using neural networks at all.
-Using a PCA-RF model, we achieved a test accuracy of $83.7\%$.
-NEED TIME STATS FOR THE REST
+Using a PCA-RF model, we achieved a test accuracy of $84.9\%$.
+Although this is lower than the other models, this model was by far the fastest when only a CPU, being $94\%$ faster than the CNN and $50\%$ faster than the PCA-NN model (Table.~\ref{modcomppneu}).
+Like with the MNIST dataset, we see that we are able to trade accuracy for performance by using simpler hybrid models to make our prediction.
 
@@
 
-Below is table showing the accuracy and average prediction time for each model on the Guangzhou Pneumonia Dataset.
+Having optimized all our models, we are again able to compare them against eachother in terms of accuracy and performance.
+Looking at table~\ref{modcomppneu}, we see much of the same story as with the MNIST dataset.
+We see that the pure CNN is the optimal choice when accuracy is the most important factor, and the undisputed champion in terms of both accuracy and prediction time if a GPU is available.
+However, when looking at CPU-prediction times, the CNN is quite slow, and the PCA-NN model makes a strong case for trading a single percentage point of accuracy for an $88\%$ increase in speed.
+While they cannot beat the CNN in terms of accuracy, these simpler models are ideal an fit for embedded systems with real-time constraints.
 
@@
 \begin{center}
-\label{pcanntable}
+\label{modcomppneu}
 \begin{tabular}{| c | l | l | l |}
@@
 
-No child was saved. More children actually died due to our actions. Nevermind childern actually were saved.
+In this paper, we have compared CNNs to different hybrid models, in an attempt to explore the accuracy/performance-tradeoff of model complexity and prediction time.
+We have applied four different machine learning models on the MNIST and Guangzhou Pneumonia dataset, and compared their accuracy and prediction time against eachother.
+In our application of these models, we found that the CNN performed best in terms of accuracy on both datasets, achieving test accuracies of $99.0\%$ and $88.8\%$ on the MNIST and Pneumonia dataset, respectively.
+However, we also found that the PCA-NN model performed well, achieving accuracies of $98.1\%$ and $88.1\%$ on the same datasets.
+Furthermore, these results were achieved at a much lower computational cost, with the PCA-NN model being $88\%$ faster at making predictions (when not using a GPU).
+These results make a compelling case for using simpler models when using systems which are time constrained and lacking in dedicated machine learning hardware.
+To explore these findings further, one could look at other combinations of machine learning models, such as Gradient Boosted Trees or Support Vector Machines.
+Another option would be to use more effective architectures than AlexNet, such as DenseNet, which may achieve good accuracies without incurring long prediction times.
 
@@
 
-\section*{Appendix}
-
-Here are some pictures of Eric Reber, formerly known as the criminal Ed "Gurgler" Bickley of Texas State Penistentiary.
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{/home/jona/pictures/cat.jpg}}
-\caption{Eric Reber in his most devious form}
-\label{real1msetraintest}
-\end{figure}
-
 \end{document}

### 2022-12-09 src/CNN_models.py
Comments added to almost all files

@@ -1,0 +1,85 @@
+
+import tensorflow as tf
+import os
+from tensorflow.keras import layers
+from tensorflow.keras import regularizers
+from functools import partial
+
+"""
+This is a script similiar to utils.py containing the architectures mentioned by Geron in his book 
+'Hands On Machine Learning with SciKit-learn and Tensorflow" which we use throughout this project.
+"""
+
+
+# --------------------------------------------- Alex-net ------------------------------------------------------
+kernel = tf.keras.regularizers.L2(l2=0.001)
+bias = tf.keras.regularizers.L2(l2=0.001)
+
+model = tf.keras.Sequential(
+    [
+        #tf.keras.layers.Rescaling(1.0 / 255),
+        tf.keras.layers.Conv2D(
+            96, 11, strides=(4,4), padding='valid', activation="relu"), #kernel_regularizer=kernel, bias_regularizer=bias),
+
+        tf.keras.layers.MaxPooling2D(pool_size=(3,3), strides=2, padding='valid'),
+        tf.keras.layers.Conv2D(
+            256, 5, strides=(1,1), padding='same', activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias), 
+          
+        tf.keras.layers.MaxPooling2D(pool_size=(3,3), strides=2, padding='valid'), # S4
+        tf.keras.layers.Conv2D(
+            384, 3, strides=(1,1), padding='same', activation="relu"), #kernel_regularizer=kernel, bias_regularizer=bias), #C5
+        tf.keras.layers.Conv2D(
+            384, 3, strides=(1,1), padding='same', activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias), #C6
+        tf.keras.layers.Conv2D(
+            256, 3, strides=(1,1), padding='same', activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias), 
+        
+        tf.keras.layers.MaxPooling2D(pool_size=(3,3), strides=2, padding='valid'), 
+        tf.keras.layers.Flatten(),
+        #tf.keras.layers.Dropout(0.5),
+        tf.keras.layers.Dense(
+            4096, activation="relu"),# kernel_regularizer=kernel, bias_regularizer=bias
+        #),
+        #tf.keras.layers.Dropout(0.5),
+        tf.keras.layers.Dense(
+            4096, activation="relu"), #, kernel_regularizer=kernel, bias_regularizer=bias
+        #),
+        tf.keras.layers.Dense(1),
+    ]
+)
+
[…]

### 2022-12-09 src/conv_nn_pneu.py

@@ -5,13 +5,23 @@ from tensorflow.keras import regularizers
 import keras_tuner as kt
 from utils import *
 
+"""
+This file contains code for a CNN model using AlexNet architecture ( 
+won the ILSVRC challenge in 2012) which is applied to the Pneumonia dataset.
+The script performs outputs the final training and test accuracies, and plots 
+a confusion matrix after the model is finished predicting.
+"""
+
+
 filedir = os.path.dirname(__file__)
 from functools import partial
 
 seed = 1336
 tf.keras.utils.set_random_seed(seed)
+# Module from tensorflow enabling override of GPU settings
 tf.config.experimental.enable_op_determinism()
 
+# ------------------------------- Loading data ------------------------------------
 TRAINDIR = filedir + "/../data/chest_xray/train"
 TESTDIR = filedir + "/../data/chest_xray/test"
 BATCHSIZE = 256
@@ -47,7 +57,7 @@ test_ds = tf.keras.utils.image_dataset_from_directory(
     batch_size=BATCHSIZE,
     color_mode="grayscale",
 )
-
+# --------------------------- Adjusting class ratio ----------------------------
 COUNT_NORMAL = 1071
 COUNT_PNEUMONIA = 3114
 
@@ -61,7 +71,7 @@ AUTOTUNE = tf.data.AUTOTUNE
 train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
 test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
 
-
+# ------------------------------ Hyperparameter tuner ---------------------------------
 def model_builder(hp):
     hp_learning_rate = hp.Choice(
         "learning_rate", values=[2 * 1e-5, 2 * 1e-4, 2 * 1e-3, 2 * 1e-2, 2 * 1e-1]
@@ -146,8 +156,7 @@ tuner = kt.Hyperband(
 # best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
 learning_rate = 0.001  # best_hps.get("learning_rate")
 
-print(learning_rate)
-
+# --------------------------------- AlexNet CNN ----------------------------------
 model = tf.keras.Sequential(
     [
         tf.keras.layers.Rescaling(1.0 / 255),
@@ -211,7 +220,7 @@ model.compile(
     loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
[…]

### 2022-12-09 src/conv_rf_mnist.py

@@ -9,12 +9,17 @@ from utils import *
 import tensorflow_decision_forests as tfdf
 
 tf.keras.utils.set_random_seed(1336)
+tf.config.experimental.enable_op_determinism()
 """
-Convolutional random forest used for MNIST dataset. Builds and fits data.
+This script contains an implementation of a CNN model constructed using the 
+tensorflow API. This model is first applied to the MNIST dataset before 
+its output is fed into a Random Decision Forest as an attempt to improve upon 
+the results achieved solely by the CNN architecture used here. 
 """
 
+# ---------------------------------- Loading data ----------------------------------
 batch_size = 128
-epochs = 6
+epochs = 10
 mnist = tf.keras.datasets.mnist
 
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
@@ -34,6 +39,8 @@ test_ds = test_ds.batch(batch_size)
 width = 5
 learning_rate = 0.001
 
+
+# -------------------- Simple CNN with one Convolutional and Polling layer ------------------
 model = tf.keras.Sequential(
     [
         layers.Rescaling(1.0 / 255),
@@ -52,6 +59,7 @@ model.compile(
     metrics=["accuracy"],
 )
 
+# ----------------------- Saving best results from fitting stage ----------------------------
 checkpoint_filepath = "/tmp/checkpoint"
 model_checkpoint_callback = tf.keras.callbacks.ModelCheckpoint(
     filepath=checkpoint_filepath,
@@ -65,7 +73,7 @@ model.fit(
     train_ds,
     validation_data=test_ds,
     epochs=epochs,
-    # batch_size=batch_size,
+    batch_size=batch_size,
     callbacks=[model_checkpoint_callback],
 )
 
@@ -73,6 +81,7 @@ model.load_weights(checkpoint_filepath)
 
 model.evaluate(test_ds, batch_size=batch_size)
 
+# Extracting features (output) from the final layer of our CNN model
 feature_extractor = tf.keras.Model(
     inputs=model.inputs,
     outputs=model.layers[-2].output,
@@ -81,6 +90,7 @@ feature_extractor = tf.keras.Model(
 features_train = train_ds.map(lambda batch, label: (feature_extractor(batch), label))
 features_test = test_ds.map(lambda batch, label: (feature_extractor(batch), label))
 
+# Random Forest model
 forest = tfdf.keras.RandomForestModel(
     verbose=1,
[…]

### 2022-12-09 src/conv_rf_pneu.py

@@ -7,13 +7,23 @@ import keras_tuner as kt
 import tensorflow_decision_forests as tfdf
 from utils import *
 
+"""
+This script contains an implementation of a CNN model constructed using the 
+tensorflow API and utlizing the AlexNet architecture. This model is first applied 
+to the Pneumonia dataset before its output is fed into a Random Decision Forest 
+as an attempt to improve upon the results achieved solely by the CNN architecture 
+used here. 
+"""
+
 filedir = os.path.dirname(__file__)
 from functools import partial
 
 seed = 1336
 tf.keras.utils.set_random_seed(seed)
+# Module from tensorflow forcing GPU to run deterministically
 tf.config.experimental.enable_op_determinism()
 
+# ------------------------------- Loading data ------------------------------------
 TRAINDIR = filedir + "/../data/chest_xray/train"
 TESTDIR = filedir + "/../data/chest_xray/test"
 BATCHSIZE = 256
@@ -49,7 +59,7 @@ test_ds = tf.keras.utils.image_dataset_from_directory(
     batch_size=BATCHSIZE,
     color_mode="grayscale",
 )
-
+# --------------------------- Adjusting class ratio ----------------------------
 COUNT_NORMAL = 1071
 COUNT_PNEUMONIA = 3114
 
@@ -63,7 +73,7 @@ AUTOTUNE = tf.data.AUTOTUNE
 train_ds = train_ds.cache().prefetch(buffer_size=AUTOTUNE)
 test_ds = test_ds.cache().prefetch(buffer_size=AUTOTUNE)
 
-
+# ------------------------------ Hyperparameter tuner ---------------------------------
 def model_builder(hp):
     hp_learning_rate = hp.Choice(
         "learning_rate", values=[2 * 1e-5, 2 * 1e-4, 2 * 1e-3, 2 * 1e-2, 2 * 1e-1]
@@ -148,8 +158,7 @@ tuner.search(train_ds, epochs=10, validation_data=val_ds, class_weight=class_wei
 best_hps = tuner.get_best_hyperparameters(num_trials=1)[0]
 learning_rate = best_hps.get("learning_rate")
 
-print(learning_rate)
-
+# --------------------------------- AlexNet CNN ----------------------------------
 model = tf.keras.Sequential(
     [
         tf.keras.layers.Rescaling(1.0 / 255),
@@ -227,7 +236,7 @@ model.fit(
     train_ds,
     batch_size=BATCHSIZE,
[…]

### 2022-12-09 src/pca_rf_pneu.py

@@ -15,10 +15,14 @@ filedir = os.path.dirname(__file__)
 
 tf.keras.utils.set_random_seed(1336)
 """
-PCA neural network used for pneumonia dataset. Builds and fits data, takes time.
+This file contains a combination of the PCA algorithm from SciKit-learn 
+library used for feature extraction and dimensionality reduction, and 
+Random Decision Forest algorithm from Tensorflow API. The first n 
+principal components output by the PCA are fed as input to the Random 
+Forest model, and used for classification of images from the Pneumonia dataset. 
 """
 
-
+# -------------------------------- Hyperparameter tuner -----------------------------------
 def model_builder(hp):
     hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
     hp_learning_rate = hp.Choice("learning_rate", values=[1e-3, 1e-2, 1e-1])
@@ -58,13 +62,13 @@ def model_builder(hp):
 
     return model
 
-
+# ---------------------------------- Loading data -----------------------------------
 with tf.device("/cpu:0"):
     TRAINDIR = filedir + "/../data/chest_xray/train"
     TESTDIR = filedir + "/../data/chest_xray/test"
     batch_size = 128
-    IMG_HEIGHT = 200
-    IMG_WIDTH = 200
+    IMG_HEIGHT = 227
+    IMG_WIDTH = 227
 
     train_ds = tf.keras.utils.image_dataset_from_directory(
         TRAINDIR,
@@ -86,7 +90,7 @@ with tf.device("/cpu:0"):
 
     COUNT_NORMAL = 1071
     COUNT_PNEUMONIA = 3114
-
+# --------------------------------- Adjusting class ratio -------------------------------------
     weight_for_0 = (1 / COUNT_NORMAL) * (COUNT_NORMAL + COUNT_PNEUMONIA) / 2.0
     weight_for_1 = (1 / COUNT_PNEUMONIA) * (COUNT_NORMAL + COUNT_PNEUMONIA) / 2.0
 
@@ -103,28 +107,28 @@ with tf.device("/cpu:0"):
     X_test = None
     y_train = None
     y_test = None
-
+# ----------------------------------- Preproessing -------------------------------------------
     n_components = 9
 
     for batch, labels in train_ds:
         if X_train is None:
[…]

### 2022-12-09 src/conv_rf_mnist.py
quality control

@@ -24,8 +24,6 @@ mnist = tf.keras.datasets.mnist
 
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
 
-print(x_train.shape)
-
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
 
@@ -93,9 +91,7 @@ features_test = test_ds.map(lambda batch, label: (feature_extractor(batch), labe
 # Random Forest model
 forest = tfdf.keras.RandomForestModel(
     verbose=1,
-    # max_depth=25,
     random_seed=1337,
-    # num_trees=300,  # , tuner=tuner#, check_dataset=False
 )
 forest.fit(x=features_train)
 
@@ -104,9 +100,8 @@ forest.compile(metrics=["accuracy"])
 print(forest.evaluate(features_train, return_dict=True))
 print(forest.evaluate(features_test, return_dict=True))
 
+
 # timing:
-
-
 def predict():
     features_test = test_ds.map(lambda batch, label: (feature_extractor(batch), label))
     forest.predict(features_test)

### 2022-12-09 src/pca_nn_pneu.py

@@ -13,8 +13,8 @@ from sklearn.decomposition import PCA
 """
 This script contains an implementation of a neural network constructed by using the 
 tensorflow API which is then fed the first n principal components extracted through the 
-use of the SciKit-learns PCA implementation from X-Ray images of Pneumonia obesrved in children.
-The sript outputs the training and test accurcies achieved, and average time spen on prediction (time 
+use of the SciKit-learns randomized PCA implementation from X-Ray images of Pneumonia obesrved in children.
+The sript outputs the training and test accurcies achieved, and average time spent on prediction (time 
 spent on training exluded). 
 """
 
@@ -24,7 +24,7 @@ tf.keras.utils.set_random_seed(1336)
 # Module from tensorflow enabling override of GPU settings
 tf.config.experimental.enable_op_determinism()
 
-# ------------------------------------ Hyperparameter tuner  ------------------------------------  
+# ------------------------------------ Hyperparameter tuner  ------------------------------------
 def model_builder(hp):
     hp_lambda = hp.Choice("lambda", values=[1e-5, 1e-4, 1e-3])
     hp_learning_rate = hp.Choice("learning_rate", values=[1e-5, 1e-4, 1e-3, 1e-2, 1e-1])
@@ -65,11 +65,12 @@ def model_builder(hp):
 
     return model
 
+
 # ------------------------------------- Loading data -------------------------------------------
 with tf.device("/cpu:0"):
     TRAINDIR = filedir + "/../data/chest_xray/train"
     TESTDIR = filedir + "/../data/chest_xray/test"
-    BATCHSIZE = 256#128
+    BATCHSIZE = 256
     IMG_HEIGHT = 227
     IMG_WIDTH = 227
 
@@ -91,7 +92,7 @@ with tf.device("/cpu:0"):
         color_mode="grayscale",
     )
 
-# ----------------------------------- Preprocessing ---------------------------------------------
+    # ----------------------------------- Preprocessing ---------------------------------------------
     COUNT_NORMAL = 1071
     COUNT_PNEUMONIA = 3114
 
@@ -171,11 +172,11 @@ with tf.device("/gpu:0"):
     learning_rate = best_hps.get("learning_rate")
     units = best_hps.get("units")
 
-    print(lam)
-    print(learning_rate)
-    print(units)
+    print(f"{lam=}")
+    print(f"{learning_rate=}")
+    print(f"{units=}")
 
[…]

### 2022-12-09 src/pca_rf_mnist.py

@@ -11,7 +11,9 @@ from sklearn.decomposition import PCA
 
 tf.keras.utils.set_random_seed(1336)
 """
-PCA random forest used for MNIST dataset. Builds and fits data.
+PCA random forest used for MNIST dataset. Builds PCA Random Forest hybrid model using
+SciKit Learns randomized PCA and Tensorflow decision forest libraries. Trains and tests
+model.
 """
 
 batch_size = 128
@@ -21,11 +23,13 @@ mnist = tf.keras.datasets.mnist
 (x_train, y_train), (x_test, y_test) = mnist.load_data()
 x_train, x_test = x_train / 255.0, x_test / 255.0
 
+# components to use after PCA
 n_components = 50
 
 x_train_flat = tf.reshape(x_train, shape=[-1, 784])
 x_test_flat = tf.reshape(x_test, shape=[-1, 784])
 
+# define and fit PCA
 pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
 pca.fit(x_train_flat)
 
@@ -38,15 +42,13 @@ test_ds = tf.data.Dataset.from_tensor_slices((x_test_pca, y_test))
 train_ds = train_ds.batch(batch_size)
 test_ds = test_ds.batch(batch_size)
 
+# tune model
 tuner = tfdf.tuner.RandomSearch(num_trials=20)
 tuner.choice("max_depth", [5, 10, 15, 20, 25, 30])
 tuner.choice("min_examples", [5, 7, 9, 11, 13])
 
-
-# forest = tfdf.keras.RandomForestModel(max_depth=3, min_examples=9, check_dataset=False)
+# random forest
 forest = tfdf.keras.RandomForestModel(random_seed=1336, tuner=tuner, check_dataset=False)
-
-
 forest.fit(x=train_ds)
 
 forest.compile(metrics=["accuracy"])
@@ -61,6 +63,7 @@ def predict():
     test_ds = test_ds.batch(batch_size)
     forest.predict(test_ds)
 
+# time CPU runtime
 with tf.device("/cpu:0"):
     print("Timing prediction")
     timeit(predict)

### 2022-12-09 doc/paper.tex
Update paper

@@
 
-\title{Hybrid models as an alternative to traditional Convolutional Neural Networks for image classification}
+\title{Hybrid models as an alternative to Convolutional Neural Networks for image classification}
 
@@
 
-We compared convolutional neural networks, the de facto standard for image classification problems, with hybrid models created by combining neural networks with Principal Component Analysis (PCA) and Random Forests (RF).
+We compared convolutional neural networks, the de facto standard for image classification problems, with hybrid models created by combining neural networks with principal component analysis (PCA) and random forests (RF).
 By doing this, we explored the tradeoff between prediction accuracy and computational cost for four different ML models: A pure CNN model, a CNN-RF model, a PCA-NN model and a PCA-RF model.
-We used these models on the well known MNIST dataset and the Guangzhou Pneumonia Dataset consisting of x-ray images of potential cases of pneumonia.
+We used these models on the well known MNIST dataset and the Guangzhou pneumonia dataset consisting of x-ray images of potential cases of pneumonia.
 From this, we found that pure CNNs performed best in terms of accuracy, achieving test accuracies of $99.0\%$ and $88.8\%$ on the MNIST and pneumonia dataset, respectively.
-However, we found that the PCA-NN model performed similarly, at a much lower computational cost, achieving an impressive $88\%$ increase in prediction speed at a cost of only $1.1$ percentage points in accuracy.
+However, we found that the PCA-NN model performed similarly, at a much lower computational cost, achieving an impressive $88\%$ increase in prediction speed at the cost of only $1.1$ percentage points in accuracy.
 
@@
 
-The task of image recognition and classification is a great challenge in machine learning, as there are a large number of features present in the data which slow down our training.
+The task of image recognition and classification is a great challenge in machine learning, as there are a large number of features present in the data which slow down our training and prediction speed.
 For a relatively small one-channel image of $227$ by $227$ pixels, the naive approach of using every pixel as a feature becomes unfeasable, as we now have $51$ thousand features.
@@
 CNNs are able to reduce the dimensionality while extracting useful information in each layer, for example by finding edges in the first layer, combinations of edges that make a shape in the next layer, and combinations of shapes that resemble an object in the last layer.
-However, all these layers of convolution introduce significant computational costs, and CNNs are often reliant on Graphical Processing Units (GPUs) for efficient training and prediction.
+However, all these layers of convolution introduce significant computational costs, and CNNs are often reliant on graphical processing units (GPUs) for efficient training and prediction.
 In many embedded systems or mobile devices, the computational cost introduced by using deep CNNs may not be ideal, as the absence of dedicated hardware may lead to long prediction times.
 Furthermore, these systems may have real-time constraints, needing to make many predictions per second.
-In this paper, we explore alternative models for image classification, using Principal Component Analysis and Random Forests in combination with parts of the traditional CNN model.
+In this paper, we explore alternative models for image classification, using principal component analysis and random forests in combination with parts of the pure CNN model.
 We will investigate how these architectures perform for a simple image classification problem such as the MNIST dataset, as well as a more difficult classification problem in the form of pneumonia prediction from X-ray images.
@@
 More specifically, we will compare four different architectures.
-First we will use a traditional CNN, where a series convolutional layers is followed by a series fully connected layers.
-We will then modify our CNN to create a hybrid CNN-RF model, where the features extracted by the CNN are fed into a Random Forest, which makes the final prediction.
-Finally, we will make use of Principal Component Analysis instead of convolution and pooling for dimensionality reduction, feeding the results of this stage into either a Neural Network or a Random Forest.
+First we will use a pure CNN, where a series convolutional layers\footnote{We follow the terminology used in Goodfellow et Al.
+\cite[336]{gbc}, where what is denoted as a single convolutional layer includes the convolutional, detector and pooling stage} is followed by a series of fully connected layers.
+We will then modify our CNN to create a hybrid CNN-RF model, where the features extracted by the CNN are fed into a random forest, which makes the final prediction.
+Finally, we will make use of principal component analysis instead of convolution and pooling for dimensionality reduction, feeding the results of this stage into either a neural network or a random forest.
 In doing this, we hope to answer the following question: could hybrid models be a viable alternative to CNNs, either by directly outperforming them in accuracy, or by offering an appealing tradeoff between performance and computational effort?
@@
 -1 & -1 & -1
-\end{bmatrix}
+\end{bmatrix}.
 \end{equation}
@@
 
-The mathematical equation for the convolution of a kernel $w$ on an image $f(x,y)$ (represented as a discrete function in two variables) is as follows:
+The mathematical equation for the convolution of a kernel $w$ on an image $f(x,y)$ (represented as a discrete function in two variables) is as follows\footnote{In reality, we should have $x-s$ and $y-t$, as convolution involves rotating the kernel by 180 degrees.
+However, the term convolution is often used for both the mathematical operations of correlation and convolution\cite[160]{dip}.
+Furthermore, the kernels used in CNNs have randomly initialized weights, making rotation inconsequential.}:
 
 \begin{equation}
-w \star f(x,y) = \sum^a_{s=-a} \sum^b_{t=-b}f(x+s, y+t)\footnote{In reality, we should have $x-s$ and $y-t$, as correlation involves rotating the kernel by 180 degrees.
-However, the term convolution is often used for both the mathematical operations of correlation and convolution\cite[160]{dip}.
-Furthermore, the kernels used in CNNs have randomly initialized weights, making rotation inconsequential.}
+w \star f(x,y) = \sum^a_{s=-a} \sum^b_{t=-b}f(x+s, y+t).
 \end{equation}
 
-In a convolutional neural network, the values of each kernel is initialized randomly, and training involves tuning these values.
+In a convolutional neural network, the values of each kernel are initialized randomly, and training involves tuning these values.
 When performing convolution at the edges of an image, some values in the image will be undefined (i.e, some values of the kernel will not overlap with the image).
@@
 
-After several layers of convolution, the input has been usually been downsampled, and important features about the input have hopefully been extracted by use of convolution and pooling.
+After several layers of convolution, the input has usually been downsampled, and important features about the input have hopefully been extracted by the use of convolution and pooling.
 The output is then flattened, and the features are sent to one or more fully connected layers.
@@
 
-Here, the coloumns $U$ is known as the left singular vectors of $X$, $D$ is a diagonal matrix containing the singular values of $X$ (the square roots of the eigenvalues of $X^TX$) and the coloumns in $V$ are known as the right singular values.
+Here, the coloumns of $U$ are known as the left singular vectors of $X$, $D$ is a diagonal matrix containing the singular values of $X$ (the square roots of the eigenvalues of $X^TX$) and the coloumns of $V$ are known as the right singular values.
 The right singular values are actually the principal components we need, ordered by the variance in that direction.
 Thus, we can project our dataset down to a subspace spanned by a subset of these vectors (for example the first 3 vectors) and reduce the dimensionality of our dataset while losing as little information as possible.
-Furthermore, the coloumn vectors of $V$ are ortogonal, and as such we can project our dataset down to this subspace with a single matrix multiplication:
+Furthermore, the coloumn vectors of $V$ are orthogonal, and as such we can project our dataset down to this subspace with a single matrix multiplication:
 
@@
 A Random Forest is a type of ensemble method used to improve the performance of a single decision tree, by using bagging.
-Instead of using a single tree, a random forest consists of many similar trees, and the final class is determined majority voting by all the trees.
+Instead of using a single tree, a random forest consists of many similar trees, and the final class is determined by majority voting by all the trees.
 Like decision trees, they are relatively simple to understand.
@@
 To create a random forest, we perform a bootstrap of the original dataset, creating many similar instances with small permutations introduced by resampling with replacement.
-We now create one decision tree per bootstrap, only using a random subset of the total amount of features every time we perform a split\footnote{The number of features used is often set to $\sqrt{p}$}.
+We now create one decision tree per bootstrap, only using a random subset of the total amount of features every time we perform a split\footnote{The number of features used is often set to $\sqrt{p}$, where $p$ is the number of features in the dataset}.
 The result of this is that we create many slightly different trees.
@@
 For hyperparameter tuning, we used the Keras Tuner library \cite{kerastuner}.
+For the CNN-RF model, we extract the outputs of the final hidden layer of the CNN per instance in the dataset, and use this as inputs to the random forest.
 
@@
 
-To test our hybrid models against traditional methods, we applied them all to the MNIST dataset.
+To test our hybrid models against a pure CNN model, we applied them all to the MNIST dataset.
 
@@
 As is well known, convolutional neural networks perform excellent on the MNIST dataset, achieving test accuracies rivaling humans.
-As an initial guess for the architecture, we used only a single convolutional layer\footnote{We follow the terminology used in Goodfellow et Al.
-\cite[336]{gbc}, where what is denoted as a single convolutional layer includes the convolutional, detector and pooling stage} followed by a dense layer.
+As an initial guess for the architecture, we used only a single convolutional layer followed by a dense layer.
 The convolutional layer had $32$ filters.
@@
 Given that we have achieved excellent results by using a CNN, we may not expect to achieve better results by sending the extracted features to a random forest.
-Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of $98.1\%$, which is slightly worse than a pure CNN model.
+Regardless, we found that by using the same CNN model we found previously, and sending the extracted features into a random forest model, we achieved a testing accuracy of $98.9\%$, which is essentially equivalent to our pure CNN model.
 Furthermore, this method introduced a significant computational cost to the prediction, as can be seen from table~\ref{modcompmnist}.
@@
 After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $98.1\%$.
-This is comparable to the accuracy obtained by a traditional convolutional neural network.
+This was achieved with $224$ nodes per layer and a learning rate of $0.001$.
+These results are comparable to the accuracy obtained by a convolutional neural network.
 Looking at figure~\ref{mnistheatmap_pcann}, we again see that we are making accurate predictions across all classes.
@@
 
+\begin{table}[H]
+\caption{Optimal hyperparameters for the PCA-NN model}
+\begin{center}
+\label{pcanntable}
+\begin{tabular}{| l | l | l | l |}
+\hline
+Principal Components & Nodes per hidden layer & Learning Rate & Lambda \\
+\hline
+9 & 9 & 0.1 & 0.0001 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{figure}[H]
@@
 \caption{Confusion matrix for MNIST dataset classified using PCA followed by a neural network.
-Here we can see that misclassifications are slightly more common than when using a traditional CNN, but the results are still good.}
+Here we can see that misclassifications are slightly more common than when using a CNN, but the results are still good.}
 \label{mnistheatmap_pcann}
@@
 Therefore, we initially had problems designing an architecture which achieved much higher than $80\%$ test accuracy.
-Clearly, we are not able to use a simple one convolutional layer model like we were able to on the MNIST dataset, and designing a general convolutional neural network architecture is a very complex task.
+Clearly, we are not able to use a simple one model consisting of only one convolutional layer, like we were able to on the MNIST dataset, and designing a general convolutional neural network architecture is a very complex task.
 Thus, we turn to the literature, and explore models which have been proven to be effective at image recognition.
@@
 Comparing our result to those obtained by Kermany et Al.
-($92.8\%$), we see that our model is a few percentage points worse, however when considering that we are not using transfer learning and are completing training on a single laptop, we consider our results to be reasonably good.
+($92.8\%$), we see that our model is a few percentage points worse.
+Looking at the confusion matrix (Fig.~\ref{conf_cnn_pneu}), we can also see that our model is biased, being more prone to create false positives than false negatives.
+This is due to the imbalance of the training dataset, which has more cases of pneumonia than not.
+We have used class weighting when training our model, which greatly improves upon the accuracy without weighting, however even with this, we are not able to remove the bias completely\footnote{We have been able to make the model less biased by arbitrarily adjusting the class weights, however this hurts the overall accuracy}.
+All in all, the model is worse than the one created by Kermany et Al., but when considering that we are not using transfer learning and are completing training on a single laptop, we consider our results to be reasonable.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=5in]{figure/conf_cnn_pneu.png}}
+\caption{Confusion matrix for the pneumonia dataset, using the CNN model}
+\label{conf_cnn_pneu}
+\end{figure}
 
@@
 Here, we achieved a test accuracy of $88.1\%$, which is $0.7\%$ lower than simply using the CNN as is to make the prediction.
-Furthermore, using a random forests increases the computational complexity considerably as well, as can be seen from Table~\ref{modcomppneu}.
+Furthermore, using a random forest increases the computational complexity considerably as well, as can be seen from Table~\ref{modcomppneu}.
 
@@
 
-Again we then attempt to replace the feature extraction stage of a CNN with a simple application of Principal Component Analysis.
+Again, we then attempt to replace the feature extraction stage of a CNN with a simple application of Principal Component Analysis.
 Given the complexity of the network needed to achieve good results when using a CNN, one might not expect to find good results using PCA.
@@
 This is an impressive accuracy considering the simplicity of our model, and shows that it is possible to achieve reasonable results without turning to complex convolutional neural networks.
-Furthermore, this small decrease in accuracy came with a considerable gain in prediction speed when predicting using only the CPU, with the PCA-NN model being $88\%$ faster than the CNN on average (Table.~\ref{modcomppneu}).
+Looking at the confusion matrix, we again see a bias towards classifying images as pneumonia (Fig.~\ref{conf_pca_nn_pneu}).
+Regardless of the bias, we can see that the small decrease in accuracy compared to the CNN model came with a considerable gain in prediction speed when predicting using only the CPU, with the PCA-NN model being $88\%$ faster than the CNN on average (Table.~\ref{modcomppneu}).
 Of course, prediction of pneumonia is not a problem where one would trade accuracy for prediction speed, but there are many real-time use cases for image recognition where this is a tradeoff that is preferable.
@@
 
-The fact that only $9$ components where needed is quite surprising, but could be explained by the homogeneity of the dataset.
+\begin{figure}[H]
+\centerline{\includegraphics[width=5in]{figure/conf_pca_nn_pneu.png}}
+\caption{Confusion matrix for the pneumonia dataset, using the PCA-NN model}
+\label{conf_pca_nn_pneu}
+\end{figure}
+
+The fact that only $9$ components were needed is quite surprising, but could be explained by the homogeneity of the dataset.
 Unlike the MNIST dataset, where each class is very different, both the positive and negative classes of pneumonia have images which are very similar, as can be seen in figure~\ref{pneupcacomp}.
@@
 
-\subsubsection{Pneumonia recap}
+\subsubsection{Pneumonia model comparison}
 
@@
 However, when looking at CPU-prediction times, the CNN is quite slow, and the PCA-NN model makes a strong case for trading a single percentage point of accuracy for an $88\%$ increase in speed.
-While they cannot beat the CNN in terms of accuracy, these simpler models are ideal an fit for embedded systems with real-time constraints.
+While they cannot beat the CNN in terms of accuracy, these simpler models are an ideal fit for embedded systems with real-time constraints.
 
@@
 However, we also found that the PCA-NN model performed well, achieving accuracies of $98.1\%$ and $88.1\%$ on the same datasets.
-Furthermore, these results were achieved at a much lower computational cost, with the PCA-NN model being $88\%$ faster at making predictions (when not using a GPU).
+Furthermore, these results were achieved at a much lower computational cost, with the PCA-NN model being $34\%$ and $88\%$ faster at making predictions for each dataset (when not using a GPU).
 These results make a compelling case for using simpler models when using systems which are time constrained and lacking in dedicated machine learning hardware.

### 2022-12-11 src/conv_nn_mnist.py
Add code for additional metrics
It never ends...

@@ -22,7 +22,10 @@ batch_size = 128
 epochs = 10
 mnist = tf.keras.datasets.mnist
 
-(x_train, y_train), (x_test, y_test) = mnist.load_data()
+(x_train, y_train_integer), (x_test, y_test_integer) = mnist.load_data()
+
+y_train = tf.one_hot(y_train_integer, 10)
+y_test = tf.one_hot(y_test_integer, 10)
 
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
@@ -47,7 +50,7 @@ def model_builder(hp):
 
     model.compile(
         optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
-        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+        loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
         metrics=["accuracy"],
     )
 
@@ -84,8 +87,12 @@ model = tf.keras.Sequential(
 
 model.compile(
     optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
-    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
-    metrics=["accuracy"],
+    loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
+    metrics=[
+        "accuracy",
+        tf.keras.metrics.Precision(top_k=1),
+        tf.keras.metrics.Recall(thresholds=0),
+    ],
 )
 
 checkpoint_filepath = "/tmp/checkpoint"
@@ -117,7 +124,7 @@ with tf.device("/cpu:0"):
     timeit(model.predict, (x_test, y_test), batch_size=batch_size)
 
 predictions = tf.math.argmax(model.predict(x_test), axis=1)
-conf = conf_mat(predictions, y_test, num_cls=10)
+conf = conf_mat(predictions, y_test_integer, num_cls=10)
 conf = perc(conf)
 
 plot_confusion(conf, title="Confusion matrix - CNN - MNIST")

### 2022-12-11 src/conv_nn_pneu.py

@@ -205,7 +205,11 @@ model = tf.keras.Sequential(
 model.compile(
     optimizer=tf.keras.optimizers.Adam(learning_rate=2 * 10e-6),
     loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
-    metrics=["accuracy"],
+    metrics=[
+        "accuracy",
+        tf.keras.metrics.Precision(top_k=1),
+        tf.keras.metrics.Recall(thresholds=0),
+    ],
 )
 # Saves the weights from the best trainig epoch
 checkpoint_filepath = "/tmp/checkpoint"

### 2022-12-11 src/pca_nn_mnist.py

@@ -24,8 +24,11 @@ on the test set.
 # --------------------------------------- Loading data --------------------------------------------
 mnist = tf.keras.datasets.mnist
 
-(x_train, y_train), (x_test, y_test) = mnist.load_data()
+(x_train, y_train_integer), (x_test, y_test_integer) = mnist.load_data()
 x_train, x_test = x_train / 255.0, x_test / 255.0
+
+y_train = tf.one_hot(y_train_integer, 10)
+y_test = tf.one_hot(y_test_integer, 10)
 
 n_components = 50
 
@@ -37,7 +40,8 @@ pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
 pca.fit(x_train_flat)
 
 x_train_pca = pca.transform(x_train_flat)
-# ------------------------------------ Hyperparameter tuner  ------------------------------------ 
+# ------------------------------------ Hyperparameter tuner  ------------------------------------
+
 
 def model_builder(hp):
     hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
@@ -54,11 +58,12 @@ def model_builder(hp):
 
     model.compile(
         optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
-        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+        loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
         metrics=["accuracy"],
     )
 
     return model
+
 
 # --------------------------------- Tuning hyperparameters --------------------------------------
 tuner = kt.Hyperband(
@@ -95,10 +100,14 @@ model = tf.keras.Sequential(
 
 model.compile(
     optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
-    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
-    metrics=["accuracy"],
+    loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
+    metrics=[
+        "accuracy",
+        tf.keras.metrics.Precision(top_k=1),
+        tf.keras.metrics.Recall(thresholds=0),
+    ],
 )
-# ----------------------------------- Fitting the model -------------------------------------- 
+# ----------------------------------- Fitting the model --------------------------------------
 model.fit(
     x_train_pca,
     y_train,
[…]

### 2022-12-11 doc/paper.tex
Add precision and recall dummy numbers

@@
 With these hyperparameters we achieved a test accuracy of $99.0\%$, a very good result.
+Furthermore, we acieved a recall of NUMBER and precision of NUMBER.
 Based on this result, we conclude that no more hidden layers or regularization parameters need to be included.
@@
 Using different amount of principal components, we sent the reduced dataset into a neural network with three hidden layers.
-After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $98.1\%$.
+After testing different amounts of principal components to keep, and tuning the hyperparameters based on these, we found that even with only $50$ principal components, we achieved a test accuracy of $98.1\%$, a recall of NUMBER and a precision of NUMBER.
 This was achieved with $224$ nodes per layer and a learning rate of $0.001$.
@@
 AlexNet is both a relatively simple model to train using our available hardware, and a simple model to understand.
-Using AlexNet with the Adam optimizer, we achieved a testing accuracy of $88.9\%$.
+Using AlexNet with the Adam optimizer, we achieved a testing accuracy of $88.9\%$, a recall of NUMBER and a precision of NUMBER.
 Comparing our result to those obtained by Kermany et Al.
@@
 This is due to the imbalance of the training dataset, which has more cases of pneumonia than not.
-We have used class weighting when training our model, which greatly improves upon the accuracy without weighting, however even with this, we are not able to remove the bias completely\footnote{We have been able to make the model less biased by arbitrarily adjusting the class weights, however this hurts the overall accuracy}.
+We have used class weighting when training our model, which greatly improves upon the accuracy gained without weighting, however even with this, we are not able to remove the bias completely\footnote{That is, we have not been able to remove the bias without also severely decreasing the accuracy}.
 All in all, the model is worse than the one created by Kermany et Al., but when considering that we are not using transfer learning and are completing training on a single laptop, we consider our results to be reasonable.
@@
 Tuning the model, we found that optimal hyperparameters were as described by Table~\ref{pcanntable}.
-Using these parameters, we were able to achieve a test accuracy of $87.8\%$.
+Using these parameters, we were able to achieve a test accuracy of $87.8\%$, with a recall of NUMBER and a precision of NUMBER.
 This is an impressive accuracy considering the simplicity of our model, and shows that it is possible to achieve reasonable results without turning to complex convolutional neural networks.

### 2022-12-12 src/conv_nn_pneu.py
Fix precision

@@ -207,7 +207,7 @@ model.compile(
     loss=tf.keras.losses.BinaryCrossentropy(from_logits=True),
     metrics=[
         "accuracy",
-        tf.keras.metrics.Precision(top_k=1),
+        tf.keras.metrics.Precision(thresholds=0),
         tf.keras.metrics.Recall(thresholds=0),
     ],
 )

### 2022-12-12 src/conv_nn_mnist.py
Restore old files

@@ -22,10 +22,7 @@ batch_size = 128
 epochs = 10
 mnist = tf.keras.datasets.mnist
 
-(x_train, y_train_integer), (x_test, y_test_integer) = mnist.load_data()
-
-y_train = tf.one_hot(y_train_integer, 10)
-y_test = tf.one_hot(y_test_integer, 10)
+(x_train, y_train), (x_test, y_test) = mnist.load_data()
 
 x_train = tf.reshape(x_train, shape=[-1, 28, 28, 1])
 x_test = tf.reshape(x_test, shape=[-1, 28, 28, 1])
@@ -50,7 +47,7 @@ def model_builder(hp):
 
     model.compile(
         optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
-        loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
+        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
         metrics=["accuracy"],
     )
 
@@ -87,12 +84,8 @@ model = tf.keras.Sequential(
 
 model.compile(
     optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
-    loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
-    metrics=[
-        "accuracy",
-        tf.keras.metrics.Precision(top_k=1),
-        tf.keras.metrics.Recall(thresholds=0),
-    ],
+    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+    metrics=["accuracy"],
 )
 
 checkpoint_filepath = "/tmp/checkpoint"
@@ -124,7 +117,7 @@ with tf.device("/cpu:0"):
     timeit(model.predict, (x_test, y_test), batch_size=batch_size)
 
 predictions = tf.math.argmax(model.predict(x_test), axis=1)
-conf = conf_mat(predictions, y_test_integer, num_cls=10)
+conf = conf_mat(predictions, y_test, num_cls=10)
 conf = perc(conf)
 
 plot_confusion(conf, title="Confusion matrix - CNN - MNIST")

### 2022-12-12 src/pca_nn_mnist.py

@@ -24,11 +24,8 @@ on the test set.
 # --------------------------------------- Loading data --------------------------------------------
 mnist = tf.keras.datasets.mnist
 
-(x_train, y_train_integer), (x_test, y_test_integer) = mnist.load_data()
+(x_train, y_train), (x_test, y_test) = mnist.load_data()
 x_train, x_test = x_train / 255.0, x_test / 255.0
-
-y_train = tf.one_hot(y_train_integer, 10)
-y_test = tf.one_hot(y_test_integer, 10)
 
 n_components = 50
 
@@ -40,8 +37,7 @@ pca = PCA(n_components=n_components, svd_solver="randomized", random_state=1336)
 pca.fit(x_train_flat)
 
 x_train_pca = pca.transform(x_train_flat)
-# ------------------------------------ Hyperparameter tuner  ------------------------------------
-
+# ------------------------------------ Hyperparameter tuner  ------------------------------------ 
 
 def model_builder(hp):
     hp_units = hp.Int("units", min_value=32, max_value=512, step=32)
@@ -58,12 +54,11 @@ def model_builder(hp):
 
     model.compile(
         optimizer=tf.keras.optimizers.Adam(learning_rate=hp_learning_rate),
-        loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
+        loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
         metrics=["accuracy"],
     )
 
     return model
-
 
 # --------------------------------- Tuning hyperparameters --------------------------------------
 tuner = kt.Hyperband(
@@ -100,14 +95,10 @@ model = tf.keras.Sequential(
 
 model.compile(
     optimizer=tf.keras.optimizers.Adam(learning_rate=learning_rate),
-    loss=tf.keras.losses.CategoricalCrossentropy(from_logits=True),
-    metrics=[
-        "accuracy",
-        tf.keras.metrics.Precision(top_k=1),
-        tf.keras.metrics.Recall(thresholds=0),
-    ],
+    loss=tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
+    metrics=["accuracy"],
 )
-# ----------------------------------- Fitting the model --------------------------------------
+# ----------------------------------- Fitting the model -------------------------------------- 
 model.fit(
     x_train_pca,
     y_train,
[…]

### 2022-12-12 doc/paper.tex
Update report with precision and recall

--- 2022-12-09 doc/paper.tex
@@
 AlexNet is both a relatively simple model to train using our available hardware, and a simple model to understand.
-Using AlexNet with the Adam optimizer, we achieved a testing accuracy of $88.9\%$.
+Using AlexNet with the Adam optimizer, we achieved a testing accuracy of $88.9\%$, a recall of $0.9359$ and a precision of $0.8838$.
 Comparing our result to those obtained by Kermany et Al.
-($92.8\%$), we see that our model is a few percentage points worse.
+($92.8\%$), we see that our model is a few percentage points worse in terms of testing accuracy.
 Looking at the confusion matrix (Fig.~\ref{conf_cnn_pneu}), we can also see that our model is biased, being more prone to create false positives than false negatives.
 This is due to the imbalance of the training dataset, which has more cases of pneumonia than not.
-We have used class weighting when training our model, which greatly improves upon the accuracy without weighting, however even with this, we are not able to remove the bias completely\footnote{We have been able to make the model less biased by arbitrarily adjusting the class weights, however this hurts the overall accuracy}.
+We have used class weighting when training our model, which greatly improves upon the accuracy gained without weighting, however even with this, we are not able to remove the bias completely\footnote{That is, we have not been able to remove the bias without also severely decreasing the accuracy}.
 All in all, the model is worse than the one created by Kermany et Al., but when considering that we are not using transfer learning and are completing training on a single laptop, we consider our results to be reasonable.
@@
 Tuning the model, we found that optimal hyperparameters were as described by Table~\ref{pcanntable}.
-Using these parameters, we were able to achieve a test accuracy of $87.8\%$.
+Using these parameters, we were able to achieve a test accuracy of $87.8\%$, with a recall of $0.9103$ and a precision of $0.8703$.
 This is an impressive accuracy considering the simplicity of our model, and shows that it is possible to achieve reasonable results without turning to complex convolutional neural networks.

### 2022-12-12 doc/paper.tex
Update abstract

@@
 From this, we found that pure CNNs performed best in terms of accuracy, achieving test accuracies of $99.0\%$ and $88.8\%$ on the MNIST and pneumonia dataset, respectively.
-However, we found that the PCA-NN model performed similarly, at a much lower computational cost, achieving an impressive $88\%$ increase in prediction speed at the cost of only $1.1$ percentage points in accuracy.
+However, we found that the PCA-NN model performed similarly, at a much lower computational cost, achieving increases of $34$ and $88$ percent in prediction speed on each dataset, without losing more than $1.1$ percentage point in testing accuracy in both instances.
 
